### 2011-03-22 aux/material_exp2/Pics/Sequences/GenauEin-Einige/seguence.tex

\documentclass{article}

\usepackage[palatino]{mypackages}\usepackage{mycommands}

\usepackage{verbatim}
\usepackage[active,tightpage]{preview}
\PreviewEnvironment{tikzpicture}
\setlength\PreviewBorder{5pt}

\begin{document}

\begin{tikzpicture}[rounded corners]

  \node at (6,18) {\Huge{Genau einer der Briefe ist mit einigen seiner Kreise verbunden.}};

  \node at (0,0) {\includegraphics{105_pdfsam_Brief-Kreise-Blau.pdf}};

  \node at (12,0) {\includegraphics{105_pdfsam_Brief-Kreise-Blau.pdf}};

  \node at (12,12) {\includegraphics{105_pdfsam_Brief-Kreise-Blau.pdf}};

  \node at (0,12) {\includegraphics{105_pdfsam_Brief-Kreise-Blau.pdf}};

  \node[draw=black!80,fill=blue!30,ultra thick] at (0,-6)
  {\Huge{Wahr}};

  \node[draw=black!80,fill=blue!30,ultra thick] at (6,-6)
  {\Huge{Falsch}};

  \node[draw=black!80,fill=blue!30,ultra thick] at (12,-6) {\Huge{Mehr Info}};

\end{tikzpicture}

\begin{tikzpicture}[rounded corners]

  \node at (6,18) {\Huge{Genau einer der Briefe ist mit einigen seiner Kreise verbunden.}};

  \node at (0,0) {\includegraphics{105_pdfsam_Brief-Kreise-Blau.pdf}};

  \node at (12,0) {\includegraphics{105_pdfsam_Brief-Kreise-Blau.pdf}};

  \node at (12,12) {\includegraphics{72_pdfsam_Brief-Kreise-Blau.pdf}};

  \node at (0,12) {\includegraphics{105_pdfsam_Brief-Kreise-Blau.pdf}};

  \node[draw=black!80,fill=blue!30,ultra thick] at (0,-6)
  {\Huge{Wahr}};

  \node[draw=black!80,fill=blue!30,ultra thick] at (6,-6)
  {\Huge{Falsch}};

  \node[draw=black!80,fill=blue!30,ultra thick] at (12,-6) {\Huge{Mehr Info}};

\end{tikzpicture}

\begin{tikzpicture}[rounded corners]

  \node at (6,18) {\Huge{Genau einer der Briefe ist mit einigen seiner Kreise verbunden.}};

  \node at (0,0) {\includegraphics{105_pdfsam_Brief-Kreise-Blau.pdf}};

  \node at (12,0) {\includegraphics{105_pdfsam_Brief-Kreise-Blau.pdf}};

  \node at (12,12) {\includegraphics{64_pdfsam_Brief-Kreise-Blau.pdf}};

  \node at (0,12) {\includegraphics{105_pdfsam_Brief-Kreise-Blau.pdf}};

  \node[draw=black!80,fill=blue!30,ultra thick] at (0,-6)
  {\Huge{Wahr}};

  \node[draw=black!80,fill=blue!30,ultra thick] at (6,-6)
  {\Huge{Falsch}};

  \node[draw=black!80,fill=blue!30,ultra thick] at (12,-6) {\Huge{Mehr Info}};

\end{tikzpicture}

\begin{tikzpicture}[rounded corners]

  \node at (6,18) {\Huge{Genau einer der Briefe ist mit einigen seiner Kreise verbunden.}};

  \node at (0,0) {\includegraphics{31_pdfsam_Brief-Kreise-Blau.pdf}};

  \node at (12,0) {\includegraphics{105_pdfsam_Brief-Kreise-Blau.pdf}};

  \node at (12,12) {\includegraphics{64_pdfsam_Brief-Kreise-Blau.pdf}};

  \node at (0,12) {\includegraphics{105_pdfsam_Brief-Kreise-Blau.pdf}};

  \node[draw=black!80,fill=blue!30,ultra thick] at (0,-6)
  {\Huge{Wahr}};

  \node[draw=black!80,fill=blue!30,ultra thick] at (6,-6)
  {\Huge{Falsch}};

  \node[draw=black!80,fill=blue!30,ultra thick] at (12,-6) {\Huge{Mehr Info}};

\end{tikzpicture}

\begin{tikzpicture}[rounded corners]

  \node at (6,18) {\Huge{Genau einer der Briefe ist mit einigen seiner Kreise verbunden.}};

  \node at (0,0) {\includegraphics{31_pdfsam_Brief-Kreise-Blau.pdf}};

  \node at (12,0) {\includegraphics{64_pdfsam_Brief-Kreise-Blau.pdf}};

  \node at (12,12) {\includegraphics{64_pdfsam_Brief-Kreise-Blau.pdf}};

  \node at (0,12) {\includegraphics{64_pdfsam_Brief-Kreise-Blau.pdf}};

  \node[draw=black!80,fill=blue!30,ultra thick] at (0,-6)
  {\Huge{Wahr}};

  \node[draw=black!80,fill=blue!30,ultra thick] at (6,-6)
  {\Huge{Falsch}};

  \node[draw=black!80,fill=blue!30,ultra thick] at (12,-6) {\Huge{Mehr Info}};

\end{tikzpicture}

\end{document}