### 2012-05-21 text/paper/paper.tex
paper svnInfo fiddling

\documentclass[fleqn,reqno,10pt]{article}

\usepackage{myarticlestyledefault}

\usepackage{mypackages}
\usepackage{mycommands}
\usepackage{myenvironments}

\usepackage[]{svninfo}

\newcommand{\lit}{\acro{lit}}
\newcommand{\glb}{\acro{glb}}
\newcommand{\loc}{\acro{loc}}

\newcommand{\as}{\acro{as}}
\renewcommand{\es}{\acro{es}}

\newcommand{\lc}{\acro{lc}}
\newcommand{\ec}{\acro{ec}}

\title{Scalar Items in Embedded Position: {A}nother Experimental Approach}
\author{Fabian Schlotterbeck, Michael Franke and Petra Augurzky}
\date{}

\begin{document}
\maketitle

\begin{abstract}
  \dots make concrete \dots
\end{abstract}

\svnInfo $Id$

\section{Introduction}
\label{sec:introduction}

This paper deals with the interpretation of just two types of
sentences, namely what we will call \as and \es sentences, mnemonic
for \emph{all \dots some \dots} and \emph{exactly one \dots some
  \dots}.

\section{Get to know your readings}
\label{sec:get-know-your}

\section{Theories and predictions}
\label{sec:theories-predictions}

\section{Previous studies}
\label{sec:previous-studies}

\printbibliography[heading=bibintoc]

\end{document}