### 2012-05-21 text/paper/paper.tex
paper sec com

\documentclass[fleqn,reqno,10pt]{article}

\usepackage{myarticlestyledefault}

\usepackage{mypackages}
\usepackage{mycommands}
\usepackage{myenvironments}

\usepackage[]{svninfo}

\newcommand{\lit}{\acro{lit}}
\newcommand{\glb}{\acro{glb}}
\newcommand{\loc}{\acro{loc}}

\newcommand{\as}{\acro{as}}
\renewcommand{\es}{\acro{es}}
\renewcommand{\AE}{\as}
\newcommand{\GE}{\es}

\newcommand{\lc}{\acro{lc}}
\newcommand{\ec}{\acro{ec}}

\DefineNamedColor{named}{mycol}{cmyk}{0.6,0.6,0,0}
\DefineNamedColor{named}{mygray}{cmyk}{0.05,0.05,0.05,0.05}
\DefineNamedColor{named}{mygraylight}{cmyk}{0.017,0.017,0.017,0.017}

\newcommand{\mymark}[1]{{\color{mycol}{#1}}}

\title{Scalar Items in Embedded Position: {A}nother Experimental Approach}
\author{Fabian Schlotterbeck, Michael Franke and Petra Augurzky}
\date{}

\begin{document}
\maketitle

\begin{abstract}
  \dots make concrete \dots
\end{abstract}

\svnInfo $Id$

\section{Introduction}
\label{sec:introduction}

This paper deals with the interpretation of two types of
sentences, exemplified by:

\begin{exe}
\ex \label{bsp:AE}
  \mymark{All} of the students read \mymark{some} of the
  papers. \hfill (\AE)
\ex \label{bsp:GE} 
  \mymark{Exactly one} of the students read \mymark{some} of the
  papers. \hfill (\GE)
\end{exe}

\noindent In \as-sentences a scalar item \emph{some} takes scope under
a universal quantifier \emph{all}. In \es-sentences a scalar item
\emph{some} takes scope under a non-monotonic quantifier \emph{exactly
  one}. Scalar \emph{some} is usually assumed to receive a semantic
interpretation similar to logical $\exists$, so that the expression
\emph{Some boys cried.} is literally true in a situation where all
boys cried. But it is also known to invite upper-bounding inferences
in plain utterances:

\begin{exe}
  \ex \label{bsp:Plain-SI}
    \begin{xlist}
      \ex \label{bsp:Plain-SI-Target} Hans solved some of the problems.
      \ex \label{bsp:Plain-SI-Implicature} $\implicates$ Hans solved some but not all of the problems.
    \end{xlist}
\end{exe}

\noindent The classical explanation of this inference, following the
pioneering work of \citet{Grice1975:Logic-and-Conve}, has it that
(\ref{bsp:Plain-SI-Implicature}) is a pragmatic inference, a so-called
\emph{quantity implicature}, derived by an abductive inference as the
best explanation of why am informed, knowledgable and cooperative
speaker has uttered (\ref{bsp:Plain-SI-Target}) when she could also
have uttered the semantically stronger and relevant:\dn{redo labels of
  examples}

\begin{exe}
  \exr{bsp:Plain-SI}
    \begin{xlist}
      \exi{c.} \label{bsp:Plain-SI-Alternative} Hans solved all of the problems.
    \end{xlist}
\end{exe}

\noindent If this upper-bounding inference would occur often and
systematically enough, then it may well be that also embedded
occurrences of \emph{some} get enriched, in some fashion or other, to
contribute \emph{in situ} the enriched meaning \emph{some but not
  all}. Indeed, even \citeauthor{Grice1975:Logic-and-Conve} envisaged
this possibility when he wrote: ``It may not be impossible for what
starts life, so to speak, as a conversational implicature to become
conventionalized'' \citep[p.58]{Grice1975:Logic-and-Conve}. But then
there are at least three relevant candidate readings for \as- and
\es-sentences: (i) a \emph{literal reading} like in
(\ref{bsp:AE-Literal}) and (\ref{bsp:GE-Literal}) where \emph{some}
has only its literal meaning as in; (ii) a \emph{global reading} like
in (\ref{bsp:AE-Literal}) and (\ref{bsp:GE-Literal}) where, according
to the Gricean intuition, we compare utterances of (\ref{bsp:AE}) or
(\ref{bsp:GE}) with utterances of the corresponding sentences
(\ref{bsp:AE-Alternative}) and (\ref{bsp:GE-Alternative}) where
\emph{some} is replaced by \emph{all}; and also (iii) a \emph{local
  reading} like in (\ref{bsp:AE-Local}) and (\ref{bsp:GE-Local}) where
\emph{some} is read to mean \emph{some but not all} in the scope of
the embedding quantifier.

% DIRTY!!!
\setcounter{exx}{0}

\begin{exe}

  \ex \mymark{All} of the students read {\mymark{some}} of the
  papers.

  \begin{xlist}
  \ex \label{bsp:AE-Literal} \mymark{All} of the students read
    {\mymark{some and maybe all}} of the papers. \hfill (\as-\lit)
  \ex \label{bsp:AE-Global}
    \mymark{All} of the students read \mymark{some} 
    and  \hfill (\as-\glb)\\
    \mymark{not all} of the students read \mymark{all} of the papers.
  \ex \label{bsp:AE-Local}
    \mymark{All} of the students read {\mymark{some  but not all}} of the
    papers. \hfill (\as-\loc)
  \end{xlist}

\ex \mymark{Exactly one} of the students read {\mymark{some}} of the
  papers.

  \begin{xlist}
  \ex \label{bsp:GE-Literal} \mymark{Exactly one} of the students read
    {\mymark{some and maybe all}} of the papers. \hfill (\es-\lit)
  \ex \label{bsp:GE-Global}
    \mymark{Exactly one} of the students read \mymark{some} 
    and  \hfill (\es-\glb)\\
    \mymark{not all} of the students read \mymark{all} of the papers.
  \ex \label{bsp:GE-Local}
    \mymark{Exactly one} of the students read {\mymark{some  but not all}} of the
    papers. \hfill (\es-\loc)
  \end{xlist}

\ex \label{bsp:AE-Alternative} \mymark{All} of the students read
  {\mymark{all}} of the papers.

\ex \label{bsp:GE-Alternative} \mymark{Exactly one} of the students
  read {\mymark{all}} of the papers.

\end{exe}

\bigskip

\section{Get to know your readings}
\label{sec:get-know-your}

\section{Theories and predictions}
\label{sec:theories-predictions}

\section{Previous studies}
\label{sec:previous-studies}

\printbibliography[heading=bibintoc]

\end{document}