### 2013-01-21 text/paper/paper3.tex
paper3 - up to general discussion

\documentclass[fleqn,reqno,10pt,draft]{article}

\usepackage[natbib=true,
            style=authoryear-comp,
            backend=bibtex,
            sortcites=false,
            maxnames=2,
            doi=false,
            url=false]{biblatex}
% \bibliography{../helpers/MyRefGlobal}
\bibliography{paper}

\usepackage[final,            % override "draft" which means "no nothing"
            colorlinks,       % rather than outlining them in boxes
            linkcolor=gray,   % override truly awful color choices
            citecolor=gray,   %   (ditto)
            urlcolor=gray,    %   (ditto)
            plainpages=false, % to overcome complaints with multiple
            pdfpagelabels,    % multiple page 1-s due to preface
            hypertexnames=false % solves warning, but interferes with
                                % index and \autoref apparently
            ]{hyperref}

\usepackage{amsmath}            % Formeln
\usepackage{amsfonts}           % Fonts for Formulas
\usepackage{amssymb}
\usepackage[final]{graphicx}
\usepackage{booktabs}
\usepackage{enumerate}
\usepackage{lipsum}
\usepackage{txfonts} % for strict implication symbols
\usepackage{soul}
\usepackage{relsize} % provides command \relsize{+/-x} for relative
                     % font size changes
\usepackage[german,english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc} 
\usepackage{subfig}
\usepackage{xypic}
\usepackage{url}
\usepackage{tikz}
\usetikzlibrary{arrows,shapes,automata,backgrounds,petri,fit,decorations.pathmorphing}
\usepackage{refcount}
\usepackage{xspace} % for \xspace in definition of acronyms etc.
\usepackage{attrib} % for right-aligned references at the end of
                    % quotes; part of Frankenstein bundle, but 
\renewcommand\PreTrib {}   % overwrite these commands, from attrib.sty
\renewcommand\PostTrib {}  % to suppress additional brackets around
                           % attributions
\usepackage{eurosym}
\usepackage{pgfplots}

\usepackage{../helpers/gb4e_micha}
\usepackage{../helpers/proofing}

\usepackage{../helpers/mycommands}
% \usepackage{../helpers/myenvironments}

\usepackage[]{svninfo}
\usepackage{subfig}

% \usepackage{geometry}
% \linespread{1.2}

%%%% Additional Macros

\newcommand{\lit}{\acro{lit}}
\newcommand{\glb}{\acro{glb}}
\newcommand{\loc}{\acro{loc}}

\newcommand{\as}{\acro{as}}
\renewcommand{\es}{\acro{es}}
\renewcommand{\AE}{\as}
\newcommand{\GE}{\es}
\newcommand{\ea}{\acro{ea}}

\newcommand{\lc}{\acro{lc}}
\newcommand{\ec}{\acro{ec}}
\newcommand{\LC}{\lc}
\newcommand{\EC}{\ec}

\newcommand{\exh}{\ensuremath{\mathrm{Exh}}}
\newcommand{\alt}{\ensuremath{\mathrm{Alt}}}

\newcommand{\ivt}{\acro{ivt}}

%%%% Document

\title{Scalar Items in Embedded Position: {A}n Experimental Revisit}
\author{Fabian Schlotterbeck, Michael Franke and Petra Augurzky}
\date{}

\begin{document}
\maketitle

\begin{abstract}
  \begin{itemize}
  \item What is needed to complement the picture is data from a
    verification experiment that is as independent as possible of
    potentially conflating properties of the pictorial material and
    that is nonetheless sensitive enough to detect local readings even
    if they are dispreferred.
  \end{itemize}

\end{abstract}

\tableofcontents

\svnInfo $Id: paper.tex 102 2012-07-04 10:19:59Z michael.franke $

\section{Introduction}
\label{sec:introduction}

The existential quantifier \emph{some} is usually assumed to receive a
semantic interpretation similar to logical $\exists$, so that the
sentence \emph{Some boys cried} is literally true in a situation where
all boys cried. But it is also usually considered to be a
\mymark{scalar item} in that its use invites comparison with (at
least) the semantically stronger universal quantifier \emph{all}
\citep[c.f.][]{Horn1972:On-the-Semantic,Gazdar1979:Pragmatics:-Imp,AtlasLevinson1981}. This
comparison can lead to an upper-bounding meaning enrichment, e.g.,
when an utterance of (\ref{bsp:Plain-SI-Target}) is taken to invite
the inference in (\ref{bsp:Plain-SI-Implicature}).

\begin{exe}
  \ex \label{bsp:Plain-SI}
    \begin{xlist}
      \ex \label{bsp:Plain-SI-Target} Hans solved some of the
        problems.
      \ex \label{bsp:Plain-SI-Implicature} $\implicates$ Hans solved
        some but not all of the problems.
      \ex \label{bsp:Plain-SI-Alternative} Hans solved all of the problems.
    \end{xlist}
\end{exe}

\noindent The classical explanation of this inference, following the
pioneering work of \citet{Grice1975:Logic-and-Conve} \citep[see][for
recent overview]{Geurts2010:Quantity-Implic}, is that
(\ref{bsp:Plain-SI-Implicature}) is a pragmatic inference, a so-called
\emph{quantity implicature}, derived by an abductive inference as the
best explanation of why informed, knowledgable and cooperative
speakers would utter (\ref{bsp:Plain-SI-Target}) when they could also
utter the semantically stronger and relevant
(\ref{bsp:Plain-SI-Alternative}).

% If this upper-bounding inference would occur often and
% systematically enough, then it may well be that also embedded
% occurrences of \emph{some} get enriched, in some fashion or other, to
% contribute the enriched meaning \emph{some but not all} also under the
% scope of other logical operators.
% Indeed, even
% \citeauthor{Grice1975:Logic-and-Conve} envisaged this possibility when
% he wrote: ``It may not be impossible for what starts life, so to
% speak, as a conversational implicature to become conventionalized''
% \citep[p.58]{Grice1975:Logic-and-Conve}.

This paper deals with the interpretation of two types of sentences,
where the scalar item \emph{some} occurs in the scope of other logical
operators. In \as-sentences (mnemonic for \textit{all \dots some \dots})
as in (\ref{bsp:AE}) the scalar item \emph{some} is embedded under
universal quantifier \emph{all}. In \es-sentences (mnemonic for
\textit{exactly one \dots some \dots}) like (\ref{bsp:GE}) \emph{some}
takes scope under the non-monotonic quantifier \emph{exactly one}.
According to current pragmatic theory, there are at least three
relevant candidate readings for \as- and \es-sentences: (i) a
\emph{literal reading} like in (\ref{bsp:AE-Literal}) and
(\ref{bsp:GE-Literal}) where \emph{some} has only its literal meaning;
(ii) a \emph{global reading} like in (\ref{bsp:AE-Global}) and
(\ref{bsp:GE-Global}) where, according to Gricean intuition, we enrich
utterances of (\ref{bsp:AE}) and (\ref{bsp:GE}) with the negation of
alternative utterances of the corresponding sentences
(\ref{bsp:AE-Alternative}) and (\ref{bsp:GE-Alternative}) where
\emph{some} is replaced by \emph{all}; and also (iii) a \emph{local
  reading} like in (\ref{bsp:AE-Local}) and (\ref{bsp:GE-Local}) where
\emph{some} is interpreted as \emph{some but not all} in the scope of
the embedding quantifier.

\begin{exe}
  \ex \label{bsp:AE} \mymark{All} of the students read {\mymark{some}} of the
  papers. \hfill{(\as)}

  \begin{xlist}
  \ex \label{bsp:AE-Literal} \mymark{All} of the students read
    {\mymark{some and maybe all}} of the papers. \hfill (\as-\lit)
  \ex \label{bsp:AE-Global}
    \mymark{All} of the students read \mymark{some and maybe all} 
    and  \hfill (\as-\glb)\\
    it's not the case that \mymark{all} of the students read \mymark{all} of the papers.
  \ex \label{bsp:AE-Local}
    \mymark{All} of the students read {\mymark{some  but not all}} of the
    papers. \hfill (\as-\loc)
  \end{xlist}
\end{exe}

\begin{exe}
\ex  \label{bsp:GE} \mymark{Exactly one} of the students read {\mymark{some}} of the
  papers. \hfill{(\es)}

  \begin{xlist}
  \ex \label{bsp:GE-Literal} \mymark{Exactly one} of the students read
    {\mymark{some and maybe all}} of the papers. \hfill (\es-\lit)
  \ex \label{bsp:GE-Global}
    \mymark{Exactly one} of the students read \mymark{some and maybe all} 
    and  \hfill (\es-\glb)\\
    it's not the case that \mymark{exactly one} of the students read \mymark{all} of the papers.
  \ex \label{bsp:GE-Local}
    \mymark{Exactly one} of the students read {\mymark{some  but not all}} of the
    papers. \hfill (\es-\loc)
  \end{xlist}
\end{exe}

\begin{exe}
\ex \label{bsp:AE-Alternative} \mymark{All} of the students read
  {\mymark{all}} of the papers.

\ex \label{bsp:GE-Alternative} \mymark{Exactly one} of the students
  read {\mymark{all}} of the papers.
\end{exe}

% Section~\ref{sec:get-know-your} will discuss these readings
% in more detail, showing that all three readings of each sentence are
% distinct but logically dependent on each other in interesting ways.

\noindent Given this relative abundance of theoretically conceivable
readings, two interlocked empirical questions arise:
\begin{enumerate}[Q1:]
\item which of these readings are available to na\"{i}ve subjects; and
\item which of the available readings are preferred.\fn{It may
    ultimately be impossible to keep unavailability and strong
    dispreference strictly apart. For our purposes this is not
    important. Every time we speak of availability, the so-inclined
    reader may think of non-negligible preference.}
\end{enumerate}
Addressing these questions empirically is relevant because they lie at
the heart of the current debate about the exact location and nature of
the interface between semantics and pragmatics. The controversy
concerns the question under which circumstances and to what extent the
upper-bounding inference from \emph{some} to \emph{some but not all}
can occur within the compositional computation of semantic values. Two
opposing theoretical positions can be distinguished here:
\emph{traditionalism} holds in particular that local readings should
not be readily attested unless marked by special intonation;
\emph{grammaticalism} holds that local readings are available and
indeed preferred. To distinguish between these opposing positions, it
is particularly relevant to assess the availability and relative
preference of local readings for \as- and
\es-sentences. %(Section~\ref{sec:theories-predictions}
%will discuss the relevant theoretical positions and the significance
%of alleged local readings in more detail.)

% Firstly, there are \mymark{pragmatic traditionalists} who seek to
% conserve the spirit of \citeauthor{Grice1975:Logic-and-Conve}'s
% (\citeyear{Grice1975:Logic-and-Conve}) original ideas as much as
% possible
% \citep[e.g.][]{Spector2006:Scalar-Implicat,Sauerland2004:Scalar-Implicat,Russell2006:Against-Grammat,vanRooijSchulz:ExhaustiveInterpretation,Geurts2010:Quantity-Implic,Franke2011:Quantity-Implic}. Traditionalists
% acknowledge the existence of global readings, but might consider local
% readings either unavailable or a beast distinct from scalar
% implicatures. An often phrased intuition of the traditionalists is
% that local readings, since they are special kinds of inferences,
% require special intonation, in particular emphatic stress on the
% scalar item.\dn{give references}

% Opposed to that is the camp of \mymark{lexical conventionalists}
% \citep[e.g.][]{LevinsonPresumptiveMeanings2000,Chierchia:2004_ScalarImplicatures}
% who maintain that scalar \emph{some} is lexically ambiguous between
% the standard logical meaning \emph{some and maybe all} and the
% upper-bounded meaning \emph{some but not all}. As the latter is
% considered a default, lexical conventionalism has no problem
% accounting for local readings, and in fact would consider these the
% preferred readings.

% Thirdly and finally, there is the camp of \mymark{grammaticalists} who
% defend that the distribution of upper-bounded readings of \emph{some}
% is best explained by postulating a silent operator, akin to the
% meaning of the particle \emph{only}
% \citep{Chierchia2006:Broaden-Your-Vi,Fox2007:Free-Choice-and,Magri2011:Another-Argumen,Sauerland2012:The-Computation,ChierchiaFox2008:The-Grammatical,Chierchia2012:FC-Nominals-and}. According
% to the grammatical view, this silent operator may be applied in
% compositional semantics also in the scope of other logical operators,
% but, so as not to overgenerate readings, the availability of readings
% is constraint by the \emph{strongest meaning hypothesis}
% \citep{DalrympleKanazawa1998:Reciprocal-Expr}. Grammaticalist theories
% predict that all three types of readings are available. Moreover,
% according to the grammatical view, the local reading is preferred for
% \as-sentences, while the global one is preferred for
% \es-sentences. (More precisely, as we will see
% Section~\ref{sec:theories-predictions}, depending on the variant of
% the strongest meaning hypothesis employed, either the global reading
% is predicted to be preferred for \es-sentences, or the global and the
% local reading are predicted to be equally preferred.)

A number of empirical studies have already targeted this issue
\citep[e.g.][]{GeurtsPouscoulous2009:Embedded-Implic,CliftonDube2010:Embedded-Implic,ChemlaSpector2010:Experimental-Ev}. Unfortunately,
results from these studies do not provide as clear-cut evidence for or
against either of the two rivaling positions as one would hope
for. This is due to a number of reasons, the most important of which
are the following. Firstly, previous studies have only accumulated
limited evidence pertaining to the preference relations between
putative readings of both types of sentences. Secondly, as
\citet{Tielvan-Tiel2012:Embedded-Scalar} and
\citet{GeurtsTielvan-Tiel2013:Scalar-expressi} argue, the putative
evidence for the availability of local readings of \as-sentences
suggested by studies of \citet{CliftonDube2010:Embedded-Implic} and
\citet{ChemlaSpector2010:Experimental-Ev}, might well be confounds
with what van Tiel and Geurts call \emph{typicality effects}. Finally,
and most importantly, previous studies presented target sentences
visually. This is highly problematic, because traditionalists admit to
the availability of local readings as long as it is intonationally
marked
\citep[e.g.][]{Horn2006:The-Border-Wars,Geurts2009:Scalar-Implicat,ChemlaSpector2010:Experimental-Ev,Geurts2010:Quantity-Implic,Tielvan-Tiel2012:Embedded-Scalar,GeurtsTielvan-Tiel2013:Scalar-expressi}. We
will address this extant position as the \emph{intonational markedness
  hypothesis}. But if intonation plays a role, it is absolutely vital
to present target sentences auditorily, so as to be able to control
for effects of ``silent prosody'' that subjects may apply when reading
a sentence \citep[see][]{Bader98,Fodor98}.

This paper therefore presents results form an experimental study
designed to deal with the above mentioned problems. We presented
sentence material auditorily and explicitly manipulated intonational
stress on embedded scalar items so as to explicitly so as to test the
intonational markedness hypothesis. To cater for potential pictorial
confounds, such as from typicality
\citep{Tielvan-Tiel2012:Embedded-Scalar,GeurtsTielvan-Tiel2013:Scalar-expressi},
we employed an \emph{incremental verification task} in which parts of
a to-be-evaluated picture were hidden and only revealed piece-wise at
the participants' request until they felt able to give a binary
truth-value judgement (SEE CONROY ????). This way we obtained
categorical behavioral data that is both indicative of the reading
subjects assumed and independent of any potential pictorial
confounds. We also made sure that our method is able to reveal
information about reading preferences by including a control condition
with ambiguous sentences with a well-known preference on available
readings.

The data obtained by this method contradicts both traditionalism and
grammaticalism. Contrary to traditionalism and the intonational
markedness hypothesis, we found evidence for local readings for both
kinds of critical sentences, irrespective of intonational stress. Even
the intonational markedness hypothesis on its own is contradicted by
our data: stress on embedded scalar items did not lead to an increase
in answers suggesting a local reading. But also grammaticalism is
incongruent with the observed pattern in our data.  Grammaticalism
predicts local readings to be the preferred readings, but we found no
evidence for that. More strongly even, the preference order among
putative readings suggested by our data cannot even in principle be
explained by the general kind of second-order selection criterion that
is currently favored by grammaticalists, namely by comparison of
logical strength of putative readings
\citep[e.g.][]{ChierchiaFox2008:The-Grammatical}. In sum, our data
sheds doubt on both traditionalism and grammaticalism in their
currently most common forms. We conclude the paper by some speculation
about how either position might be amended in the light of our data.

The paper is structured as follows. Section~\ref{sec:get-know-your}
elaborates on the three kinds of relevant readings for our target
sentences. Section~\ref{sec:theories-predictions} works out the
predictions of traditionalism and grammaticalism about availability
and preference of readings of our target sentences. The intonational
markedness hypothesis, which is frequently relied on to back-up
traditionalism, is elaborated on in
Section~\ref{sec:inton-mark-hypoth}, together with a concise
assessment of the sparse empirical data on the effects of intonation
on the interpretation of (embedded) scalar
items. Section~\ref{sec:previous-studies} recaps the results of
previous related studies, pointing out the most important necessary
amendments that motivated out own design. Section~\ref{sec:exp}
eventually 
describes our experimental design and continues to state, analyze and interpret the
results we obtained. We defend our design and the interpretation of
the results based on it in Section~\ref{sec:conclusions}.

\section{Get to know your readings}
\label{sec:get-know-your}

Three readings are \emph{prima facie} conceivable for the \as- and
\es-sentences in (\ref{bsp:AE}) and (\ref{bsp:GE}). These are
logically dependent in intricate ways.

\paragraph{\as-sentences.}

An \as-sentence like (\ref{bsp:AE}), repeated below, has a literal
reading (\lit) as in (\ref{bsp:AE-Literal}), a global reading (\glb) as in
(\ref{bsp:AE-Global}) and a local reading (\loc) as in (\ref{bsp:AE-Local}).

\begin{exer}{bsp:AE}

  \ex \mymark{All} of the students read {\mymark{some}} of the
  papers.

  \begin{xlist}
  \ex \mymark{All} of the students read
    {\mymark{some and maybe all}} of the papers. \hfill (\as-\lit)
  \ex
    \mymark{All} of the students read \mymark{some and maybe all} 
    and  \hfill (\as-\glb)\\
    it's not the case that \mymark{all} of the students read \mymark{all} of the papers.
  \ex
    \mymark{All} of the students read {\mymark{some  but not all}} of the
    papers. \hfill (\as-\loc)
  \end{xlist}
\end{exer}

\noindent These readings stand in a strict entailment relation: the
local reading asymmetrically entails the global reading, which
asymmetrically entails the literal reading:
\begin{exe}
  \ex \label{bsp:Entailments-AS} \loc $\subset$ \glb $\subset$ \lit
\end{exe}
\glb entails \lit because, in general, global readings are defined as
the conjunction of the literal reading and the negated
(relevant/feasible) alternative(s) of the to-be-interpreted
utterance. This entailment is asymmetric, because the information that
not all of the students read all of the papers is not entailed by the
literal reading. To see that \loc $\subset$ \glb, notice that the case
where all of the students read some but not all of the papers is a
special case of the set of situations where all of the students read
some (and maybe all), while not all of the students read all of the
papers.

Given these entailment relations, there are four kinds of situations,
the names for which we borrow from
\citet{ChemlaSpector2010:Experimental-Ev}, that we can distinguish
based on different truth values for our candidate readings. These are
given in Table~\ref{fig:situations-AE}.
%
\begin{table}
  \centering
  \subfloat[][\as-sentences]{
  \label{fig:situations-AE}
    \begin{tabular}{lccc}
    \toprule
    situation    & \multicolumn{3}{c}{truth value} 
  \\ 
  \cmidrule(r){2-4}
     & \lit & \glb & \loc \\ \midrule
    false   & 0 & 0 & 0 \\
    literal & 1 & 0 & 0 \\
    weak    & 1 & 1 & 0 \\
    strong  & 1 & 1 & 1 \\ \bottomrule
  \end{tabular}
} \qquad \qquad
  \subfloat[][\es-sentences]{
  \label{fig:situations-GS}
  \begin{tabular}{lccc}
    \toprule
    situation    & \multicolumn{3}{c}{truth value} 
  \\ 
  \cmidrule(r){2-4}
    & \textsc{lit} & \textsc{glb} & \textsc{loc} \\ \midrule
    false   & 0 & 0 & 0 \\
    literal & 1 & 0 & 0 \\
    local   & 0 & 0 & 1 \\
    all     & 1 & 1 & 1 \\ \bottomrule
  \end{tabular}
}
\caption{Possible truth-value distributions for readings of \as- and \es-sentences}
\label{tab:situations}
\end{table}
%
Examples of these kinds of situations are given in
Figure~\ref{fig:AS-distinguishing-pics}, where the dots on the right
of each diagram represent students, the dots on the left represent
papers and an arrow from a student to a paper indicates that the
student read the paper. Notice that other arrangements of arrows might
equally well serve as examples for the various situations.

\begin{figure}[]
  \centering

\subfloat[fig:false][false]{
  \label{fig:false-AE}

\begin{tikzpicture}[node distance = 1cm]
        % Nodes

        \node (X-1) {$\Large{\bullet}$};

        \node (X-2) [below of = X-1] {$\Large{\bullet}$};

        \node (X-3) [below of = X-2] {$\Large{\bullet}$};

        \node (Y-1) [right of = X-1, node distance = 2cm]{$\Large{\bullet}$};

        \node (Y-2) [below of = Y-1] {$\Large{\bullet}$};

        \node (Y-3) [below of = Y-2] {$\Large{\bullet}$};

        \node (Y-4) [below of = Y-3] {$\Large{\bullet}$};

        % table added from here

        \node (X-4) [below of = X-3, node distance = 2cm] {};

        \node (table) [right of = X-4] {
          \begin{tabular}{ll}
            \lit & false \\
            \glb & false \\
            \loc & false 
          \end{tabular}};

        % Arrows

        \path [draw=blue,->] (X-1) -> (Y-1);

        \path [draw=blue,->] (X-1) -> (Y-2);

        \path [draw=red,->] (X-3) -> (Y-3);

        \path [draw=red,->] (X-3) -> (Y-4);

      \end{tikzpicture}

}
\subfloat[fig:literal][literal]{
  \label{fig:literal-AE}

  \begin{tikzpicture}[node distance = 1cm]
        % Nodes

        \node (X-1) {$\Large{\bullet}$};

        \node (X-2) [below of = X-1] {$\Large{\bullet}$};

        \node (X-3) [below of = X-2] {$\Large{\bullet}$};

        \node (Y-1) [right of = X-1, node distance = 2cm]{$\Large{\bullet}$};

        \node (Y-2) [below of = Y-1] {$\Large{\bullet}$};

        \node (Y-3) [below of = Y-2] {$\Large{\bullet}$};

        \node (Y-4) [below of = Y-3] {$\Large{\bullet}$};

        % table added from here

        \node (X-4) [below of = X-3, node distance = 2cm] {};

        \node (table) [right of = X-4] {
          \begin{tabular}{ll}
            \lit & true \\
            \glb & false \\
            \loc & false 
          \end{tabular}};

        % Arrows

        \path [draw=blue,->] (X-1) -> (Y-1);

        \path [draw=blue,->] (X-1) -> (Y-2);

        \path [draw=blue,->] (X-1) -> (Y-3);

        \path [draw=blue,->] (X-1) -> (Y-4);

        \path [draw=black,->] (X-2) -> (Y-1);

        \path [draw=black,->] (X-2) -> (Y-2);

        \path [draw=black,->] (X-2) -> (Y-3);

        \path [draw=black,->] (X-2) -> (Y-4);

        \path [draw=red,->] (X-3) -> (Y-1);

        \path [draw=red,->] (X-3) -> (Y-2);

        \path [draw=red,->] (X-3) -> (Y-3);

        \path [draw=red,->] (X-3) -> (Y-4);

      \end{tikzpicture}

}
\subfloat[fig:weak][weak]{
  \label{fig:weak}

\begin{tikzpicture}[node distance = 1cm]
        % Nodes

        \node (X-1) {$\Large{\bullet}$};

        \node (X-2) [below of = X-1] {$\Large{\bullet}$};

        \node (X-3) [below of = X-2] {$\Large{\bullet}$};

        \node (Y-1) [right of = X-1, node distance = 2cm]{$\Large{\bullet}$};

        \node (Y-2) [below of = Y-1] {$\Large{\bullet}$};

        \node (Y-3) [below of = Y-2] {$\Large{\bullet}$};

        \node (Y-4) [below of = Y-3] {$\Large{\bullet}$};

        % table added from here

        \node (X-4) [below of = X-3, node distance = 2cm] {};

        \node (table) [right of = X-4] {
          \begin{tabular}{ll}
            \lit & true \\
            \glb & true \\
            \loc & false 
          \end{tabular}};

        % Arrows

        \path [draw=blue,->] (X-1) -> (Y-1);

        \path [draw=blue,->] (X-1) -> (Y-2);

        \path [draw=black,->] (X-2) -> (Y-1);

        \path [draw=black,->] (X-2) -> (Y-2);

        \path [draw=black,->] (X-2) -> (Y-3);

        \path [draw=black,->] (X-2) -> (Y-4);

        \path [draw=red,->] (X-3) -> (Y-3);

        \path [draw=red,->] (X-3) -> (Y-4);

      \end{tikzpicture}

}
\subfloat[fig:strong][strong]{
  \label{fig:strong}

\begin{tikzpicture}[node distance = 1cm]
        % Nodes

        \node (X-1) {$\Large{\bullet}$};

        \node (X-2) [below of = X-1] {$\Large{\bullet}$};

        \node (X-3) [below of = X-2] {$\Large{\bullet}$};

        \node (Y-1) [right of = X-1, node distance = 2cm]{$\Large{\bullet}$};

        \node (Y-2) [below of = Y-1] {$\Large{\bullet}$};

        \node (Y-3) [below of = Y-2] {$\Large{\bullet}$};

        \node (Y-4) [below of = Y-3] {$\Large{\bullet}$};

        % table added from here

        \node (X-4) [below of = X-3, node distance = 2cm] {};

        \node (table) [right of = X-4] {
          \begin{tabular}{ll}
            \lit & true \\
            \glb & true \\
            \loc & true 
          \end{tabular}};

        % Arrows

        \path [draw=blue,->] (X-1) -> (Y-1);

        \path [draw=blue,->] (X-1) -> (Y-2);

        \path [draw=black,->] (X-2) -> (Y-2);

        \path [draw=black,->] (X-2) -> (Y-3);

        \path [draw=red,->] (X-3) -> (Y-3);

        \path [draw=red,->] (X-3) -> (Y-4);

      \end{tikzpicture}

}

  \caption{Distinguishing scenarios for \as-sentences}
  \label{fig:AS-distinguishing-pics}
\end{figure}

\paragraph{\es-sentences.}

The situation for \es-sentences like (\ref{bsp:GE}) is similar but a
little more complicated because the embedding quantifier is
non-monotonic. Again, we consider a literal reading as in
(\ref{bsp:GE-Literal}), repeated below, a global reading as in
(\ref{bsp:GE-Global}) and a local reading as in (\ref{bsp:GE-Local}).

\begin{exer}{bsp:GE}
\ex \mymark{Exactly one} of the students read {\mymark{some}} of the
  papers.

  \begin{xlist}
  \ex \mymark{Exactly one} of the students read
    {\mymark{some and maybe all}} of the papers. \hfill (\es-\lit)
  \ex 
    \mymark{Exactly one} of the students read \mymark{some and maybe all} 
    and  \hfill (\es-\glb)\\
    it's not the case that \mymark{exactly one} of the students read \mymark{all} of the papers.
  \ex 
    \mymark{Exactly one} of the students read {\mymark{some  but not all}} of the
    papers. \hfill (\es-\loc)
  \end{xlist}
\end{exer}

\noindent Entailment relations in this case are non-linear:
\begin{exe}
  \ex \label{bsp:Entailments-GE}       \lit\ $\supset$ \glb\ $\color{mycol}{\subset}$ \loc 
    % \hspace{1cm} and \hspace{1cm}  \lit\ $\not \supset$ \loc
\end{exe}
By definition of global readings $\glb$ entails $\lit$. This
entailment is asymmetric because the extra information that it is not
the case that exactly one student read all of the papers is not
entailed by the literal reading (\ref{bsp:GE-Literal}). However,
unlike for \as-sentences, \loc is not stronger than \glb, but
asymmetrically entailed by the latter. To see this, notice that the
global reading is equivalent to:

\begin{exe}
  \exp{bsp:GE-Global} \mymark{Exactly one} of the students read \mymark{some but not
    all} and \\
  \mymark{everybody else} read \mymark{none} of the papers.
\end{exe}

\noindent Finally, \loc and \lit are logically independent: all
possible combinations of truth-values for \loc and \lit are possible.

Given these entailment relations, there are again four different
situations corresponding to the four possible distributions of truth
values for candidate readings. These are given in
Table~\ref{fig:situations-GS} and named following
\citet{ChemlaSpector2010:Experimental-Ev}. Examples for each situation
are given in Figure~\ref{fig:ES-distinguishing-pics}.

\begin{figure}[]
  \centering

\subfloat[fig:false][false]{
  \label{fig:false-GE}

  \begin{tikzpicture}[node distance = 1cm]
        % Nodes

        \node (X-1) {$\Large{\bullet}$};

        \node (X-2) [below of = X-1] {$\Large{\bullet}$};

        \node (X-3) [below of = X-2] {$\Large{\bullet}$};

        \node (Y-1) [right of = X-1, node distance = 2cm]{$\Large{\bullet}$};

        \node (Y-2) [below of = Y-1] {$\Large{\bullet}$};

        \node (Y-3) [below of = Y-2] {$\Large{\bullet}$};

        \node (Y-4) [below of = Y-3] {$\Large{\bullet}$};

        % table added from here

        \node (X-4) [below of = X-3, node distance = 2cm] {};

        \node (table) [right of = X-4] {
          \begin{tabular}{ll}
            \lit & false \\
            \glb & false \\
            \loc & false 
          \end{tabular}};

        % Arrows

        \path [draw=blue,->] (X-1) -> (Y-1);

        \path [draw=blue,->] (X-1) -> (Y-2);

        \path [draw=red,->] (X-3) -> (Y-3);

        \path [draw=red,->] (X-3) -> (Y-4);

      \end{tikzpicture}

}
\subfloat[fig:literal][literal]{
  \label{fig:literal-GE}

      \begin{tikzpicture}[node distance = 1cm]
        % Nodes

        \node (X-1) {$\Large{\bullet}$};

        \node (X-2) [below of = X-1] {$\Large{\bullet}$};

        \node (X-3) [below of = X-2] {$\Large{\bullet}$};

        \node (Y-1) [right of = X-1, node distance = 2cm]{$\Large{\bullet}$};

        \node (Y-2) [below of = Y-1] {$\Large{\bullet}$};

        \node (Y-3) [below of = Y-2] {$\Large{\bullet}$};

        \node (Y-4) [below of = Y-3] {$\Large{\bullet}$};

        % table added from here

        \node (X-4) [below of = X-3, node distance = 2cm] {};

        \node (table) [right of = X-4] {
          \begin{tabular}{ll}
            \lit & true \\
            \glb & false \\
            \loc & false 
          \end{tabular}};

        % Arrows

        \path [draw=black,->] (X-2) -> (Y-1);

        \path [draw=black,->] (X-2) -> (Y-2);

        \path [draw=black,->] (X-2) -> (Y-3);

        \path [draw=black,->] (X-2) -> (Y-4);

      \end{tikzpicture}

}
\subfloat[fig:local][local]{
  \label{fig:local}

            \begin{tikzpicture}[node distance = 1cm]
        % Nodes

        \node (X-1) {$\Large{\bullet}$};

        \node (X-2) [below of = X-1] {$\Large{\bullet}$};

        \node (X-3) [below of = X-2] {$\Large{\bullet}$};

        \node (Y-1) [right of = X-1, node distance = 2cm]{$\Large{\bullet}$};

        \node (Y-2) [below of = Y-1] {$\Large{\bullet}$};

        \node (Y-3) [below of = Y-2] {$\Large{\bullet}$};

        \node (Y-4) [below of = Y-3] {$\Large{\bullet}$};

        % table added from here

        \node (X-4) [below of = X-3, node distance = 2cm] {};

        \node (table) [right of = X-4] {
          \begin{tabular}{ll}
            \lit & false \\
            \glb & false \\
            \loc & true 
          \end{tabular}};

        % Arrows

        \path [draw=blue,->] (X-1) -> (Y-1);

        \path [draw=blue,->] (X-1) -> (Y-2);

        \path [draw=blue,->] (X-1) -> (Y-3);

        \path [draw=blue,->] (X-1) -> (Y-4);

        \path [draw=black,->] (X-2) -> (Y-2);

        \path [draw=black,->] (X-2) -> (Y-3);

        \path [draw=black,->] (X-2) -> (Y-4);

        \path [draw=red,->] (X-3) -> (Y-1);

        \path [draw=red,->] (X-3) -> (Y-2);

        \path [draw=red,->] (X-3) -> (Y-3);

        \path [draw=red,->] (X-3) -> (Y-4);

      \end{tikzpicture}

}
\subfloat[fig:all][all]{
  \label{fig:all}

      \begin{tikzpicture}[node distance = 1cm]
        % Nodes

        \node (X-1) {$\Large{\bullet}$};

        \node (X-2) [below of = X-1] {$\Large{\bullet}$};

        \node (X-3) [below of = X-2] {$\Large{\bullet}$};

        \node (Y-1) [right of = X-1, node distance = 2cm]{$\Large{\bullet}$};

        \node (Y-2) [below of = Y-1] {$\Large{\bullet}$};

        \node (Y-3) [below of = Y-2] {$\Large{\bullet}$};

        \node (Y-4) [below of = Y-3] {$\Large{\bullet}$};

        % table added from here

        \node (X-4) [below of = X-3, node distance = 2cm] {};

        \node (table) [right of = X-4] {
          \begin{tabular}{ll}
            \lit & true \\
            \glb & true \\
            \loc & true 
          \end{tabular}};

        % Arrows

        \path [draw=blue,->] (X-1) -> (Y-1);

        \path [draw=blue,->] (X-1) -> (Y-2);

      \end{tikzpicture}

}

  \caption{Distinguishing scenarios for \es-sentences}
  \label{fig:ES-distinguishing-pics}

\end{figure}

\section{Theories and predictions}
\label{sec:theories-predictions}

There are two main theoretical positions which make different
predictions about the readings of \as- and \es-sentences
\citep[c.f.][for
overview]{Horn2006:The-Border-Wars,Geurts2010:Quantity-Implic,Sauerland2012:The-Computation}. We
will refer to these here as \mymark{traditionalism} and
\mymark{grammaticalism} and treat each one in turn. Since there is
some leeway in assessing the predictions for each (depending on which
of several reasonable additional assumptions we should adopt), we will
distinguish different varieties of both. For convenience, the
predictions of each (variety of each) position are also summarized in
Table~\ref{tab:predictions} at the end of this section.

\subsection{Traditionalism}
\label{sec:traditionalism}

We refer to traditionalism as traditionalism because of its
conservative stance towards Grice's original theory of conversational
implicatures \citep{Grice1975:Logic-and-Conve}. Many author's have
defended traditionalist positions in this sense. Of the more recent
literature, we would consider as traditionalist, among others,
contributions by \citet{Spector2006:Scalar-Implicat},
\citet{Sauerland2004:Scalar-Implicat},
\citet{Russell2006:Against-Grammat},
\citet{vanRooijSchulz:ExhaustiveInterpretation},
\citet{Geurts2010:Quantity-Implic} or
\citet{Franke2011:Quantity-Implic}.

According to Grice, conversational implicatures, of which quantity
implicatures are a special case, are to be thought of as
rationalizations of speaker behavior. Central in this reasoning is the
assumption that the speaker's behavior is efficient (if not optimal)
and goal-oriented. Usually, the assumed goal of conversation is the
cooperative exchange of relevant information from speaker to hearer.

Consequently, the \mymark{Gricean recipe}
\citep[c.f.][]{Geurts2010:Quantity-Implic} for deriving a simple
scalar inference like that in (\ref{bsp:Plain-SI-Implicature}) from an
utterance of (\ref{bsp:Plain-SI-Target}) is as follows: if the issue
whether Hans solved only some or all of the problems is relevant, then
a cooperative and knowledgable speaker would utter
(\ref{bsp:Plain-SI-Alternative}) if in a position to do so; hence, one
of the most natural explanations of why such a speaker has not uttered
(\ref{bsp:Plain-SI-Alternative}), but only (\ref{bsp:Plain-SI-Target})
is that she is uncertain of whether (\ref{bsp:Plain-SI-Alternative})
is true; but on the assumption that she is knowledgeable (competent,
opinionated, informed \dots) it follows that
(\ref{bsp:Plain-SI-Implicature}) should in fact be true.\fn{We are
  glossing here somewhat swiftly over the more nuanced details of the
  derivation of implicatures targeting the speaker's epistemic state
  \citep[e.g.][]{Gazdar1979:Pragmatics:-Imp,Soames1982:How-Presupposit},
  as this is not crucially relevant for the issues we are interested
  in here.}\textsuperscript{,}\dn{do we need to enlarge on epistemic
  implicatures?}

\begin{exer}{bsp:Plain-SI}
  \ex 
    \begin{xlist}
      \ex \label{bsp:Plain-SI-Target} Hans solved \mymark{some} of the problems.
      \ex \label{bsp:Plain-SI-Implicature} $\implicates$ Hans solved
        \mymark{some but not all} of the problems.
      \ex  \label{bsp:Plain-SI-Alternative}  Hans solved \mymark{all} of the problems.
    \end{xlist}
\end{exer}

The Gricean recipe applies also to \as- and \es-sentences and derives
the global reading in a straightforward way. Consequently,
traditionalism predicts that both literal and global readings are
available: literal readings, because these form the starting point of
pragmatic reasoning; global readings because these may be arrived at
by the Gricean recipe.

Which of these readings, if any, does traditionalism predict to be
preferred? This depends on whether the auxiliary assumptions necessary
to derive global readings by the Gricean recipe are plausibly met in
the particular case of utterance of \as- and \es-sentences. These
auxiliary assumptions include relevance of the extra information
provided in the global reading, mutual awareness that the stronger
alternative has been a speaker option, the speaker's competence about
the issue, etc. Normally, traditionalist accounts would assume that
these extra assumptions are met. In that case, traditionalism would
predict that global readings should be preferred over literal
readings. On the other hand, it might also be hypothesized that, for
example, a competence assumption is harder to justify for \as- and
\es-sentences in general than for simpler sentences such as
(\ref{bsp:Plain-SI}): it might be less clear that the speaker knows
exactly how many students solved how many problems, than that the
speaker knows exactly how many problems, e.g., Hans solved. In that
case, or if any other assumption of the Gricean recipe cannot be
maintained, traditionalism would predict that the literal reading
would be preferred over the global one. But that means that there are
at least two varieties of traditionalism that, depending on which
additional assumptions we would make, yield slightly different
predictions: \mymark{the strong variety of traditionalism} maintains
that the auxiliary assumptions of the Gricean recipe hold
usually/strongly/unless-completely-intenable and therefore predicts
that global readings are preferred over literal ones; \mymark{the weak
  variety of traditionalism} holds that the auxiliary assumptions are
more fragile and predicts that literal readings are preferred over the
global ones.\fn{We mention for clarity that although weak
  traditionalism does not predict the global reading to be preferred,
  it might still predict an \emph{epistemically weak implicature},
  similar to the global reading, that the speaker is uncertain whether
  the stronger alternative is true. Whether it does predict that
  depends on which auxiliary assumptions are assumed to be met and
  which are not.}

Does traditionalism also predict local readings to be available? The
answer is slightly different for \as- and the \es-sentences.\fn{Some
  traditionalists concede that (something like) local readings can
  occur for both kinds of sentences if scalar items are marked with
  special accentuation, albeit then as a signal of a different
  pragmatic process
  \citep[e.g.][]{Horn2006:The-Border-Wars,Geurts2009:Scalar-Implicat,Geurts2010:Quantity-Implic}. We
  will address this addition to traditionalism in more detail in
  Section~\ref{sec:inton-mark-hypoth}.} For \es-sentences,
traditionalism does not predict that a local reading is available, at
least not as a quantity implicature
\citep[c.f.][]{GeurtsPouscoulous2009:Embedded-Implic,ChemlaSpector2010:Experimental-Ev}. This
is because traditionalism assumes that quantity implicatures are
pragmatic enrichments of the literal meaning of an utterance, obtained
by conjoining the literal meaning with a suitable set of negated
alternatives. But since the literal and the local reading of
\es-sentences are logically independent, there is no way that local
readings can be derived in a traditionalist manner as a quantity
implicature: if $X$ and $Y$ are logically independent propositions,
then there is no proposition $Z$ such that $X$ would be equivalent to
$Y \wedge \neg Z$.

On the other hand, as for \as-sentences, there is a traditionalist way
of deriving the local reading, namely by assuming that not only
(\ref{bsp:AE-Alternative}) is an alternative to (\ref{bsp:AE}), but
also the sentence in (\ref{bsp:AE-Alternatives-Extended}).

\begin{exer}{bsp:AE}
  \ex \mymark{All} of the students solved \mymark{some} of the problems.
\end{exer}

\begin{exer}{bsp:AE-Alternative}
  \ex \mymark{All} of the students solved \mymark{all} of the problems.
\end{exer}

\begin{exe}
\ex \label{bsp:AE-Alternatives-Extended} \mymark{Some} of the students solved \mymark{all} of the problems.
\end{exe}

\noindent Clearly, if we conjoin a literal reading of (\ref{bsp:AE})
with the negation of (\ref{bsp:AE-Alternatives-Extended}) we obtain
exactly the local reading. (Notice that the negation of
(\ref{bsp:AE-Alternatives-Extended}) entails the negation of
(\ref{bsp:AE-Alternative}), so it makes no difference (not) to add it
in this case.) Consequently, traditionalism does predict that the
local reading of \as-sentences is available if
(\ref{bsp:AE-Alternatives-Extended}) is an available alternative. We
should therefore introduce another distinction: \mymark{restricted
  traditionalism} assumes that (\ref{bsp:AE-Alternatives-Extended}) is
not available and therefore predicts that the local reading for
\as-sentences is not available; in contrast, \mymark{unrestricted
  traditionalism} assumes that (\ref{bsp:AE-Alternatives-Extended}) is
available and so is the local reading.

Still, even unrestricted traditionalism would not predict that the
local reading is preferred over the global reading. This is because
the derivation of the local reading via
(\ref{bsp:AE-Alternatives-Extended}) hinges on yet another auxiliary
assumption, namely the availability of the alternative in
(\ref{bsp:AE-Alternatives-Extended}), which seems much less obvious
and therefore presumably is less readily available than the
alternative in (\ref{bsp:AE-Alternative}). Consequently, we derive the
following predictions about preferences for unrestricted
traditionalism: weak unrestricted traditionalism prefers the literal
over the global over the local reading, while strong unrestricted
traditionalism prefers the global over the local over the literal
reading (see also Table~\ref{tab:predictions}).

\subsection{Grammaticalism}
\label{sec:grammaticalism}

The other currently prominent theoretical position is
\mymark{grammaticalism}. Recently, a lot of evidence based on
intuitive judgements and more theoretical considerations has been
discussed in favor of this position
\citep[c.f.][]{Chierchia2006:Broaden-Your-Vi,Fox2007:Free-Choice-and,Magri2011:Another-Argumen,Sauerland2012:The-Computation,ChierchiaFox2008:The-Grammatical,Chierchia2012:FC-Nominals-and}. % When
% it comes to the predictions of available readings of \as- and
% \es-sentences, grammaticalism appears like the conjunction of
% traditionalism and conventionalism. But conceptually speaking,
% grammaticalism goes down a road of its own.
Grammaticalism approaches quantity implicatures by postulating a
silent operator that can be variably applied during compositional
computation of a sentence's truth-value
\citep{Chierchia2006:Broaden-Your-Vi}, if necessary multiple times
\citep{Fox2007:Free-Choice-and}. This silent operator is variably
referred to as $\mathrm{O}(\cdot)$ or $\exh(\cdot)$, because it is
assumed to be similar in effect to the meaning of particle \emph{only}
or of the mechanism of \emph{exhaustive interpretation}
\citep{GroenendijkStokhofThesis1984,Stechowvon-StechowZimmermann1984:Term-Answers-an,Rooijvan-RooijSchulz2013:Exhaustive-Inte,vanRooijSchulz:ExhaustiveInterpretation,Fox2007:Free-Choice-and}. For
our purposes, it is enough to note that $\exh(\cdot)$ is a
poly-typed function that enriches an expression $X$, which crucially
need not be a full proposition, based on a set $\alt(X)$ of
(suitable, relevant) alternatives to $X$ that yields an enriched
meaning of the form:\fn{More sophisticated formulations of exhaustive
  interpretation have been proposed
  \citep[e.g.][]{Schulz2005:A-Pragmatic-Sol,vanRooijSchulz:ExhaustiveInterpretation,Spector2006:Scalar-Implicat,Fox2007:Free-Choice-and}
  but this simple formulation is sufficient for the purposes of this
  paper. Also, we gloss here over non-trivial details in the way this
  operation is to be specified exactly within a compositional
  semantics \citep[c.f.][]{Chierchia2006:Broaden-Your-Vi}.}
\begin{exe}
  \ex \label{bsp:Exh-Def} $\exh(X,\alt(X)) = X \bigwedge_{A \in
      \alt(X)} \neg A$.
\end{exe}

As this operator can apply at various scope sites, the grammatical
approach predicts that all three readings of \as- and \es-sentences
are available: literal readings arise if no exhaustification operator
is applied; global readings arise if the exhaustification operator
takes sentence-wide scope as in (\ref{Grammar-Global}); and local
readings arise if the exhaustification operator takes scope under the
respective quantifiers as in (\ref{Grammar-Local}).

\begin{exe}
  \ex \label{Grammar-Global}
    \begin{xlist}
      \ex \label{Grammar-Global-AE} \mymark{$\exh$}(\mymark{All} of the students solved
        \mymark{some} of the problems).
      \ex \label{Grammar-Global-GE} \mymark{$\exh$}(\mymark{Exactly one} of the students solved
        \mymark{some} of the problems).
    \end{xlist}
\end{exe}

\begin{exe}
  \ex \label{Grammar-Local}
    \begin{xlist}
      \ex \label{Grammar-Local-AE} \mymark{All} of the students \mymark{$\exh$}(solved
        \mymark{some} of the problems).
      \ex \label{Grammar-Local-GE} \mymark{Exactly one} of the
        students \mymark{$\exh$}( solved
        \mymark{some} of the problems).
    \end{xlist}
\end{exe}

The grammatical approach, as described so far, is rather flexible in
that it makes many readings available. It is so flexible, in fact,
that it must be constrained in some way or other to shield the
approach against overgeneration. Grammaticalism is therefore aided by
a principle that specifies which readings are to be preferred if
several readings are generated by optional applications of
$\exh(\cdot)$ in the compositional computation of semantic values. The
principle that grammaticalists adhere to
\citep[c.f.][]{FoxSpector:Economy-and-Emb,ChierchiaFox2008:The-Grammatical,Chierchia2012:FC-Nominals-and}
is the \mymark{strongest meaning hypothesis} of
\citet{DalrympleKanazawa1998:Reciprocal-Expr}. Roughly speaking, if a
given sentence has several conceivable readings, the strongest meaning
hypothesis selects for the strongest of
these. \citet{ChierchiaFox2008:The-Grammatical} discuss two concrete
variants of the strongest meaning hypothesis: fix a sentence with
propositional content $S$ with candidate readings $C(S)$, where $C(S)$
contains $S$ and all readings derivable from inserting $\exh(\cdot)$
at suitable scope sites;\fn{Strictly speaking, we should compare
  \emph{parses} of sentences with $\exh(\cdot)$ at various scope sites
  with respect to the readings that these parses give rise to
  \citep[c.f.][]{ChierchiaFox2008:The-Grammatical}, but we may ignore
  this detail here for ease of exposition.} then define \emph{ceteris
  paribus} preferences among members of $X,Y \in C(S)$ as either
\begin{exe}
\ex \label{bsp:SMH-Total} 
  $X$ is preferred to $Y$ iff $X \subset Y$ 
  \attrib{\citep[(104)]{ChierchiaFox2008:The-Grammatical}}
\end{exe}
or
\begin{exe}
\ex \label{bsp:SMH-Partial}  $X$ is preferred to $Y$ iff $X \subset Y$
  and $X$ and $Y$ are readings obtained from applications of
  $\exh(\cdot)$ at exactly the same scope sites, except for
  one. \attrib{\citep[(105)]{ChierchiaFox2008:The-Grammatical}} 
\end{exe}
Since the former is entailed by the latter, we will speak of
\mymark{weak grammaticalism} and \mymark{strong grammaticalism},
depending on whether (\ref{bsp:SMH-Total}) or (\ref{bsp:SMH-Partial})
is used.

These two varieties of grammaticalism give rise to different
predictions about preferred readings. The preference for readings that
weak grammaticalism predicts mirrors the entailment relations in
(\ref{bsp:Entailments-AS}) and (\ref{bsp:Entailments-GE}): for
\as-sentences weak grammaticalism predicts that local readings are
preferred over global readings which in turn are preferred over
literal readings; for \es-sentences weak grammaticalism predicts that
global readings are most preferred, while literal and local readings
are not ranked with respect to preference. Strong grammaticalism, on
the other hand, predicts that \as-sentences preferably get either a
local or a global reading rather than a literal reading, but does not
rank these former two with respect to each other because they differ
in more than one application of the $\exh(\cdot)$ operator. For
\es-sentences, then, strong grammaticalism does not predict \emph{any}
preference relations between candidate readings at all.

\medskip

Taking stock, traditionalism, conventionalism and grammaticalism
predict different patterns of availability and preference for \as- and
\es-sentences. Most notably, traditionalism does not predict local
readings for \es-sentences to be available, while conventionalism does
not predict any global readings to be available. Grammaticalism
predicts that all readings are available, but, given the strongest
meaning hypothesis, also comes with a clear indication of preference:
literal meanings are dispreferred. All predictions are summarized in
Table~\ref{tab:predictions}.

\begin{table}[t]
  \centering
  \begin{tabular}{lcc}
    & \multicolumn{2}{c}{preference \& availability}
    \\ \cmidrule(r){2-3}
    theoretical position
    & \as
    & \es
    \\ \midrule
    traditionalism
    \\
    \ \ \ \ weak restricted 
    & \lit > \glb 
    & \lit > \glb
    \\
    \ \ \ \ weak unrestricted
    & \lit > \glb > \loc 
    & \lit > \glb
    \\
    \ \ \ \ strong restricted
    & \glb > \lit 
    & \glb > \lit
    \\
    \ \ \ \ strong unrestricted
    & \glb > \loc > \lit 
    & \glb > \lit
    \\
    % conventionalism
    % \\
    % \ \ defaultism
    % & \loc >  \lit 
    % & \loc >  \lit
    % \\
    % \ \ literalism
    % & \lit >  \loc 
    % & \lit >  \loc
    % \\
    grammaticalism
    \\
    \ \ \ \ weak
    & \loc > \glb > \lit 
    & \glb > \lit, \loc
    \\
    \ \ \ \ strong
    & \glb, \loc > \lit 
    & \glb, \loc , \lit
    \\
  \end{tabular}
  \caption{Predictions of the relevant theoretical positions. Listed
    are the predicted preference relations. If a reading is not
    listed, it is predicted to be unavailable.}
  \label{tab:predictions}
\end{table}

% \begin{table}[t]
%   \centering
%   \begin{tabular}{lcccccc}
%     & \multicolumn{4}{c}{availability} 
%     & \multicolumn{2}{c}{preference}
%     \\ \cmidrule(r){2-5}  \cmidrule(r){6-7}
%     theoretical position
%     & \as-\glb
%     & \as-\loc
%     & \es-\glb
%     & \es-\loc
%     & \as
%     & \es
%     \\ \midrule
%     traditionalism
%     \\
%     \ \ weak restricted 
%     & $\checkmark$
%     & $-$
%     & $\checkmark$
%     & $-$
%     & \lit > \glb 
%     & \lit > \glb
%     \\
%     \ \ weak unrestricted
%     & $\checkmark$
%     & $\checkmark$
%     & $\checkmark$
%     & $-$
%     & \lit > \glb > \loc 
%     & \lit > \glb
%     \\
%     \ \ strong restricted
%     & $\checkmark$
%     & $-$
%     & $\checkmark$
%     & $-$
%     & \glb > \lit 
%     & \glb > \lit
%     \\
%     \ \ strong unrestricted
%     & $\checkmark$
%     & $\checkmark$
%     & $\checkmark$
%     & $-$
%     & \glb > \loc > \lit 
%     & \glb > \lit
%     \\
%     conventionalism
%     \\
%     \ \ defaultism
%     & $-$
%     & $\checkmark$
%     & $-$
%     & $\checkmark$
%     & \loc >  \lit 
%     & \loc >  \lit
%     \\
%     \ \ literalism
%     & $-$
%     & $\checkmark$
%     & $-$
%     & $\checkmark$
%     & \lit >  \loc 
%     & \lit >  \loc
%     \\
%     grammaticalism
%     \\
%     \ \ weak
%     & $\checkmark$
%     & $\checkmark$
%     & $\checkmark$
%     & $\checkmark$
%     & \loc > \glb > \lit 
%     & \glb > \lit, \loc
%     \\
%     \ \ strong
%     & $\checkmark$
%     & $\checkmark$
%     & $\checkmark$
%     & $\checkmark$
%     & \glb, \loc > \lit 
%     & \glb, \loc >  \lit
%     \\
%   \end{tabular}
%   \caption{Predictions of the relevant theoretical positions}
%   \label{tab:predictions}
% \end{table}

\section{The Intonational Markedness Hypothesis}
\label{sec:inton-mark-hypoth}

Although traditionalism does not of itself predict local readings to
be available, many authors who adhere to a traditionalist position
acknowledge that local readings are available if and only if the
embedded scalar item is intonationally marked by contrastive stress
\citep[e.g.][]{Horn2006:The-Border-Wars,Geurts2009:Scalar-Implicat,Geurts2010:Quantity-Implic,GeurtsTielvan-Tiel2013:Scalar-expressi}. This
is taken to be continuous with other cases where pragmatic enrichments
can take scope under logical operators if and only if the relevant
focal accent is present (\ref{bsp:Stress}).

\begin{exe}
\ex \label{bsp:Stress} This coffee is not \textsc{hot}, it's \textsc{HOT}.
\end{exe}

\noindent The
intonational markedness hypothesis, as we will call it here, is then
an addition to the traditionalist's explanation of scalar implicatures
along the lines of the Gricean recipe. The crucial point is that under
this extra assumption, local readings are a marked phenomenon. If
supplemented with the intonational markedness hypothesis,
traditionalists would therefore predict that local readings are
available for \as- and \es-sentences, but only if intonationally
marked by contrastive stress; since this reading is marked, presumably
it would still be predicted to be dispreferred even if available.

Though there seems to be general consensus about such potential
prosodic effects, empirical evidence for this assumption is still
scarce. One notable exception comes from a pilot study reported in
\citet{Frazier08}, in which participants were instructed to read
embedded and non-embedded versions of English \as-sentences. In the
embedded sentences (e.g. {\it All of the students wrote some of the
  official memos}), {\it some} was either capitalized or
non-capitalized. Prior to the experiment, participants were informed
that capitalizing corresponded to contrastive stress. The task
consisted in a forced-choice questionnaire with two given alternatives
(such as {\it All of the students wrote some but not all of the
  official memos} vs. {\it All of the students wrote at least some of
  the official memos}). Interestingly, strengthened interpretations
occurred rather often across all sentences in this task (59 \%), but
were not affected by capitalizing.\fn{Note that, unfortunately, the
  authors do not provide percentages for the embedded conditions
  alone, but only the means across sentence types.}
\citeauthor{Frazier08} concludes that accentuation did not affect
their silent reading data, but raises the possibility that prosodic
effects may have been present at least in a subset of the items (see
footnote~1 in \citet{Frazier08}, p.~330).

Another study relevant for the present considerations comes from
\citet{SchwarzClifton2008:Strengthening-o}, examining the effects of
contrastive accentuation of "or" in non-embedded sentences like {\it
  Mary will invite Fred or/OR Sam to barbecue}. After listening to
these sentences, participants were presented two alternative readings
({\it 1. She will invite Fred or Sam or possibly both}, or {\it 2.
  She will invite Fred or Sam but not both}). Contratry to the
silent-prosodic effects reported in the pilot experiment by Frazier
and colleagues, the authors found a significant impact of overtly
realized contrastive accents on implicature processing: Besides a
general advantage for non-strengthened readings, focal accents
significantly increased the strengthened interpretation (from 16.4 \%
to 28.6 \%). These results are compatible with the "focus
strengthening hypothesis" proposed by the authors.

To sum up so far, though an effect of contrastive accentuation has
been claimed tobe essential for triggering local implicatures,
experimental evidence is currently mixed, suggesting potential effects
of presentation mode (visual vs.~auditory), item-specific
characteristics, as well as the specific construction under
investigation (e.g., embedded cases vs.~non-embedded cases such as in
\citet{SchwarzClifton2008:Strengthening-o}). It is for these reasons
that an experimental investigation into the availability of readings
of \as- and \es-sentences should present sentential material
auditorily and explicitly address the intonational markedness
hypothesis.

\section{Previous studies}
\label{sec:previous-studies}

In order to test the divergent predictions of the relevant theoretical
positions, a number of empirical studies have been carried out. In
this section, we will focus on the influential studies by
\citet{GeurtsPouscoulous2009:Embedded-Implic},
\citet{CliftonDube2010:Embedded-Implic} and
\citet{ChemlaSpector2010:Experimental-Ev}. Unfortunately, as we argue
here, the conjoined evidence from all of these studies is inconclusive
as to the availability and preference of relevant readings.

\subsection{\citet{GeurtsPouscoulous2009:Embedded-Implic}}
\label{sec:Geurts-and-Pouscoulous}

\citet{GeurtsPouscoulous2009:Embedded-Implic} conducted a
picture-verification task to find out whether local readings of \as-
and \es-sentences are available.\fn{We will focus here on a subset of
  the conditions tested by
  \citet{GeurtsPouscoulous2009:Embedded-Implic}. Actually, these
  authors did not use \es-sentences, but sentences where scalar
  \emph{some} was embedded under non-monotonic quantifier
  \emph{exactly two} (with appropriate pictures, of course). This case
  is a little more complex, but we will gloss over this here, treating
  their data, as if it was obtained for \es-sentences.} Subjects were
presented with pictures like those in Figure~\ref{fig:weak} and
\ref{fig:local} where the local reading gets a different truth-value
than the literal and the global reading. For \as-sentences, the local
reading is false for the critical picture in Figure~\ref{fig:weak}
whereas the literal and global readings are true; for \es-sentences
the local reading is true for the critical picture in
Figure~\ref{fig:local}, whereas the literal and global readings are
false.

The results of \citeauthor{GeurtsPouscoulous2009:Embedded-Implic} were
strikingly unambiguous: there were \emph{no} responses indicative of a
local reading; \emph{all} of the subjects judged \as-sentences true in
a situation like in Figure~\ref{fig:weak} and \emph{all} of the
subjects judged \es-sentences false in situations like
\ref{fig:local}.

These results were criticized on theoretical grounds (e.g.,
\citet{Sauerland2010:Embedded-Implic}, but see
\citet{Ippolito2010:Embedded-Implic} for support), as well as based on
empirical observations
\citep{CliftonDube2010:Embedded-Implic,ChemlaSpector2010:Experimental-Ev}. In
the following, we will review these empirical studies in some detail,
as they constitute an important background for our own experiment. We
will first consider a comment by
\citet{CliftonDube2010:Embedded-Implic}, which explicitly
distinguishes between the availability of a reading and its
preference.

\subsection{\citet{CliftonDube2010:Embedded-Implic}}
\label{sec:clifton-dube}

In a reply to \citeauthor{GeurtsPouscoulous2009:Embedded-Implic}'s
study \citeauthor{CliftonDube2010:Embedded-Implic} raised the question
whether the use of a picture-verification paradigm might have been
infelicitous for testing the availability of strong readings, at least
in the case of \as-sentences. Asking whether a sentence fits a picture
might have created a bias for accepting sentences also on a weaker and
probably dispreferred reading. Clifton and Dube's study was therefore
aimed at finding out about a potential preference relation between
local and literal readings. To this end, Clifton and Dube developed a
picture-choice task where subjects were presented with an AS-sentence
and a pair of pictures, hence introducing the option to choose between
different alternatives. Subjects were asked to ``indicate which shape
is best described by the sentence'' and could choose either picture,
or options `both' and `neither.' There were two versions of this
experiment, differing in which kind of picture pairs were presented on
critical trials. In version 1, the picture pair consisted of the weak
and strong situations in Figures~\ref{fig:weak} and
\ref{fig:strong}. The response percentages observed by
\citeauthor{CliftonDube2010:Embedded-Implic} were:

\begin{center}
  \begin{tabular}{cccc}
    weak & strong & both & neither
    \\ \midrule 
 3 & 39 & 57 & 1 
  \end{tabular}
\end{center}

\noindent That the majority answer is ``both'' could be taken as
evidence that the literal reading is the preferred one. But the almost
40\% of choices for the strong situation, so
\citeauthor{CliftonDube2010:Embedded-Implic} argue, is indicative of
the availability of the local reading. In version 2 of the experiment,
the picture pair consisted of the literal and weak situations in
Figures~\ref{fig:literal-AE} and \ref{fig:weak}. In this case,
response percentages were:

\begin{center}
  \begin{tabular}{cccc}
    weak & literal & both & neither
    \\ \midrule 
    28 & 6 & 50 & 17 
  \end{tabular}
\end{center}

\noindent Again, the majority response ``both'' might speak for a
preference for the literal reading, but, as
\citeauthor{CliftonDube2010:Embedded-Implic} propose, the 17\% of
``neither'' answers in this case again suggest that the local reading
is available. Taken together,
\citeauthor{CliftonDube2010:Embedded-Implic} take these results to
contradict \citeauthor{GeurtsPouscoulous2009:Embedded-Implic}'s
findings. Local readings are, after all, attested if subjects are
given a choice as to which situation they consider most fitting for an
\as-sentence.

The diverging results of
\citeauthor{GeurtsPouscoulous2009:Embedded-Implic} and
\citeauthor{CliftonDube2010:Embedded-Implic} seem to indicate that
participants' choices for readings of \as-sentences might be affected
by the specific experimental paradigm used \citep[but see][for
critique of the latter
design]{GeurtsTielvan-Tiel2013:Scalar-expressi}. Unfortunately,
\citet{CliftonDube2010:Embedded-Implic} did not test both universal
and non-monotonic quantifiers within the same
experiment. Additionally, it would be informative to also probe into
the availability of global readings. The study of
\citet{ChemlaSpector2010:Experimental-Ev} did both of that.

\subsection{\citet{ChemlaSpector2010:Experimental-Ev}}
\label{sec:Chemla-Spector}

\citet{ChemlaSpector2010:Experimental-Ev} also took issue with
\citeauthor{GeurtsPouscoulous2009:Embedded-Implic}'s design, arguing
that, firstly, the pictorial material used in
\citeauthor{GeurtsPouscoulous2009:Embedded-Implic}'s study was unduly
difficult; that, secondly, these pictures also may have failed to make
the local reading sufficiently relevant; and that, thirdly, the
restriction to a categorial choice (whether the sentence fits the
picture of not) may induce a bias against non-preferred readings in
cases where candidate readings stand in entailment relations
\citep[c.f.][for this latter
criticism]{Sauerland2010:Embedded-Implic}. To meet these potential
problems, \citet{ChemlaSpector2010:Experimental-Ev} presented subjects
with pictorial material like that in Figure~\ref{fig:Chemla-Spector},
which was assumed to be easier to assess and better at highlighting
the relevance of the local readings. Albeit in a different format, the
pictures used by \citeauthor{ChemlaSpector2010:Experimental-Ev} were
instantiations of the situation types in
Figures~\ref{fig:AS-distinguishing-pics} and
\ref{fig:ES-distinguishing-pics}. Additionally, subjects were asked,
not for categorial judgements, but for \textit{graded judgements}:
subjects could freely click on a scale, as shown in
Figure~\ref{fig:Chemla-Spector}, to indicate how much they considered
a picture fitting for a given sentence \citep[c.f.][for more on this
method]{Chemla2009:Presuppositions}.

\begin{figure}[t]
  \centering

  \subfloat[][Example of the \as-weak condition of
  \citet{ChemlaSpector2010:Experimental-Ev}]{
      \label{fig:Chemla-Spector}
  \begin{minipage}{0.45\linewidth}
    \begin{center}
      \includegraphics[scale=0.25]{../pictures/paper/Chemla_Spector_2010_Critical_AE.png}\\
      \includegraphics[scale=0.25]{../pictures/paper/Chemla_Spector_2010_RatingBar.png}
    \end{center}

  \end{minipage}
    }
   \begin{minipage}{6cm}
     \subfloat[][Results \as-condition]{
    \label{fig:CS-Results-as}
        \includegraphics[width=5.5cm]{../pictures/paper/Chemla_Spector_2010_Results_AE.png}
  }

  \subfloat[test][Results \es-condition]{
    \label{fig:CS-Results-es}
    \includegraphics[width=5.5cm]{../pictures/paper/Chemla_Spector_2010_Results_GE.png}
  }  
   \end{minipage}
        \caption{Example trial and results of
          \citeauthor{ChemlaSpector2010:Experimental-Ev}'s
          (\citeyear{ChemlaSpector2010:Experimental-Ev}) study.}
\end{figure}

% \begin{figure}[t]
%   \centering
%       \includegraphics[scale=0.25]{../pictures/paper/Chemla_Spector_2010_Critical_AE.png}\\
%       \includegraphics[scale=0.25]{../pictures/paper/Chemla_Spector_2010_RatingBar.png}
%         \caption{Example of critical trial for \as-condition by
%           \citeauthor{ChemlaSpector2010:Experimental-Ev}'s
%           (\citeyear{ChemlaSpector2010:Experimental-Ev})}
%   \label{fig:Chemla-Spector}
% \end{figure}

\citeauthor{ChemlaSpector2010:Experimental-Ev} hypothesized that the
degree to which a sentence is rated acceptable is proportional to the
number of available true readings. Observed averaged clicking
positions are shown in Figures~\ref{fig:CS-Results-as} and
\ref{fig:CS-Results-es}.
%
% \begin{figure}
%   \centering
%   \subfloat[][\as-condition]{
%     \label{fig:CS-Results-as}
%         \includegraphics[width=5.5cm]{../pictures/paper/Chemla_Spector_2010_Results_AE.png}
%   }
%   \subfloat[test][\es-condition]{
%     \label{fig:CS-Results-es}
%     \includegraphics[width=5.5cm]{../pictures/paper/Chemla_Spector_2010_Results_GE.png}
%   }
%   \caption{Results from
%     \citeauthor{ChemlaSpector2010:Experimental-Ev}'s
%     (\citeyear{ChemlaSpector2010:Experimental-Ev}) study}
%   \label{fig:Chemla-Spector-Results}
% \end{figure}
%
According to \citeauthor{ChemlaSpector2010:Experimental-Ev}, the
crucial piece of evidence for the availability of local readings for
\as-sentences is that these sentences yielded higher graded
acceptability scores for the strong situation than for the weak
situation (although these differ only with respect to the truth value
of the local reading). Evidence for the availability
of the local reading for \es-sentences comes from the difference
between the local and the literal situation. Strikingly, \es-sentences
received an average 73\% degree of acceptability in the local
situation although the literal and global readings are false in this
case.

\subsection{Reflection: Categorical Judgements \& Typicality}
\label{sec:local-read-categ}

Summing up so far, three previous studies have presented diverging
pieces of evidence concerning the status of putative local
readings. In their picture verification experiment
\citet{GeurtsPouscoulous2009:Embedded-Implic} did not observe any
truth value-judgments that would indicate local readings. On the basis
of this observation, they argue that these readings are simply not
available. On the other hand,
\citeauthor{CliftonDube2010:Embedded-Implic} and
\citeauthor{ChemlaSpector2010:Experimental-Ev} argue that the findings
of \citeauthor{GeurtsPouscoulous2009:Embedded-Implic} may be due to a
bias in their experimental design and employed different methods
intended to be sensitive enough to reveal effects of local readings:
either by a choice of the situation that better befits a given
sentence (\citeauthor{CliftonDube2010:Embedded-Implic}), or by a
graded truth-value judgment
(\citeauthor{ChemlaSpector2010:Experimental-Ev}). Both studies found
putative evidence in support of local readings.

To account for the diverging results we could draw a distinction
between \emph{categorical} truth-value judgments and other more
sensitive measures. If we do so, it could be hypothesized along with
\citeauthor{CliftonDube2010:Embedded-Implic} and
\citeauthor{ChemlaSpector2010:Experimental-Ev} that local readings are
available, but so dispreferred that they do not affect categorical
truth-value judgments (see, however, \citet{Crain1998} for arguments
in favour of truth-value judgments as a means of detecting highly
dispreferred readings).  It could then also be argued, in line with
traditionalism, that the apparent difference in sensitivity of the
different judgment types can be explained by the stipulation that the
results reported by \citeauthor{CliftonDube2010:Embedded-Implic} and
\citeauthor{ChemlaSpector2010:Experimental-Ev} reflect something other
than (strongly) truth-relevant speaker-intended meaning enrichments
after all, so that no case can be made for lexicalism or
grammaticalism on the basis of these data \citep[c.f.][for arguments
along these
lines]{Tielvan-Tiel2012:Embedded-Scalar,GeurtsTielvan-Tiel2013:Scalar-expressi}. To
circumvent this issue, a method for probing into the readings
for \as- and \es-sentences would therefore ideally evoke categorical
truth-value judgements: if evidence for local readings can also be
found using categorical truth-value judgements, it cannot be discarded
as a secondary non-truth-conditional effect.

Note, however, that the type of judgment is not the only crucial
difference between the methods of
\citeauthor{CliftonDube2010:Embedded-Implic} and
\citeauthor{ChemlaSpector2010:Experimental-Ev} on the one hand and
\citet{GeurtsPouscoulous2009:Embedded-Implic} on the other. These
studies also differed with respect to the \emph{types} of situations
(as distinguished in Figures~\ref{fig:AS-distinguishing-pics}
and~\ref{fig:ES-distinguishing-pics}) that were shown to the
participants during the course of the experiment. The presence or
absence of competing situations may have led to skewed distributions
of judgments for situations like in Figure \ref{fig:weak}. Such
effects might emerge simply because the type of sentence-picture
combinations shown during the experiment affects the standard of
judgment but could also --maybe even more elegantly-- be explained by
findings from \citet{Raffray2010} suggesting that exposing
participants to models may prime corresponding logical forms and
thereby shift reading preferences. Taking these possibilities into
account, it remains unclear whether the effects reported by
\citeauthor{CliftonDube2010:Embedded-Implic} and
\citeauthor{ChemlaSpector2010:Experimental-Ev} would not also show up
using ordinary picture verification, in the first place.

For these reasons, we cannot ascribe the effects reported by
\citeauthor{CliftonDube2010:Embedded-Implic} and
\citeauthor{ChemlaSpector2010:Experimental-Ev} to the availability of
local readings with
certainty. \citet{ChemlaSpector2010:Experimental-Ev} themselves
acknowledge the possibility that the \emph{typicality} of a picture
with respect to some meaning may affect graded truth-value
judgments. Strengthening this possibility, they show that graded
judgments of \as-sentences differ significantly for different
instantiations of the situation types from
Figure~\ref{fig:AS-distinguishing-pics}, that vary in the amount of
connections without changing the assignment of truth values to
candidate readings. \citeauthor{ChemlaSpector2010:Experimental-Ev}
suggest that typicality of the pictorial material can account for
these differences, but submit that this does not explain away the high
acceptance of pictures like Figure \ref{fig:weak}. The latter point is
disputed by \citet{Tielvan-Tiel2012:Embedded-Scalar}, who demonstrates
that a huge chunk of variance in the data of
\citeauthor{CliftonDube2010:Embedded-Implic} and
\citeauthor{ChemlaSpector2010:Experimental-Ev} on \as-sentences can be
explained as typicality effects, dispensing the need to appeal to the
distribution of different readings. In support of this idea,
\citet{Tielvan-Tiel2012:Embedded-Scalar} elicited what he calls the
typicality structures associated with the quantifiers {\it all} and
{\it some} \citep[as done also by][]{DegenTanenhaus2011:Making-Inferenc} and predicted the judgments of \as-sentences obtained by
\citeauthor{ChemlaSpector2010:Experimental-Ev} based on these
data. Thereby, he obtained an excellent model fit.

We are generally sympathetic towards
\citeauthor{Tielvan-Tiel2012:Embedded-Scalar}'s innovative line of
reinterpretation of the data, but note, as
\citet{ChemlaSpector2010:Experimental-Ev} already observed, that his
typicality-based explanation does not extend to \es-sentences in an
obvious way. Moreover, typicality itself is not a satisfactory
primitive in a putative explanation of the pragmatic use of
quantifiers. Plausibly, the reason why quantifier \emph{some} is more
typical for some set sizes than others is at least dependent on, if not
entirely derived from, its contrast with other conceivable
alternatives, e.g., numerals within the subitizing range, as well as
\emph{few}, \emph{most} or \emph{many}
\citep[c.f.][]{DegenTanenhaus2012:Processing-Scal}. For these reasons,
we believe that typicality and other pictorial effects need to be
taken seriously in the experimental design, but do not give a
satisfactory account of the observed variation all by themselves.

To conclude, although considerable progress has been made in
understanding which factors have to be controlled for in order to
asses the status of local readings, the evidence about availability,
preference and nature of local readings is still inconclusive. What is
needed to complement the picture is data that is as independent as
possible of potentially conflating properties of the pictorial
material and that is nonetheless sensitive enough to detect local
readings even if they are dispreferred. Given the logical dependencies
of the relevant readings this is difficult, but, as we will argue
below, not impossible.

On top of this, it would be desirable to obtain further information
about the relative preferences among all putative readings, so as to
be able to decide between theoretical positions, as outlined in
Section~\ref{sec:theories-predictions}. Neither of the previous
studies gives us that, either because they simply have not
investigated all the relevant comparisons
(\citet{GeurtsPouscoulous2009:Embedded-Implic},
\citet{CliftonDube2010:Embedded-Implic}), or because they cannot
safely derive the required information for danger of conflating
preferences for readings with preferences for pictures
(\citet{ChemlaSpector2010:Experimental-Ev}).

Finally, the above-mentioned studies did not control for possible
conflating effects of ``silent prosody''. A growing number of evidence
suggests that numerous factors might influence accent placement and
prosodic phrasing even while reading, amongst them default
accentuation, constituent length, rhythmic phenomena, and individual
variation \citep[e.g.][]{Augurzky08,Bader98,Fodor98,Fodor02,Kentner12,
  Steinhauer01}. If we adopt the intonational markedness hypothesis
described in Section~\ref{sec:inton-mark-hypoth}, as many
traditionalists do, we predict that the availability of a local
reading hinges on the realization of contrastive stress on the scalar
item
\citep[e.g.][]{Horn2006:The-Border-Wars,Geurts2009:Scalar-Implicat,Geurts2010:Quantity-Implic,Tielvan-Tiel2012:Embedded-Scalar,GeurtsTielvan-Tiel2013:Scalar-expressi}.
But if intonation can have this role, it is necessary to control for
this potential variation due to potential differences in silent accent
placement. Ideally, therefore, we should present sentences auditorily
and test the intonational markedness hypothesis explicitly by
manipulating contrastive stress systematically.

\section{An Incremental Verification Task}
\label{sec:exp}

\subsection{Design}
\label{sec:design}

\paragraph{General Idea.} In order to test the availability and
preferences of different readings of \as- and \es-sentences in German, 
we were looking for a way to unambiguously map responses from a picture
verification task to specific readings. We were particularly
interested in the question whether local readings would be observed at
all, and whether their presence or absence would be modulated by
a contrastive accent on the scalar item. Moreover, we were  
interested in obtaining information about the relative preferences 
of the different readings. In addition, we wanted
to minimize all effects of alleged typicality of pictures. To achieve this, we designed an experiment employing a
modified version of picture verification, namely the
\mymark{incremental verification task} (\acro{ivt}, SEE CONROY
????). The general idea is that subjects are requested to judge
sentence material based on pictures that do not necessarily contain
all the information necessary to judge a certain reading true or
false. In that case, participants can demand that more information be
revealed. Participants are instructed to make a truth-value judgment
as soon as they are able to. When they do, the trial ends.

\paragraph{Target Conditions.} In the present study, participants were
presented with pictures depicting a set of four identical central
elements (e.g., letters), which could be connected to surrounding
elements (e.g., triangles). Initially, any potential connections
between central and surrounding elements were covered by dark gray
color (see Figure~\ref{fig:InitialPicture}).
%
\begin{figure}
  \centering
      \fbox{\includegraphics[width=5cm]{../pictures/paper/ae_01_1.pdf}}
      \caption{Initial picture presented to participants where all
        possible connections from each letter to its surrounding
        triangles are covered.}
  \label{fig:InitialPicture}
\end{figure}
%
Sentences to be judged 
%were in German (see (\ref{ex:as}) and (\ref{ex:es}) below) and 
were presented auditorily, and participants were asked to listen to a
given sentence and then uncover the potential links step by step until
they felt able to give a truth value judgment. Three options were
available for participants at each step: (i) demand more information,
(ii) judge the sentence as true, and (iii) judge the sentence as
false. Trials ended with a truth-value judgement, and a judgement was
required to end the trial when all connections had been
revealed. Importantly, each of the three potentially available
readings corresponded to a specific step in the uncovering process
where the truth value of that reading (and only of that reading) could
be assessed for the first time. We refer to this step in the sequence
as the {\it critical position} of a given reading. The critical
position and the corresponding truth-value judgment differed between
\as- and \es-sentences, as described presently. (This is, partly,
because of the logical dependencies between readings, and, partly,
also to rule out positional biases.)

Consider the German \as-sentence in (\ref{ex:as}), included in our study.
\begin{exe}
\ex \gll \mymark{Alle} diese Briefe sind mit \mymark{einigen} ihrer Dreiecke
  verbunden.\label{ex:as}\\
All these letters are with some their triangles connected.\\
\trans All of these letters are connected to some of their triangles.
\end{exe}
Figure~\ref{fig:exseqAS} shows the three corresponding critical
positions for this sentence in the order of the uncovering
process. First, the situation in Figure~\ref{fig:exseqAS1} becomes
available, which is true under a literal reading, while the local and
global readings cannot be evaluated yet. Second, in
Figure~\ref{fig:exseqAS2}, the literal reading is still true, but now
the global reading can be additionally confirmed by a \emph{true}
response, as all the connections of one of the letters are now
uncovered, and it is now visible that one of the letters is not
connected to all of its triangles. Finally, decisions concerning the
local reading are possible as soon as all connections have been
uncovered. In this case, the critical position shown in
Figure~\ref{fig:exseqAS3} is incompatible with a local reading: One of
the central elements is linked to all of its surrounding elements,
thus yielding to \emph{false} answers if participants indeed adopted
the local reading. Note that associating the final position with
\emph{false} answers allows us to exclude both literal and global
readings, which correspond to a \emph{true} answer at this
position. In sum, \emph{true} or \emph{false} answers on particular
positions in the incrementally revealed picture sequence can be mapped
uniquely to candidate reading. All other \emph{true} or \emph{false}
answer were counted as errors. The mapping between responses and
readings is summarized in Table~\ref{tab:mappingAS}.

\begin{figure}[ht]
	\centering
	\subfloat[][Step 1]{ 
		\fbox{\includegraphics[width=3.5cm]{../pictures/paper/ae_3_v_l.pdf}}
	    \label{fig:exseqAS1}
	}
	\subfloat[][Step 2]{
		\fbox{\includegraphics[width=3.5cm]{../pictures/paper/ae_5_v_l.pdf}}
	    \label{fig:exseqAS2}
	}
	\subfloat[][Step 3]{ 
		\fbox{\includegraphics[width=3.5cm]{../pictures/paper/ae_7_v_l.pdf}}
	    \label{fig:exseqAS3}
	}
	\caption[]{Example sequence for \as-sentences}
	\label{fig:exseqAS}
\end{figure}

\begin{table}[ht]
	\centering
	\subfloat[][\as-Sentences]{ 
\begin{tabular}{cccc}
&more info&true&false\\ \midrule
Step 1&-&\lit&error\\
Step 2&-&\glb&error\\
Step 3&n.a.&error&\loc
\end{tabular}

	    \label{tab:mappingAS}
	} \qquad
	\subfloat[][\es-Sentences]{
\begin{tabular}{cccc}
&more info&true&false\\ \midrule
Step 1&-&error&\glb\\
Step 2&-&error&\lit\\
Step 3&n.a.&\loc&error
\end{tabular}
	    \label{tab:mappingES}
	}
\caption{Mapping from response types to readings. Notice that on steps
3 the option ``more info'' is not available because the whole picture
has been revealed.}
\label{tab:subfigureExample}
\end{table}

\es-sentences, due to the non-linear relationship of candidate
readings described in Section~\ref{sec:get-know-your}, require a
slightly different sequence of unfolding which yields a different
order in which truth-value judgments can be made.
%
\begin{figure}[ht]
	\centering
	\subfloat[][Step 1]{ 
		\fbox{\includegraphics[width=3cm]{../pictures/paper/ge_3_v_l.pdf}}
	    \label{fig:subfig1}
	}
	\subfloat[][Step 2]{
		\fbox{\includegraphics[width=3cm]{../pictures/paper/ge_4_v_l.pdf}}
	    \label{fig:subfig2}
	}
	\subfloat[][Step 3]{ 
		\fbox{\includegraphics[width=3cm]{../pictures/paper/ge_6_v_l.pdf}}
	    \label{fig:subfig1}
	}
	\caption[]{Example sequence for \es-sentences}
	\label{fig:exseqES}
\end{figure}
%
For a sentence like (\ref{ex:es}) we obtain an unambiguous
mapping between truth values and readings with respect to the sequence
in \ref{fig:exseqES}.
\begin{exe}
\ex \label{ex:es} \gll \mymark{Genau} eine der Glocken ist mit
  \mymark{einigen} ihrer Halbkreise verbunden.\\ 
  Exactly one of-the bells is with some its semicircles connected.\\
  \trans Exactly one bell is connected to some of its semicircless.
\end{exe}
At the first critical position shown in Figure~\ref{fig:subfig1}, the
global reading corresponds to a \emph{false} response, as one of the
bells is already linked to all of its semicircles. The literal reading
can be evaluated at the critical position shown in
Figure~\ref{fig:subfig2}, where it would trigger a \emph{false}
response, as there are two bells that are connected to at least some
of their surrounding elements. Finally, confirming \ref{fig:subfig1}
by a \emph{true} response corresponds to the local reading, as one of
the bells is linked to some but not all of its semicircles, while all
of the other bells are linked to all of them. Note that again, global
and literal readings differ from local readings with respect to their
truth values. Mappings between responses and readings are summarized
in Table \ref{tab:mappingES}.

\paragraph{Positional controls.} In order to make sure that subjects
understood the task, i.e., gave truth-value judgements neither sooner
not later than at the first possible position in a sequence, we
included a set of control conditions. These also controlled for
response biases. For each of the three readings in the \as- and
\es-conditions, we constructed an unambiguous control sentence as in
(\ref{bsp:controls-as}) and (\ref{bsp:controls-es}), requiring the
same judgement at the identical position in the same sequence used for
the respective targets. For instance, example
(\ref{bsp:controls-as-1}) requires a \emph{true}-response analogous to
the \as-sentence in (\ref{ex:as}) under its literal reading,
(\ref{bsp:controls-as-2}) corresponds to the \as-sentence under its
global reading and (\ref{bsp:controls-as-3}) corresponds to its local
reading. With regard to the \es-sentences, controls like
(\ref{bsp:controls-es-1}) correspond to the global reading,
(\ref{bsp:controls-es-2}) to the literal and (\ref{bsp:controls-es-3})
to the local reading. If, independently of sentence meaning, there was
any bias to respond in a certain way at any point in the sequences,
such as to preferably unravel the whole picture, this should affect
control sentences to the same degree as it affects target conditions.

\begin{exe}
  \ex \label{bsp:controls-as}
    \begin{xlist}
\ex \label{bsp:controls-as-1} \gll Alle Briefe sind mit mindestens drei ihrer Dreiecke verbunden.\\
  All letters are with at-least three their triangles connected.\\
  \trans All letters are connected to at least three of their triangles.
\ex \label{bsp:controls-as-2} \gll Mindestens ein Brief ist mit genau f\"unf seiner Dreiecke verbunden.\\
  At-least one letter is with exactly five his triangles connected.\\
  \trans At least one letter is connected with exactly five of its triangles.
\ex \label{bsp:controls-as-3} \gll Jeder Brief ist mit mindestens vier seiner Dreiecke verbunden.\\
  Every letter is with at-least four his triangles connected.\\
  \trans Every letter is connected to at least four of its triangles.
\end{xlist}
\end{exe}

\begin{exe}
\ex \label{bsp:controls-es}
  \begin{xlist}
\ex \label{bsp:controls-es-1} \gll Alle Glocken sind mit weniger als
  vier ihrer Halbkreise verbunden. \\ 
All bells are with fewer than four their semicircles connected.\\
\trans All bells are connected with fewer than four of their semicircles.  
\ex \label{bsp:controls-es-2} \gll Alle Glocken sind mit allen ihren Halbkreisen verbunden.\\
All bells are with all their semicircles connected.\\
\trans All bells are connected to all of their semicircles. 
\ex \label{bsp:controls-es-3} \gll Mindestens drei Glocken sind mit
  allen ihren Halbkreisen verbunden.\\ 
  At-least three bells are with all  their semicircles connected.\\
  \trans At least three bells are connectd with all of their semicircles.
\end{xlist}
\end{exe}

\paragraph{Preference-related controls.} In order to test whether the
order of critical positions for different readings within a sequence
had an influence on responses, we included a second type of control
conditions. These conditions also controlled whether prosodic
information can, in principle, shift reading preferences in the
present task. Note that for our target sentences, the logical
entailment relations between readings always require local readings to
be evaluated at the end of a sequence. It could thus in principle be
possible that no subject ever reaches this point, because they gave
truth-value judgements earlier, thereby ending the trial. In that case
it would be unclear if these decisions have been affected by a general
unavailability of local readings or by the fact that one of the
earlier presented readings is the preferred one. By including globally
ambiguous structures with a known preference over available readings,
we thus intended to test whether participants in our task occasionally
choose dispreferred readings, even if they were available only at a
critical position following the preferred readings.

We therefore included globally ambiguous \emph{late-closure
  structures} as in (\ref{bsp:target-related}), which have been
repeatedly shown to exhibit interpretive preferences.
\begin{exe}
\ex \gll Der Brief ist mit Kreisen und Vierecken mit Sonnen
  verbunden. \label{bsp:target-related}\\
The letter is with circles and squares with suns connected.\\
The letter is connected with circles and squares with suns.
\begin{xlist}
  \ex \label{bsp:target-related-LC} The letter is connected with squares containing suns, and it is
    also connected with circles. \hfill{(\lc)}
  \ex \label{bsp:target-related-EC} The letter is connected with circles and squares, both of which
    are containing suns.  \hfill{(\ec)} 
\end{xlist}

\ex \gll Der Brief ist mit Kreisen, die Sonnen beinhalten und Vierecken
  verbunden. \label{bsp:target-related2}\\
The letter is with circles, which suns contain and squares connected.\\
The letter is connected with circles containing suns, and with squares.

\end{exe}
In late-closure sentences, a modifier such as a relative clause or a
prepositional phrase can be attached to one of two preceding
hosts. For instance, in (\ref{bsp:target-related}), the prepositional
phrase \emph{with suns} can be attached to the preceding noun
\emph{squares}, resulting in the so-called \emph{late-closure} or
\emph{low-attachment} reading
((\ref{bsp:target-related-LC}). Alternatively, the \acro{pp} can be
attached to the whole conjoined \acro{np} \emph{circles and squares},
corresponding to the \emph{early-closure} or \emph{high-attachment}
reading (\ref{bsp:target-related-EC}).  Late-closure readings have
been observed to be preferred over the early-closure reading for
sentences comparable to those in (\ref{bsp:target-related})
\citep[e.g.][]{Frazier87,Frazier82}.

For each sentence, two sequences were designed. In the first one, the
critical position associated with the dispreferred early-closure
reading preceded the critical position for the preferred late-closure
reading. In this case, a partly covered picture as in
Figure~\ref{fig:exec1} preceded a picture as in \ref{fig:exec2}.
%
\begin{figure}[ht]
	\centering
	\subfloat[][Step 1]{ 
		\fbox{\includegraphics[width=5.5cm]{../pictures/paper/ec_01_3.pdf}}
	    \label{fig:exec1}
	}
	\subfloat[][Step 2]{
		\fbox{\includegraphics[width=5.5cm]{../pictures/paper/ec_01_5.pdf}}
	    \label{fig:exec2}
	}
	\caption[]{Critical steps of a sequence for sentence
          (\ref{bsp:target-related}) where the early-closure reading
          (\ref{bsp:target-related-EC}) can be judged first.}
	\label{fig:exec}
\end{figure}
%
If participants adopted an early-closure reading as in
(\ref{bsp:target-related-EC}), a \emph{false}-judgment would be
expected at the position illustrated in \ref{fig:exec1}, as none of
the circles connected to the letter contain any suns. Under the
late-closure reading in (\ref{bsp:target-related-LC}), in contrast, a
\emph{true}-response would be expected no sooner than in the step
shown in Figure~\ref{fig:exec2}. The order of possible judgements is
reversed in the second kind of sequence including pictures like
\ref{fig:exlc1} followed by pictures like \ref{fig:exlc2}. Here, the
late-closure reading in (\ref{bsp:target-related-LC}) can be judged
first requiring a \emph{true}-response on \ref{fig:exlc1}. The
early-closure reading in (\ref{bsp:target-related-EC}) can only be
judged later in the sequence. When we reach \ref{fig:exlc2} a
\emph{false}-response is expected under an
early-closure/high-attachment reading.

\begin{figure}[ht]
	\centering
	\subfloat[][Step 1]{ 
		\fbox{\includegraphics[width=5.5cm]{../pictures/paper/lc_01_3.pdf}}
	    \label{fig:exlc1}
	}
	\subfloat[][Step 2]{
		\fbox{\includegraphics[width=5.5cm]{../pictures/paper/lc_01_6.pdf}}
	    \label{fig:exlc2}
	}
	\caption[]{Critical steps of a sequence for sentence
          (\ref{bsp:target-related}) where the late-closure reading
          (\ref{bsp:target-related-LC}) can be judged first.}
	\label{fig:exec}
\end{figure}

\paragraph{Testing Effects of prosody.} Whereas contrastive accents
are usually assumed to be realized by a special contour in English
(i.e. L+H*, see \citet{Pierrehumbert90}), the exact phonological
classification of contrast in German is still under debate
\citep[see][]{Uhmann91,Fery93,Grabe98,Toepel06, Sudhoff10}.  These
approaches have in common that from an acoustic perceptive, prosodic
prominence of contrastively accented constituents is realized by

\begin{enumerate}[(i)]
\item a higher F0 maximum or a higher pitch range (difference 
between minimal and maximal F0 values), and
\item a longer duration of the accented element as opposed to its 
non-accented counterpart.
\end{enumerate}

\noindent We therefore recorded sentences by a trained speaker
familiar with the concept of contrastive focus and instructed her to
produce contrast as described in 1 and 2. Both \as-sentences and
\es-sentences were recorded in two versions each. In the accented
version, a pitch accent was placed on {\it einige}. The second version
had neutral prosody.  If accentuation is the driving force for local
readings, we would expect a higher proportion of local readings in the
former than in the latter version of the sentences.

Preference-related controls also differed prosodically. Modifier
attachment ambiguities comparable to the examples in
(\ref{bsp:target-related}) and (\ref{bsp:target-related2}) have been
shown to be sensitive to differences in prosodic phrasing. Though
there has been discussion as to whether speakers reliably
\textit{produce} disambiguating cues in such constructions
\citep[e.g.][]{Allbritton96,Kraljik05, Snedeker03, Schafer00}, overt
prosodic boundaries are generally acknowledged to guide interpretive
processes of late-closure sentences in \textit{comprehension}
\citep[e.g.][]{Steinhauer99, Augurzky06}.  Specifically, it has been
shown that prosodically separating a modifier from the directly
preceding material supports an early-closure reading, whereas
separating the two potential attachment sites is supportive of a late
closure reading. Thus, a phrase boundary between {\it squares} and
{\it with} in (\ref{bsp:target-related}) corresponds to a
high-attachment reading, whereas a prosodic phrase boundary following
{\it circles} corresponds to a low-attachment reading.

To test whether prosodic information is taken into account in our
task, we presented sentences in both of these prosodic variants. In
addition, in order to test prosody-independent preferences, we also
included a neutral version without any pronounced boundary. Further,
we also presented unambiguous sentences corresponding to the
early-closure reading. As in the case of the \as- and \es-sentences,
the unambiguous counterparts served as a baseline to control for
response biases that are independent of sentence meaning.

\subsection{Procedure}
\label{sec:procedure}

\subsection{Materials}
\label{sec:materials}

\paragraph{Target sentences and positional controls}

We constructed a set of 15 items for German \as- and \es-sentences,
respectively, analogous to the examples in \ref{ex:as} and
\ref{ex:es}. Sentences were introduced by a subject {\small
  \acro{DP}}, including the quantifiers {\it alle diese} ("all of
these") or {\it genau eine(r) der} ("exactly one of these"), as well
as the head noun which denoted different icons, such as letters or
bells. Subjects were followed by an auxiliary, and a {\small
  \acro{PP}} containing {\it einige}, a possessive pronoun, and a noun
denoting a geometrical object, e.g. {\it einigen seiner
  Dreiecke/Quadrate} ("some of its triangles/squares"). The possessive
pronoun was used in order to fix relative quantifier scope to surface
scope, thus ensuring that {\it einige} has embedded status. The last
word was the main verb {\it verbunden} ("connected").  Each sentence
was recorded in a stressed and an unstressed version of scalar {\it
  einige}. As described above, three control
sentences were created for each experimental item, corresponding to
the critical positions in the course of uncovering the accompanying
sequences.  The positional controls were recorded with
neutral prosody.  Items were evenly distributed across 5 lists using a
Latin square design.

\paragraph{Preference-related control sentences}
For controlling preferences, 28 sentences that were ambiguous between
a late-closure and an early-closure reading, as well as 28 of their
disambiguated counterparts were constructed as described above. In
these sentences, subject {\small \acro{DP}}s were always denoting
icons and were followed by an auxiliary. In the ambiguous sentences,
the auxiliary was followed by two {\it mit}-{\small \acro{PP}}s.  The
first of these {\small \acro{PP}}s consisted in a coordination of two
nouns denoting geometrical objects, whereas the second PP denoted
icons. In contrast to these sentences, the unambiguous sentences
contained only one PP with a conjunction the first part of which was
modified by a relative clause.

\paragraph{Acoustic properties}
A phonetically trained female native speaker of German was instructed
to realize two prosodic versions of each target sentence, and three
versions of the ambiguous preference-related controls. In addition to these
sentences, the unambiguous preference-related controls, the positional controls
and a set of unrelated fillers were also recorded. For these constructions, the speaker was instructed
to realize prosodic contours as neutral and natural as possible.

For all target sentences (\as and \es), the determiner \emph{einige} was
produced with a contrastive pitch accent as well as with a neutral
accent. Contrastive accents in German are realized by an increase in F0 (range)
as well as by a longer duration (see Section~\ref{sec:design} above). A set of 15 
experimental items was recorded for each condition (\as vs.~\es, \emph{accented} 
vs.~\emph{unaccented}), resulting in a total number of 60 target sentences.

In contrast to the accent manipulation, the ambiguous preference-related controls
differed with respect to prosodic phrasing. Prosodic phrase boundaries
in German are realized by a rise in F0 as well as by a durational
increase on the final part of the constituent preceding the boundary
(prefinal lengthening) plus an optional pause \citep[e.g.][]{Vaissiere83,Fery93}.
Boundaries for these control sentences were either realized at the position 
separating the second {\small \acro{PP}}
from the preceding material (\emph{late boundary}, corresponding to a
high-attachment reading) or directly preceding the second conjunct in
the first {\small \acro{PP}} (\emph{early boundary}, corresponding to a
low-attachment reading). As the prosodic realization of the targets
involved the comparison between an accented and a neutral variant, we
also included a third version of the ambiguous preference related controls, in which our
speaker produced the sentences without any pronounced
boundaries. For each prosodic variant, a set of
30 items was read, yielding 90 preference-related fillers.

Altogether a total number of 300 sentences consisting of 60 target sentences, 90 ambiguous preference-related controls, 30 unambiguous preference-related controls, 90 positional control sentences and 30 unrelated fillers was recorded. The session was recorded in an acoustically
shielded booth (44.1\,kHz sampling rate, 16 bit amplitude resolution).

Before entering the judgment task, experimental items and
preference-related fillers were analyzed with respect to their acoustic
properties. As both accented elements as well as prosodic boundaries
were expected to differ with respect to their F0 and/or durational
properties, we calculated durational values as well as difference
values between minimal and maximal F0 for each word. As targets
slightly differed with respect to the total number of words as well as
with respect to certain lexical properties (i.e., \emph{seinen}
vs.~\emph{ihren}, we considered the following analysis regions:

\begin{exe}
  \ex
    \begin{xlist}
      \ex $|_{\text{R}1}$~Alle  	\ $|_{\text{R}2}$~diese 
      \ $|_{\text{R}3}$~$\acro{np}_1$  
      \ $|_{\text{R}4}$~sind  \ $|_{\text{R}5}$~mit  \
      $|_{\text{R}6}$~einigen  \ $|_{\text{R}7}$~ihrer 
       \ $|_{\text{R}8}$~{$\acro{np}_2$}  \ $|_{\text{R}9}$~verbunden.
    \ex       $|_{\text{R}1}$~Genau einer  	\ $|_{\text{R}2}$~der 
      \ $|_{\text{R}3}$~$\acro{np}_1$  
      \ $|_{\text{R}4}$~ist  \ $|_{\text{R}5}$~mit  \
      $|_{\text{R}6}$~einigen  \ $|_{\text{R}7}$~seiner 
       \ $|_{\text{R}8}$~{$\acro{np}_2$}  \ $|_{\text{R}9}$~verbunden.
    \end{xlist}
    % \begin{tabbing}
    %   $|_{\text{R}1}$ Genau einer \= 	\ $|_{\text{R}2}$ diese \=
    %   \ $|_{\text{R}3}$ \acro{np}$_1$ \= 
    %   \ $|_{\text{R}4}$ sind \= \ $|_{\text{R}5}$ mit \= \
    %   $|_{\text{R}6}$ einigen \= \ $|_{\text{R}7}$ seiner 
    %   \= \ $|_{\text{R}8}$  \acro{np}$_2$ \= \ $|_{\text{R}9}$
    %   verbunden. \kill
    %   $|_{\text{R}1}$ Alle \> 	\ $|_{\text{R}2}$ dieser \>
    %   \ $|_{\text{R}3}$ \acro{np}$_1$ \> 
    %   \ $|_{\text{R}4}$ sind \> \ $|_{\text{R}5}$ mit \> \
    %   $|_{\text{R}6}$ einigen \> \ $|_{\text{R}7}$ ihrer 
    %   \> \ $|_{\text{R}8}$  \acro{np}$_1$ \> \ $|_{\text{R}9}$
    %   verbunden. \\
    %   $|_{\text{R}1}$ Genau einer \> 	\ $|_{\text{R}2}$ der \>
    %   \ $|_{\text{R}3}$ \acro{np}$_1$ \> 
    %   \ $|_{\text{R}4}$ ist \> \ $|_{\text{R}5}$ mit \> \
    %   $|_{\text{R}6}$ einigen \> \ $|_{\text{R}7}$ seiner 
    %   \> \ $|_{\text{R}8}$  \acro{np}$_1$ \> \ $|_{\text{R}9}$ verbunden.
    % \end{tabbing}
\end{exe}

\noindent Note that differences between Regions 1, 2, 4 and 7 can be expected
due to lexical differences between the \as and \es-conditions. As
preference-related fillers did not differ with respect to their lexical
properties, we carried out word-by-word analyses for these
conditions. Durational values included the respective word plus any
following silent interval. We did not include the
disambiguated fillers in these analyses as they involved very
different sentence types (i.e., constructions involving prepositional
phrases vs.~relative clauses).

\begin{exe}
  \ex $|_{\text{R}1}$ \acro{D}  	\ $|_{\text{R}2}$ $\acro{np}_1$
      \ $|_{\text{R}3}$ ist 
      \ $|_{\text{R}4}$ mit  \ $|_{\text{R}5}$ $\acro{np}_2$   \
      $|_{\text{R}6}$ und  \ $|_{\text{R}7}$ $\acro{np}_3$ 
       \ $|_{\text{R}8}$  mit  \ $|_{\text{R}9}$ $\acro{np}_4$ \ $|_{\text{R}10}$~verbunden.
\end{exe}

For the durational analyses, constituents were automatically labeled
by the \emph{Aligner} tool \citep{Rapp98}, and the
obtained values were manually corrected afterwards. For the targets,
two-factorial \acros{anova} with the factors \acro{Quantifier} (\emph{all}
vs.~\emph{exactly one}) and \acro{Prosody} (accented vs.~unaccented) were
carried out. For the preference-related fillers, we carried out
one-factorial \acros{anova} with the factor \acro{Prosody} (early
boundary, late boundary, neutral prosody). F0 values were extracted by
means of special Praat scripts
(\url{http://www.fon.hum.uva.nl/praat/}). For the present analyses,
differences between minimal and maximal F0 values for each region or
word were calculated. Again, two-factorial \acros{anova} were carried
out for statistical comparison of the target sentences, and one-factorial
\acros{anova} were carried out for preference-related controls.

The details of these analyses are found in
Appendix~\ref{sec:audit-sent-mater}. \dn{Petras Kommentar: Vielleicht
  sollten wir hier jeweils 1 Figure für akzentuiert vs. nicht und
  boundary vs. nicht reinbringen. Fabian meint zurecht, dass man nicht
  einfach die Ergebnisse nur in den Appendix packen kann. Es wären
  dann halt 4 Figures, da wir ja F0 und duration ansehen, und wir
  würden nur die experimentell interessanten veranschaulichen und die
  Statistik dann im Appendix lassen - was meint Ihr? --- Micha sagt ``Ja!''}

In sum, our speaker reliably
produced (i) differences in accent realization for the target
sentences and (ii) the expected boundary realizations for the
preference-related controls. Whereas accented elements clearly
differed from their unaccented counterpart by showing an increase in
duration and F0 range, prosodic boundaries for our preference-related
fillers were realized by pre-final lengthening (i.e., an increase in F0
and duration at the position preceding the boundary).

\paragraph{Pictures.}
The visual sequences accompanying each sentence consisted of pictures
comparable to those illustrated in Figure \ref{fig:InitialPicture}. In
each picture, four alternative sub-scenarios were presented, in which
an icon was surrounded by six geometrical shapes. Depending on the
sentence material, shapes and icons differed across sentence types and
conditions. For example, in our target conditions, 4 identical icons
were surrounded by geometrical shapes that were all of the same type,
whereas various icons and surrounding elements were used in each
sequence of the preference-related control structures. As described in
section xx, grey colour initially covered the links between icons and
shapes. As the connections were to be uncovered step by step,
different pictures were designed corresponding to different stages in
the uncovering process. Note that the single conditions impose
distinct requirements on sequences. For instance, inherent entailment
relations between \as-sentences and \es-sentences lead to different
critical positions at which participants can make their judgments. We
will therefore describe the sequences for each condition separately.

For \as-sentences, as well as for the corresponding positional control
sentences described in Section \ref{sec:design}, we designed a set of
seven pictures for each trial. In each sequence, the critical position
associated with the literal meaning was presented at the third step,
the position associated with the global meaning was shown at the fifth
step, and the local reading was presented at the final step (see
Figure \ref{fig:exseqAS} for an illustration of the critical
positions). Pictures that were interspersed between these positions
served as spillover-pictures. These were used in order to control for
potentially delayed judgments and they additionally served as
distractor items.

Sequences in the \es conditions (and their respective positional
controls) consisted of a set of five pictures with the following
critical positions. The critical positions as illustrated in Figure
\ref{fig:exseqES} are associated with the second picture (global
reading), the third picture (literal reading), as well as with the
final picture (local reading).

For the preference-related controls, we used two types of
sequences. First, a sequence in which the crirical position for the
(preferred) late-closure reading was presented preceding the critical
position for the early-closure reading ({\it LC sequence}), and
second, a sequence in which the readings were presented in reversed
order ({\it EC sequence}. In the LC sequences, seven pictures were
presented, with the LC reading given at the third picture, and the EC
reading at the sixth picture. EC sequences consisted of five pictures,
in which the critical positions were realized at the third picture (EC
reading) and at the fifth picture (LC reading).

Finally, we added pictures compatible with unrelated fillers with
critical position at random uncovering stages, including cases where
no uncovering was needed for a truth-value judgement.

\paragraph{Lists.}
Experimental items and preference-related fillers were evenly
distributed across lists. For the \as- and \es-sentences together with
their respective controls, five lists were used employing a Latin
square design. Picture sentence-pairs for the preference-related
controls were spread across eight lists. Each target list was combined
with each list from the preference-related controls, thus yielding to
a total of forty lists. The thirty unrelated filler items were
included in each of these resulting lists.

\subsection{Participants}
\label{sec:participants}

Forty native speakers of German took part in our study, none of which had
any prior exposure to logic or formal semantics. We excluded three
subjects due to insufficient performance on controls ($\le 50\%$
correct answers).

\begin{itemize}
\item possibly mention ages, backgrounds and sexes?
\end{itemize}

\subsection{Results}
\label{sec:results}

In the following, we will report the results for targets and
preference-related controls separately. After providing the
descriptive results, we present statistical analyses, which were
carried out in the following manner.  We computed log-linear models
(see SCHEEPERS 2003 ????) in two variants (by item and by participant). In
the first, the factor {\it Item} was included, and in the second
analysis, the factor {\it Participants} was included. Inclusion of
these two factors allowed us to test whether the distribution of
participants' choices was identical accros items and participants. We
report log-likelihood ratio Chi-squares ($LRCS_1$ and $LRCS_2$),
degrees of freedom ($df_1$, $df_2$) and significance levels ($p_1$ and
$p_2$). The relevant experimental factors are provided in the specific
subsections below.

\paragraph{Target conditions and positional controls}
Performance on positional controls was high (92\% correct responses on
average), indicating that subjects understood the task, giving
truth-value judgements at the right moment in a sequence without being
influenced in general by inessential features of the picture material
or showing general response biases.

The judgments obtained for the four target conditions are depicted in
Figure \ref{fig:JudgmentsK2}.\dn{we should introduce the actual
  sequences that we presented so that it is clear what it means to say
\emph{true} or \emph{false} on some position or other}
%
\begin{figure}[]
\centering
\subfloat[][\as-neutral]{
% \includegraphics[width=5cm]{../pictures/paper/graph_AE_NTR.jpg}
\begin{tikzpicture}[scale=0.8]

      \begin{axis}[ybar, 
                  enlarge x limits=0.1,
                  enlarge y limits=0.1,
                  legend style={at={(0.3,0.95)}, anchor=north, legend columns=-1},
                  ylabel={\% of answers (rounded)}, 
                  % xlabel={position}, 
                  symbolic x coords={1, 2, 3, 4, 5, 6, 7}, 
                  xtick=data, 
                  nodes near coords,
                  nodes near coords align={vertical}, 
                  x=20, y=1,
                  ymin = 0, ymax = 100,
                  bar width=2mm
        ]

        \addplot coordinates {(1,0) (2,0) (3,72) (4,1) (5,1) (6,0) (7,3)};

        \addplot coordinates {(1,0) (2,0) (3,0) (4,0) (5,0) (6,0) (7,23)};

        \legend{yes, no}

      \end{axis}
\end{tikzpicture}
\label{fig:ReadingsAS-ntr}
}
\subfloat[][\es-neutral]{
% \includegraphics[width=5cm]{../pictures/paper/graph_GE_NTR.jpg}
\begin{tikzpicture}[scale=0.8]

      \begin{axis}[ybar, 
                  enlarge x limits=0.2,
                  enlarge y limits=0.1,
                  legend style={at={(0.4,0.95)}, anchor=north, legend columns=-1},
                  % ylabel={\% of answers (rounded)}, 
                  % xlabel={position}, 
                  symbolic x coords={1, 2, 3, 4, 5}, 
                  xtick=data, 
                  nodes near coords,
                  nodes near coords align={vertical}, 
                  x=20, y=1,
                  ymin = 0, ymax = 100,
                  bar width=2mm
        ]

        \addplot coordinates {(1,1) (2,4) (3,3) (4,0) (5,30)};

        \addplot coordinates {(1,14) (2,0) (3,48) (4,2) (5,0)};

        \legend{yes, no}

      \end{axis}

\end{tikzpicture}
\label{fig:ReadingsES-ntr}
}

\subfloat[][\as-accented]{
% \includegraphics[width=5cm]{../pictures/paper/graph_AE_AKZ.jpg}
\begin{tikzpicture}[scale=0.8]

      \begin{axis}[ybar, 
                  enlarge x limits=0.1,
                  enlarge y limits=0.1,
                  legend style={at={(0.3,0.95)}, anchor=north, legend columns=-1},
                  ylabel={\% of answers (rounded)}, 
                  % xlabel={position}, 
                  symbolic x coords={1, 2, 3, 4, 5, 6, 7}, 
                  xtick=data, 
                  nodes near coords,
                  nodes near coords align={vertical}, 
                  x=20, y=1,
                  ymin = 0, ymax = 100,
                  bar width=2mm
        ]

        \addplot coordinates {(1,0) (2,1) (3,69) (4,1) (5,2) (6,0) (7,4)};

        \addplot coordinates {(1,0) (2,0) (3,1) (4,0) (5,0) (6,0) (7,23)};

        \legend{yes, no}

      \end{axis}

\end{tikzpicture}
\label{fig:ReadingsAS-acc}
}
\subfloat[][\es-accented]{
% \includegraphics[width=5cm]{../pictures/paper/graph_GE_AKZ.jpg}
\begin{tikzpicture}[scale=0.8]

      \begin{axis}[ybar, 
                  enlarge x limits=0.2,
                  enlarge y limits=0.1,
                  legend style={at={(0.4,0.95)}, anchor=north, legend columns=-1},
                  % ylabel={\% of answers (rounded)}, 
                  % xlabel={position}, 
                  symbolic x coords={1, 2, 3, 4, 5}, 
                  xtick=data, 
                  nodes near coords,
                  nodes near coords align={vertical}, 
                  x=20, y=1,
                  ymin = 0, ymax = 100,
                  bar width=2mm
        ]

        \addplot coordinates {(1,0) (2,11) (3,9) (4,1) (5,23)};

        \addplot coordinates {(1,0) (2,1) (3,50) (4,4) (5,1)};

        \legend{yes, no}

      \end{axis}

\end{tikzpicture}
\label{fig:ReadingsES-acc}
}

\caption{Positional judgments for the four target conditions. The
  plots show rounded percentages of \emph{yes}/\emph{no} judgments
  against the position within the trial. In the \as-conditions there
  were seven positions. On the third position a \emph{yes} judgment
  corresponded to a literal reading, while on the fifth position a
  \emph{yes} judgment corresponded to a global reading.  A \emph{no}
  judgment on the seventh position was only consistent with a local
  reading. The \es-conditions had five positions. \emph{No} judgments
  on positions two and three corresponded to global and literal
  readings, respectively. A \emph{yes} judgment on the last segment
  corresponded to a local reading.}
\label{fig:JudgmentsK2}
\end{figure}
%
We coded the judgments as {\it literal}, {\it global} or {\it local}
if they were as expected under one of these readings and as {\it
  error} if not. The distribution of readings thus coded is presented
in Figure~\ref{fig:JudgementPercentages}.
%
\begin{figure}[]
\centering
\subfloat[][\as-condition]{
\begin{tikzpicture}[scale=0.8]

      \begin{axis}[ybar, 
                  enlarge x limits=0.5,
                  enlarge y limits=0.25,
                  legend style={at={(0.5,-0.15)}, anchor=north, legend columns=-1},
        ylabel={\% of answers (rounded)}, symbolic x
        coords={neutral,accented}, xtick=data, nodes
        near coords, nodes near coords align={vertical},x=65, bar
        width=3mm, ymin = 8, ymax = 73 ]

        \addplot coordinates {(neutral,73) (accented,70)};

        \addplot coordinates {(neutral,23) (accented,22)};

        \addplot coordinates {(neutral,1) (accented,2)};

        \addplot coordinates {(neutral,4) (accented,6)};

        % \addplot coordinates {(ae-ntr,83) (ae-acc,80) (ge-ntr,53)
        % (ge-acc,56)};

        % \addplot coordinates {(ae-ntr,26) (ae-acc,25) (ge-ntr,33)
        % (ge-acc,26)};

        % \addplot coordinates {(ae-ntr,1) (ae-acc,2) (ge-ntr,0)
        % (ge-acc,1)};

        % \addplot coordinates {(ae-ntr,4) (ae-acc,7) (ge-ntr,28)
        % (ge-acc,31)};

        \legend{literal, local, global, error}

      \end{axis}

    \end{tikzpicture}

% \includegraphics[width=5cm]{../pictures/paper/ReadingsAE.pdf}
\label{fig:JudgementPercentagesAE}
}
\subfloat[][\es-condition]{
\begin{tikzpicture}[scale=0.8]

  \begin{axis}[ybar, 
               enlarge x limits=0.5,
               enlarge y limits=0.25, 
               legend style={at={(0.5,-0.15)},
                 anchor=north,
                 legend columns=-1}, 
               % ylabel={\% of answers}, 
               symbolic x coords={neutral,accented}, 
               xtick=data,
               nodes near coords,
               nodes near coords align={vertical},
               x=65, bar width=3mm,
               ymin = 8, ymax = 73
               ]

    \addplot coordinates {(neutral,47) (accented,49)};

    \addplot coordinates {(neutral,29) (accented,23)};

    \addplot coordinates {(neutral,0) (accented,1)};

    \addplot coordinates {(neutral,25) (accented,27)};

    % \addplot coordinates {(ae-ntr,83) (ae-acc,80) (ge-ntr,53) (ge-acc,56)};

    % \addplot coordinates {(ae-ntr,26) (ae-acc,25) (ge-ntr,33) (ge-acc,26)};

    % \addplot coordinates {(ae-ntr,1) (ae-acc,2) (ge-ntr,0) (ge-acc,1)};

    % \addplot coordinates {(ae-ntr,4) (ae-acc,7) (ge-ntr,28) (ge-acc,31)};

    \legend{literal, local, global, error}

  \end{axis}

\end{tikzpicture}

% \includegraphics[width=5cm]{../pictures/paper/ReadingsGE.pdf}
\label{fig:JudgementPercentagesGE}
}
\caption[Optional caption for list of figures]{Response type
  frequencies for target conditions}
\label{fig:JudgementPercentages}
\end{figure}
%
Participants mostly gave judgments indicating literal or local
readings. Judgments compatible with a global reading were seldomly
obtained. In the \as-conditions (see Figure~\ref{fig:JudgementPercentagesAE}),
judgments were consistent with literal readings in 65.0\% of the
trials with neutral prosody and in 67.5\% of the trials with accented
prosody.\dn{IMPORTANT: Figure contains different numbers (even if
  rounded)! Fabian, did you exclude the 3 bad performers when you
  obtained this data?} Judgments indicating local readings were given 
  in 22.0\% of both the neutral and accented trials. Global readings 
occurred in 1.6\% of the neutral and 0.8\% of the accented trials
(corresponding to a total of 3 answers across all trials).  
In the \es-conditions (see Figure~\ref{fig:JudgementPercentagesGE}) 
judgments consistent with literal readings were given in 47.2\% and 
44.7\% of the trials with neutral and accented prosody, respectively. 
Judgments indicating local readings were given in 22.8\% of
the neutral and 27.6\% of the accented trials. Global readings were
observed in 0.8\% of the trials in both the neutral and the accented
condition. 
As for the number of erroneous decisions, a discrepancy between 
\as-conditions and \es-conditions was observed: In the former, 
participants made hardly any errors, while in the latter, the amount 
of errors was considerably larger.

In order to test wether local readings occurred reliably, we tested whether the
number of local responses in each condition was higher than expected
by chance. Local responses had to be given on the last position in
each trial. If local readings didn't exist we would expect
participants, who reached the last position (ie. who did not abort the
trial prior to the last position) to give local judgments not more
often than expected by chance. On the last position two responses were
possible. One was compatible with a local reading the other was
not. Therefore, we would expect 50\% local responses by chance. As it
turns out local responses were given significantly more often than
50\% in all target conditions (all Bonferoni corrected p<.05). Note
that this finding cannot be explained by some kind of general response
bias on the last position since local readings required a \emph{yes} judgment
in the \es-conditions but a \emph{no} judgment in the \as-conditions.

In order to test whether accentuation or the quantifier had an
influence on the distribution of readings, log-linear models 
including the factors {\it Reading}, {\it Accentuation} and 
{\it Quantifier}.

Unsurprisingly, there was an effect of \emph{Reading} ($LRCS_1=364.77,
df_1 = 3, p_1<.001, LRCS_2=364.77, df_2 = 3, p_1<.001,$) because
overall the readings were distributed inhomogeneously (see Figure
\ref{fig:ReadingsGE}). The quantifier had an influence on the
distribution of readings as revealed by a reliable interaction of {\it
  Reading} and {\it Quantifier} ($LRCS_1=49.32, p_1<.01,LRCS_2=32.16,
p_1<.01,$). Looking at the data it seemed unlikely that the quantifier
affected the amount of local readings. We suspected that this
interaction was due to the higher number of errors and lower number of
literal readings in the \es-conditions as compared with the
\as-conditions. A difference in the distribution of judgment types
between participants was revealed by a reliable interaction of {\it
  Reading} and {\it Participant} ($LRCS_1=487.70, p_1<.01$). We were
interested in whether this effect was due to some of the participants
being more likely to choose literal readings than others. Finally,
there was a three-way interaction of {\it Participants}, {\it
  Construction} and {\it Reading} ($LRCS_1, = 143.92, df_1 = 117,
p_1<.05$) which could stem from the fact that some participants were
more prone than others to make errors in the \es-conditions.

In order to find out whether the interactions just reported affected
the amount of local readings, the log-linear models were computed
again with a different coding of the readings. Here, we only
considered the amount of local readings versus all other kinds of
judgments. Judgments were coded as {\it local} and {\it other}. If the
reported interactions affect the amount of local readings they should
show up again. The expected but irrelevant effect of {\it Reading} was
again significant ($LRCS_1=134.55, df_1 = 1, p_1<.001, LRCS_2=144.64,
df_2 = 1, p_1<.001,$). In addition, the interaction of {\it
  Participant} and {\it Reading} ($LRCS_1=487.70, p_1<.01$) as well as
the three-way interaction of {\it Construction}, {\it Participant} and
{\it Reading} ($LRCS_1=487.70, p_1<.01$) were significant. No other
effects were significant. In particular, the interaction of {\it
  Construction} and {\it Reading} was not significant ($LRCS_1=2.10,
p_1=.15,LRCS_2=.74, p_2=.39$).

The interaction of {\it Participant} and {\it Reading} shows that
there are indeed varying preferences for local readings among German
speakers (see Table~\ref{table:answer-types-as}).
%
\begin{table}[t]
  \centering
      \begin{tabular}{lccc}
      & \multicolumn{3}{c}{\es} \\ \cmidrule(r){2-4}
      \as & literalist & localist & incons. \\ \midrule
      literalist   & 19 & 2 & 6\\
      localist     &  0 & 7 & 0 \\
      incons. &  2 & 1 & 1\\
    \end{tabular}
    \caption{Contingency table showing the number of occurrences of
      \emph{response types} in the \as- and \es-conditions with
      neutral intonation. \emph{Literalists} are subjects who gave
      responses that were classified as either a literal response or
      an error; \emph{localist} are subjects who gave responses
      that were classified as either a 
      local response or an error. Subjects who gave at least two
      responses from different categories (literal, global, local) are
      classified as \emph{inconsistent}. (No subject's responses where
      errors exclusively and there where no \emph{globalists}, as
      defined in parallel fashion.)}
    \label{table:answer-types-as}
\end{table}
%
Further examination of the distribution of local readings revealed a
clear pattern (see Figure \ref{fig:HistogramLocalReadingsK2}). The
relative frequency of local judgments per participant strongly
correlated between all four conditions (all $r>.6, p<.001$). Appart
from the localists about half of the participants (and 41.5\% of all
participants) were consistently giving judgments that indicate literal
readings. Only a minority of participants exhibited
inconsistency. Reading preferences were more clear-cut in the
\as-conditions than in the \es-conditions. The number of consistent
literalists was lower in the \es-conditions (25.6\%) as compared to
the \as-conditions (57.3\%). Also, the number of participants that
showed inconsistency was higher in these conditions. However, the
number of consistent localists did hardly differ between
constructions.

\begin{figure}[h]
\centering
\subfloat[][\as-accented]{
\includegraphics[width=5cm]{../pictures/paper/histLocalReadingAE_AKZ.jpg}
\label{fig:ReadingsAE}
}
\subfloat[][\as-neutral]{
\includegraphics[width=5cm]{../pictures/paper/histLocalReadingAE_NTR.jpg}
\label{fig:ReadingsGE}
}

\subfloat[][\es-accented]{
\includegraphics[width=5cm]{../pictures/paper/histLocalReadingGE_AKZ.jpg}
\label{fig:ReadingsAE}
}
\subfloat[][\es-neutral]{
\includegraphics[width=5cm]{../pictures/paper/histLocalReadingGE_NTR.jpg}
\label{fig:ReadingsGE}
}
\label{fig:HistogramLocalReadingsK2}
\caption[]{Judgments for the four target conditions}
\end{figure}

The three way-interaction of {\it Participant}, {\it Construction} and
{\it Reading} is due to the fact that for some participants
preferences for local over other readings deviated between the two
construction types. These deviations were, however, not systematic to
any degree. How much local preferences deviated between the two
construction types per participant is depicted in Figure \ref{}. Most
participants had exactly the same preferences in the two
constructions. However, a few had a stronger and a few others a weaker
preference for local readings in the \es- than in the
\as-conditions. Since we did not find any systematic patterns
here, we speculate that the three-way interaction is due to the
\es-construction beeing understood non-standardly by a
few participants. This speculation is plausible given the higher
number of errors in the \es- as compared to the \as-conditions.

The absence of the interaction between {\it Construction} and {\it
  Participant} indicates that the type of construction does not affect
the amount of local readings. We assumed, based on the observed
distribution of judgments, that the type of contsruction affected the
amount of literal but not local and global readings. To further test
this assumption, we also considered literal and global readings versus
other judgments in separate analyses. Comparing global\dn{do you mean LITERAL?} readings versus
other judgments {\it Construction} and {\it Reading} were found to
interact ($LRCS_1=42.80, df= 1, p<.001; LRCS_2=22.14, df= 1,
p<.001$). Comparing global readings to other judgments no such
interaction was obeserved ($LRCS_1=1.24, df= 1, p=.29, LRCS_1=.21, df=
1, p=.65$).

\paragraph{Preference-related controls}

Participants' judgments are depicted in Figures \ref{Fig:EC_judgments}
and \ref{Fig:LC_judgments}. Judgments were coded as \ec or as \lc if
they were as expected under one of these readings, and as "other" if
they were not. The distributions of readings thus coded is illustrated
in Figure \ref{fig:response-type-freque-TRF}.  Overall, participants
mainly gave answers corresponding to the \lc-reading (65\%).
Moreover, this pattern was constrained by prosody. When prosody
supported a dispreferred \ec-reading, the number of \ec-decisions were
increased and vice versa.

In both \lc- and \ec-sequences, the number of \ec-answers was highest
when this reading was supported by prosodic phrasing (30\% and
32\%, respectively), and a bit lower with neutral prosody (24\% and
25\%). When prosodic phrasing supported the \lc-reading, the number of
\ec choices was lowest (12\% in both sequences). The number of
erroneous answers did not differ between the single
conditions. Finally, the availability of readings within a sequence
did not yield any significant effects.

In order to test statistically whether prosodic phrasing and
sequential ordering had an effect on the number of \ec-choices, we
contrasted the number of \ec-choices with non-\ec -choices (i.e.,
\lc-answers and erroneous responses together), calculating log-linear
model analyses with the factors {\it Reading}, {\it Prosody}, and {\it
  Sequence}.

An effect of {\it Reading} reflects that overall, readings were distributed inhomogeneously
($LRCS_1 = 252.8, df_1 =1, p_1 < .001$; $LRCS_2 = 252.8, df_2 =1, p_2 < 001$). 
As revealed by an interaction between {\it Prosody} and
{\it Reading}, prosodic phrasing had an influence upon the distribution of readings 
($LRCS_1 = 38.4, df_1 =2, p_1 < 001$; $LRCS_2 = 26.1, df_2 =2, p_2 < .001$).

In the analysis by participant, an interaction between {\it Participant} and {\it Reading}
indicates that reading preferences differed between participants ($LRCS_1 = 376.6,df_1 =37,  
p_1 < .001$).

In the analysis by items, a three-way interaction between {\it Item}, {\it Prosody} and 
{\it Reading} was observed ($LRCS_2 = 88.3, df_1 =54, p_2 < .01$). As this effect was 
unexpected, we carried out additional descriptive analyses (crosstables) in order to examine whether
the interaction might have been caused by specific items. When excluding the four items with
the highest deviation from expected frequencies (Residuals > 2.4 in at least one cell), 
the interaction was no longer significant ($p_2 = .18$).

\begin{figure}
\includegraphics[width=\textwidth]{../pictures/paper/graph_ec.pdf}
\caption{Judgments for prefence-related controls (EC sequence). Relative frequencies
of {\it yes/no} judgments is plotted against the position within the
trial. Each {\it EC sequences} consisted of five pictures, in which the
critical positions for a reading were the third position ("false" corresponded
to an \ec reading) and the fifth position ("true" corresponded to an \lc reading)}
\label{Fig:EC_judgments}
\end{figure}

%\clearpage

\begin{figure}
\includegraphics[width=\textwidth]{../pictures/paper/graph_lc.pdf}
\caption{Judgments for prefence-related controls (LC sequence). Relative frequencies
of {\it yes/no} judgments is plotted against the position within the
trial. Each {\it LC sequences} consisted of seven pictures, in which the
critical positions for a reading were the third position ("true" corresponded
to an \lc reading) and the sixth position ("false" corresponded to an \ec reading)}
\label{Fig:LC_judgments}.
\end{figure}

% \begin{figure}
% \includegraphics[width=\textwidth]{../pictures/graph_lc.pdf}
% \end{figure}

%\begin{itemize}
%\item see Figure~\ref{fig:response-type-freque-TRF}
%\end{itemize}

\begin{figure}[t]
  \centering
  \subfloat[][\LC-condition]{
        \begin{tikzpicture}[scale=0.8]

  \begin{axis}[ybar, 
               enlarge x limits=0.5,
               enlarge y limits=0.25, 
               legend style={at={(0.5,-0.15)},
                 anchor=north,
                 legend columns=-1}, 
               ylabel={\% of answers (rounded)}, 
               symbolic x coords={neutral,LC-pref,EC-pref}, 
               xtick=data,
               nodes near coords,
               nodes near coords align={vertical},
               x = 45, 
               bar width = 3mm,
               ymin = 8
               ]

    \addplot coordinates {(neutral, 69) (LC-pref, 77) (EC-pref, 52)};

    \addplot coordinates {(neutral, 24) (LC-pref, 12) (EC-pref, 30)};

    \addplot coordinates {(neutral, 7) (LC-pref, 11) (EC-pref, 19)};

    % \addplot coordinates {(l-ntr, 93) (l-l, 101) (l-e, 70) (e-ntr, 85)
    %   (e-l, 100) (e-e, 75)};

    % \addplot coordinates {(l-ntr, 33) (l-l, 16) (l-e, 40) (e-ntr, 34)
    %   (e-l, 16) (e-e, 43)};

    % \addplot coordinates {(l-ntr, 9) (l-l, 14) (l-e, 25) (e-ntr, 16)
    %   (e-l, 15) (e-e, 17)};

    \legend{LC, EC, error}

  \end{axis}

\end{tikzpicture}
  }
  \subfloat[][\ec-condition]{
    \begin{tikzpicture}[scale=0.8]

  \begin{axis}[ybar, 
               enlarge x limits=0.5, 
               enlarge y limits=0.25, 
               legend style={at={(0.5,-0.15)},
                 anchor=north,
                 legend columns=-1}, 
               % ylabel={\% of answers}, 
               symbolic x coords={neutral,LC-pref,EC-pref}, 
               xtick=data,
               nodes near coords,
               nodes near coords align={vertical},
               x = 45, 
               bar width = 3mm,
               ymin = 8
               ]

    \addplot coordinates {(neutral, 63)
      (LC-pref, 76) (EC-pref, 56)};

    \addplot coordinates {(neutral, 25)
      (LC-pref, 12) (EC-pref, 32)};

    \addplot coordinates {(neutral, 12)
      (LC-pref, 11) (EC-pref, 13)};

    % \addplot coordinates {(l-ntr, 93) (l-l, 101) (l-e, 70) (e-ntr, 85)
    %   (e-l, 100) (e-e, 75)};

    % \addplot coordinates {(l-ntr, 33) (l-l, 16) (l-e, 40) (e-ntr, 34)
    %   (e-l, 16) (e-e, 43)};

    % \addplot coordinates {(l-ntr, 9) (l-l, 14) (l-e, 25) (e-ntr, 16)
    %   (e-l, 15) (e-e, 17)};

    \legend{LC, EC, error}

  \end{axis}

\end{tikzpicture}
}
  \caption{Response type frequences for preference-related controls}
  \label{fig:response-type-freque-TRF}
\end{figure}

\subsection{Summary \& Interpretation}
\label{sec:interpretation}

High scores of success clearly indicate that participants understood 
the incremental verification task and were in general able to give 
judgements at the earliest possible point in a picture sequence. 
Moreover, results obtained for the preference-related controls  
show that our design is able to detect multiple readings and preferences 
among these. Crucially, the amount of response types for \lc- and \ec-readings 
reflected the known preference for the former, \emph{irrespective} of the 
order in which each reading could first be judged in a given sequence.
Finally, the incremental verification task was generally sensitive
to prosody. However, prosodic effects were restricted to the preference-related 
controls. As expected, in these conditions, the dispreferred \ec-reading was chosen 
significantly more often when supported by prosodic phrasing. Interestingly, 
contrastively accenting scalar \emph{some} had no effect on the observed 
decisions in the target conditions. Crucially, local readings occurred in both
\as- and \es-sentences, independently of prosody. Finally, closer examination of 
the participants' decisions revealed two groups: participants in the first
group consistently gave answers corresponding to local readings, whereas participants
in the second group consistently showed a literal interpretation.

Given the overall pattern of results, the most straightforward 
explanation of the observed data is the hypothesis that the actual 
preferences among readings for both \as- and \es-sentences alike are:

\begin{exe}
  \ex \label{bsp:observed-preferences} \lit \ > \ \loc \ \ ( > \glb)
\end{exe}

\noindent where it is unclear whether the global readings are entirely
unavailable or simply severely dispreferred. Crucially, our data
attests local readings independent of prosody, albeit that local
readings are not preferred to literal readings and given, as it seems,
only by a small, but substantial number of consistent localists.

These findings are hard to explain both by traditionalism and grammaticalism, 
as no version of either would predict the pattern in (\ref{bsp:observed-preferences}). 
Traditionalism, if left on its own, fails to predict local readings,
while grammaticalism, when it relies on the strongest meaning
hypothesis, fails to predict the right preference order. Let's look
more closely at each position in turn.

For traditionalism, the absence of global readings \textit{per se} can
be explained, for instance, by proposing that the extra assumptions
needed to derive global readings are not standardly available (as in
the weak variety of traditionalism).  Alternatively, traditionalists
could argue that the task's rather high processing demands lead to a
suppression of global readings, in line with the findings of
\citet{NeysDe-NeysSchaeken2007:When-People-Are} that working memory
load negatively affects the number of scalar implicature responses in
sentence verification tasks. However, it then remains unclear why
memory effects should be restricted to global readings.  As a result,
approaches like the unrestricted variety of traditionalism thereby
loose the possibility of accounting for local readings of
\as-sentences.\fn{Remember that unrestricted traditionalism, as we
  called it, can account for local readings of \as-sentences by
  assuming that a sentence of the form \emph{Some of the $X$ are
    related to some of the $Y$} is a feasible alternative.} This is
because unrestricted traditionalism explains local and global readings
of \as-sentences in essentially the same manner. On top of that,
traditionalism cannot explain local readings for
\es-sentences. Resorting to the intonational markedness hypothesis
does not help improve predictions of traditionalism. The latter would
predict an increase in local readings in the accented conditions, and
maybe even an absence of local readings in the unaccented
conditions. Neither was observed. Local readings were observed in the
accented as well as in the unaccented condition, thus countering
previous traditionalist suggestions that local readings are to be
attributed to specific pragmatic mechanisms. In sum, traditionalism
alone cannot explain (\ref{bsp:observed-preferences}), but needs an
additional explanation of local readings for \emph{both} \as-
\emph{and} \es-sentences.

Grammaticalism, on the other hand, fails to predict the preference
relation in (\ref{bsp:observed-preferences}), because it strictly
contradicts the strongest meaning hypothesis in either of the two
varieties suggested by \citet{ChierchiaFox2008:The-Grammatical}. This
shows most clearly in the fact that both versions of the strongest
meaning hypothesis predict that local readings of \as-sentences are
preferred over literal ones. The problem, therefore, seems to be the
selection of readings by the strongest meaning hypothesis. In fact, if
(\ref{bsp:observed-preferences}) is the actual preference over
readings, then the strongest meaning hypothesis is also conceptually
on the wrong track, because no account that merely looks at the
logical strengths of readings can plausibly predict the preference
pattern in (\ref{bsp:observed-preferences}) for both \as- and
\es-sentences at the same time. (Try, if you please!)

Summing up so far, the existing versions of traditionalism and
grammaticalism cannot explain the present results. Of the two
approaches, grammaticalism seems better suited to account for the
observed effects. Whereas traditionalism cannot cope with local
readings (of unaccented sentences), the only thing that grammaticalism
would have to change is the selection criterion over preferred
readings, i.e., to reconsider the strongest meaning
hypothesis. Turning away from comparisons of logical strength, a
\emph{prima facie} plausible selection criterion that produces
(\ref{bsp:observed-preferences}) is ready at hand: if we assume that
each insertion of the $\exh(\cdot)$ operator is \emph{costly}
anti-proportional to the depth of its embedding under other logical
operators, then we derive a preference of literal readings over local
ones over global ones. Whether such an \emph{economy principle} is
sound and in line with further empirical data would then be an
interesting matter of further investigation.

Interestingly, an alternative theory of scalar implicature explains
the observed preference ordering (\ref{bsp:observed-preferences}) in a
completely straightforward manner.  \emph{Conventionalism} (also
sometimes called \emph{lexicalism}) proposes that if the
upper-bounding inference from \emph{some} to \emph{some but not all}
occurs frequently, it would be inefficient to execute over and over
again the Gricean reasoning that traditionalists propose
\citep[e.g.][]{LevinsonPresumptiveMeanings2000,Chierchia:2004_ScalarImplicatures}. Conventionalism
therefore assumes that scalar inferences have become routinized,
hard-wired, so much so that there are two lexical entries for a scalar
item like \emph{some}, one with logical meaning \emph{some and maybe
  all} and one with the upper-bounded meaning \emph{some but not
  all}. \citet{LevinsonPresumptiveMeanings2000} proposed that the
latter is the preferred default meaning, a position which has
therefore ofttimes be addressed as \mymark{defaultism}. Unfortunately,
defaultism has strong empirical evidence against it, because several
processing studies show that the computation of scalar inferences
appears to be time- and effort-consuming, when they do arise,
not when they do not
\citep[c.f.][]{BrehenyKatsos2006:Are-Generalised,BrehenyKatsos2008:Experimental-In,NeysDe-NeysSchaeken2007:When-People-Are}. But
there is also competing evidence that scalar inferences are available
in immediate, low-level processes, such as the resolution of
referential expressions in a visual-world paradigm
\citep{GrodnerKlein2010:Some-and-Possib}. For our purposes, it is
interesting to note that the preference relation in
(\ref{bsp:observed-preferences}) is predicted by the dual of
defaultism, call it \emph{literalism}, which posits a lexical
ambiguity but predicts that the logical meanings are \emph{ceteris
  paribus} preferred over the upper-bounded meanings.

Unfortunately, there are problems with conventionalism, no matter
which lexical reading is assumed to be preferred, that seem to require
that it be backed up by traditionalism
\citep[e.g.][]{Sauerland2012:The-Computation}.\fn{\citet{Sauerland2012:The-Computation}
  eventually dismisses even conventionalism backed up by
  traditionalism in favor of grammaticalism based on cases of
  so-called \emph{intermediate implicatures}, where what seems to be
  necessary is the insertion of an exhaustivity operator in
  intermediate scope position \citep[see][for
  details]{Sauerland2012:The-Computation}. Appreciation of such
  intermediate implicatures requires more argumentation, involving a
  discussion of \emph{Hurford's constraint}
  \citep[c.f.][]{FoxSpector:Economy-and-Emb,ChierchiaFox2009:Hurfords-Constr,ChierchiaFox2008:The-Grammatical},
so we skip this here.}
For instance, conventionalism cannot account for \emph{indirect
  implicatures}, such as the inference from
(\ref{bsp:indirect-SI-Target}) to (\ref{bsp:indirect-SI-Inference})
which cannot be explained by positing a lexical ambiguity for
\emph{all}.

\begin{exe}
  \ex \label{bsp:indirect-SI}
    \begin{xlist}
      \ex \label{bsp:indirect-SI-Target} Hans did not solve all of the problems.
      \ex \label{bsp:indirect-SI-Inference} Hans solved some of the problems.
    \end{xlist}
\end{exe}

\noindent In the context of the present study, this would then be good
news for traditionalism, because a combined
traditionalism+conventionalism would predict the observed data
well. The combined position would even be able to explain something
that grammaticalism cannot make sense of as easily, namely the
observation that there were rather consistent localists among the
participants of our study. The combined tradionalism+conventionalism
could explain this by the assumption that some (lazy?, smart?)
speakers have made a conventionalized upper-bounding meaning entry for
scalar \emph{some} while others keep computing things on the fly. It
is not clear how grammaticalism can account for the intra-subjective
consistency. Still, this is not an argument \emph{for} traditionalism,
because the point remains that traditionalism as such fails to predict
the observed data.

In sum, if the interpretation of our data as suggested here is
correct, then both traditionalism and grammaticalism are in need of
revision. Traditionalism must explain the availability of local
readings, grammaticalism must revise its selection criterion among
readings. In the remainder of this section we would like to discuss
whether this interpretation of our data is feasible.

\subsection{General Discussion}
\label{sec:discussion}

%Several worries and objections are conceivable that might cast doubt
%on the previous interpretation of our experimental data.

In the following, we would like to address potential alternative
explanations for the present results. We will argue that besides
these interpretation variants, the main point raised
above remains valid: neither traditionalism, nor grammaticalism
is able to account for the present pattern of results.

%some of these that we consider most
%pressing. Eventually, however, we conclude that the main point raised
%above remains valid: neither traditionalism, nor grammaticalism
%explains the observations even under alternative interpretations of the
%data.

First, it could be questioned whether our design reliably 
indicates the preference relation in (\ref{bsp:observed-preferences}) 
among readings of our target sentences. Thus, the incremental verification task
(\ivt) might simply select for the weakest possible reading. This idea
can be grounded in the observation that the preferred \lc-readings
of our target-related controls entail the dispreferred
\ec-readings. It could thus be argued that the difference in responses
for preference-related controls are due to the entailment properties,
not the known preference relation over \lc- and \ec-readings. However, this
does not explain why the distribution of responses in the target
conditions are obviously not governed by logical strength. In fact, it
is a very noteworthy point that although entailment relations among
readings differ between \as- and \es-sentences, the observed response
pattern does not. (This is what is at the heart of the problem for a
grammaticalism that relies on the strongest meaning hypothesis.)
\dn{Petra: ich glaub, den Punkt verstehe ich nicht ganz - sollte man hier
auch die Prosodie erwähnen? Phrasierung wirkt sich ja nicht auf entailment
relations aus, sondern wird normalerweise dem Syntax-Prosodie-Interface
zugeschrieben}

But maybe, a similar kind of objection goes, it is not feasible to
actually compare results from the preference-related controls to
results from our target conditions, because the kind of ``ambiguity''
is of a different kind. In the former we have a syntactic
ambiguity, in the latter a small number of competing pragmatic
enrichments. The former is a case of perceived ambiguity (some of our
subjects reported this in a \emph{Nachbefragungsbogen}\dn{English?}), 
the latter arguably is
not \citep[c.f.][]{GeurtsPouscoulous2009:Embedded-Implic}. So perhaps
even if the distribution of responses is indicative of the known
preference patterns in preference-related controls, this would not
necessarily mean that the distribution of responses in target
conditions is indicative of preference relations.

There are two kinds of reply to this objection (and others like
it). The first kind of reply starts with a concession. It is true that
we cannot conclude with certainty that our data implies that
(\ref{bsp:observed-preferences}) is the actual preference relation
among candidate readings of our target sentences. But we do not have
to either. Maybe in fact, the observed distribution has nothing to do
with preferences over readings. It does not matter. The main point is
that neither traditionalism nor grammaticalism predicts the observed
distribution of responses. The second kind of reply is fatalistic but
stubborn. If we assume that the \ivt is incapable of indicating
preferences among readings for pragmatically unspecified, then it
should at least be able to indicate availability of readings, should
it not? And if we deny the observed response patterns the power to
indicate preference, what should the information about availability of
readings be that we may safely infer from it? It should be that local
readings are available and that global readings are not. But that is
again bad news for both traditionalism and grammaticalism. The former
cannot explain the substantial number of local responses, the latter
doesn't explain the absence of global responses.

\begin{itemize}
\item pictures overemphasize contrast some/all
\item subjects tricked into local responses because they are at the end
  \begin{itemize}
  \item would only show availability after all
  \end{itemize}
\item typicality after all
\end{itemize}

\subsubsection{Bits and Pieces}
\label{sec:bits-pieces}

\begin{itemize}
\item The distribution of
responses found in previous studies could in part be explained by
differences in typicality of the pictures with respect to the possible
readings (see van Tiel). In the present case this kind of reasoning
would have to be extended to the typicalilty of parts of pictures. On
intuitive grounds, we take this kind of reasoning to be
implausible. We will now shortly discuss how graphical properties
might still affect judgments in the worst case. In the first step of
the \as-condition local and global readings cannot be judged,
yet. Therefore, graphical properties should not affect judgments if
participants have one of these readings in mind. If participants have
a literal reading in mind, the worst case scenario is that they
refrain from judging the sentence as true at this point due to some,
eg. atypical, graphical property of the picture. The situation is
similar at the second step. The local reading is still not
decidable. Therefore, graphical properties should not play a role if
participants have a local reading in mind. A positive judgment might,
however, be a delayed judgment of a literal reading. Again, the worst
case scenario for our purpose is that participants refrain from
answering because of graphical properties. At the last step the local
reading requires a negative judgment. Now, we could obtain delayed
negative responses from literal or global readings because the graph
as a whole is, e.g., logically true but atypical with respect to these
readings. These kinds of judgments would be indistinguishable from
local responses. Note, however, that we have no reason to believe that
typicality has large effects, because, given the results from van
Tiel, the picture as a whole is expected to be one of the more typical
situations for \as-sentences. Still, we cannot exclude the possibility
that graphical properties of the pictures might have the described
effect. In the \es-conditions this type of effect is, however,
entirely impossible since the literal and global reading is plainly
false given the initial parts as well as the whole graph. Summing up,
in the worst case, which we consider implausible, graphical effects
might lead to an enhanced number of local responses in the \as- as
compared to the \es-condition. We have to keep this possibility in
mind. We want to stress, however, that the design presented here
improves on previous studies with respect to possible graphical
effects.
\end{itemize}

\section{Conclusions}
\label{sec:conclusions}

\begin{itemize}
\item future research: isolate consistent localists to probe into \emph{processing}
\end{itemize}

\appendix

\section{The auditory sentence material}
\label{sec:audit-sent-mater}

\paragraph{Targets.} Table~\ref{tab:table-A} shows the durational values for each of
the single regions in the sentence. Crucially, durational values of
accented determiners were significantly increased as opposed to their
non-accented counterparts. Overall, \acro{Quantifier} effects were
observed, which might be attributed to the above-mentioned lexical
differences, as well as to the fact that \es-structures generally
contained fewer words, thus leading to a tendency of a durational
decrease for each of the single words.

Differences between maximal and minimal F0 values for each of the
single words in the sentence are depicted in
Table~\ref{tab:table-B}. As is descriptively evident, accented
determiners showed a larger F0 range compared to unaccented versions,
an effect which is confirmed by statistical analyses. An additional
\acro{Quantifier}*\acro{Prosody} interaction indicates that these
differences are even more pronounced in the \as-condition. Finally,
comparable to the durational analyses, \acro{Quantifier} effects were
observed when conditions exhibited lexical differences.

Finally, differences between maximal and minimal F0 values for each of
the single words in the sentence are given in
Table~\ref{tab:table-D}. Again, the most reliable differences occur at
the boundary regions.\dn{provide caption for table D}

\begin{table}
  \centering

  \includegraphics[width=\textwidth]{../pictures/Acoustics/Table-A.png}

  \caption{Durational values in ms for each of the single regions in the target sentences. Region 6 corresponds to \emph{einigen}.}
  \label{tab:table-A}
\end{table}

\begin{table}
  \centering

  \includegraphics[width=\textwidth]{../pictures/Acoustics/Table-B.png}

  \caption{Difference between minimal and maximal F0 values in Hz for
    each of the single words in the target sentences. Region 6
    corresponds to \emph{einigen}.} 
  \label{tab:table-B}
\end{table}

\paragraph{Preference-related fillers.} Table~\ref{tab:table-C} shows the
durational values of each of the single words in the sentence. As is
descriptively evident, the largest durational differences were
realized at the boundary regions (i.e. Region 5 and Region
7). Interestingly, small differences between conditions also yielded
significance at other positions, suggesting that our speaker produced
the different conditions very consistently (i.e., with little
variance).

\begin{table}
  \centering

  \includegraphics[width=\textwidth]{../pictures/Acoustics/Table-C.png}

  \caption{Durational values in ms for each of the single regions in
    the preference-related fillers. Regions 5 and 7 correspond to the
    nouns preceding the boundaries.}  
  \label{tab:table-C}
\end{table}

\begin{table}
  \centering

  \includegraphics[width=\textwidth]{../pictures/Acoustics/Table-D.png}

  \caption{???}  
  \label{tab:table-D}
\end{table}

\printbibliography[heading=bibintoc]

\end{document}