### 2013-06-20 text/micha_stat_analysis/notes.tex

\documentclass[fleqn,reqno,10pt]{article}

\usepackage{myarticlestyledefault}

\usepackage{subfig}

\title{K2 Analysis}
\author{Michael Franke}
\date{}

\newcommand{\lit}{\acro{lit}}
\newcommand{\glb}{\acro{glb}}
\newcommand{\loc}{\acro{loc}}

\newcommand{\as}{\acro{as}}
\renewcommand{\es}{\acro{es}}
\renewcommand{\AE}{\as}
\newcommand{\GE}{\es}
\newcommand{\ea}{\acro{ea}}

\newcommand{\lc}{\acro{lc}}
\newcommand{\ec}{\acro{ec}}
\newcommand{\LC}{\lc}
\newcommand{\EC}{\ec}

\begin{document}
\maketitle

\section{Controls}
\label{sec:controls}

\section{Target Conditions}
\label{sec:target-conditions}

Each participant saw each critical sentence type three times with
neutral and three times with constrastive stress intonation on the
scalar item. We coded judgements as either \emph{literal},
\emph{local}, \emph{global} or \emph{false}. Contingency tables of the
observed frequencies are in Table~XYZ.

\begin{table}[h]

\subfloat[][\as neutral]{\begin{tabular}{lcccc}
    & \emph{literal} & \emph{local} & \emph{global} & \emph{false} \\
    \midrule
    Pos 1 & 29 &  7 & 0 & 1 \\
    Pos 2 & 26 & 10 & 0 & 1 \\
    Pos 3 & 25 &  9 & 1 & 2 \\ \addlinespace[0.15cm]
    Total & 80 & 26 & 1 & 4
\end{tabular}}
%
\hspace{0.2cm}
%
\subfloat[][\es neutral]{\begin{tabular}{lcccc}
    & \emph{literal} & \emph{local} & \emph{global} & \emph{false} \\
    \midrule
    Pos 1 & 16 &  7 & 0 & 14 \\
    Pos 2 & 21 & 13 & 0 &  3 \\
    Pos 3 & 16 & 13 & 0 &  8 \\ \addlinespace[0.15cm]
    Total & 51 & 33 & 0 & 25
\end{tabular}}

\subfloat[][\as accented]{\begin{tabular}{lcccc}
    & \emph{literal} & \emph{local} & \emph{global} & \emph{false} \\
    \midrule
    Pos 1 & 28 &  7 & 1 & 1 \\
    Pos 2 & 25 &  8 & 1 & 3 \\
    Pos 3 & 24 & 10 & 0 & 3 \\ \addlinespace[0.15cm]
    Total & 77 & 25 & 2 & 7
\end{tabular}}
%
\hspace{0.2cm}
%
\subfloat[][\es accented]{\begin{tabular}{lcccc}
    & \emph{literal} & \emph{local} & \emph{global} & \emph{false} \\
    \midrule
    Pos 1 & 16 &  7 & 0 & 14 \\
    Pos 2 & 21 & 13 & 0 &  3 \\
    Pos 3 & 16 & 13 & 0 &  8 \\ \addlinespace[0.15cm]
    Total & 51 & 33 & 0 & 25
\end{tabular}}

\caption{Contingency table of raw judgements.}
\label{tab:CC-ContingencyTable}
\end{table}

\section{Preference-Related Controls}
\label{sec:pref-relat-contr}

\printbibliography[heading=bibintoc]

\end{document}