### 2014-09-09 text/paper/submission/cover_letter.tex
this version first submission to JoS

\documentclass[fleqn,reqno,10pt]{article}

\usepackage{applicationstyle}

\begin{document}

\thispagestyle{empty}

\begin{minipage}{7.5cm}
  The editors of\\
  Journal of Semantics
\end{minipage}\hfill
\begin{minipage}{5.3cm}
  Michael Franke \\
  Universtit\"at T\"ubingen \\
  Seminar f\"ur Sprachwissenschaft \\
  Wilhelmstra\ss e 19\\
  72072 T\"ubingen\\
  Germany  \\
  \medskip \href{mailto:mchfranke@gmail.com}{\tt
    mchfranke@gmail.com}
\end{minipage}

\vspace{0.75cm}

\hfill Emden, September 9, 2014

\vspace{0.75cm}

submission of manuscript ``Embedded Scalars, Preferred Readings \& Intonation: An
  Experimental Revisit''

\vspace{0.75cm}

Dear colleagues:

\bigskip

here is a first submission of the paper ``Embedded Scalars, Preferred
Readings \& Intonation: An Experimental Revisit'' authored by Michael
Franke, Fabian Schlotterbeck and Petra Augurzky, all from the
University of T\"ubingen.

Please note that the paper contains, next to a ``classical''
statistical analysis, also a Bayesian analysis. I mention this here
because it could easily be overlooked but it might be relevant to your
choice of reviewers.

Thank you for your efforts.

\vspace{0.75cm}

\hfill Best wishes,\hspace{2.5cm}

\medskip

\hfill \includegraphics{signature.jpg}\hspace{1.7cm}

\vspace{-.3cm}

% \hfill Michael Franke\hspace{2cm}

% \vfill
% \begin{description}
% \item[enclosed:] research accomplishments; future research; approach to
%   teaching; curriculum vitae; list of three referees
% \end{description}

\end{document}