### 2014-12-11 MT/exp_01-correlation-bysubject/pic.tex
rearranged MT experiment folder

\documentclass{article}

\usepackage[palatino]{mypackages}

\usepackage{mycommands}

\usepackage{verbatim}
\usepackage[active,tightpage]{preview}
\PreviewEnvironment{tikzpicture}
\setlength\PreviewBorder{5pt}

\usepackage{pgfplots}

\begin{document}

\begin{tikzpicture}[node distance = 2cm]

  \node[rectangle, fill = blue, draw=black, minimum size = 1cm] (s1)
  at (0,0) {};

  \node[rectangle, fill = blue, draw=black, minimum size = 1cm] (s2)
  at (0,3) {};

  \node[rectangle, fill = blue, draw=black, minimum size = 1cm] at (0,6) (s3) {};

  \node[circle, fill = yellow, draw=black, minimum size = 1cm] at (4,0) (c1) {};

  \node[circle, fill = yellow, draw=black, minimum size = 1cm] at (4,2) (c2) {};

  \node[circle, fill = yellow, draw=black, minimum size = 1cm] at (4,4) (c3) {};

  \node[circle, fill = yellow, draw=black, minimum size = 1cm] at (4,6) (c4) {};

  \draw[very thick,-] (s1) -- (c1);
  \draw[very thick,-] (s1) -- (c2);
  \draw[very thick,-] (s1) -- (c3);
  \draw[very thick,-] (s1) -- (c4);

  \draw[very thick,-] (s2) -- (c1);
  \draw[very thick,-] (s2) -- (c2);
  \draw[very thick,-] (s2) -- (c3);
  \draw[very thick,-] (s2) -- (c4);

  \draw[very thick,-] (s3) -- (c4);
  \draw[very thick,-] (s3) -- (c3);

\end{tikzpicture}

\end{document}