### 2014-12-11 MT/exp_02-K2-unembedded-postTest/pics/letter_triangles_dashedLines.tex
new experiment K2 Post

\documentclass{article}

\usepackage[palatino]{mypackages}

\usepackage{mycommands}

\usepackage[]{mathdesign}
%\usepackage{MnSymbol}
% \usepackage{ulsy}
\usepackage{wasysym}
\usepackage{ascii}
\usepackage{marvosym}
\usepackage{pifont}
\usepackage{phaistos}
\usepackage{epsdice}
\usepackage{skak}

\usepackage{verbatim}
\usepackage[active,tightpage]{preview}
\PreviewEnvironment{tikzpicture}
\setlength\PreviewBorder{20pt}

\usetikzlibrary{shapes.geometric}
\pgfdeclarelayer{foreground}
\pgfsetlayers{background,main,foreground}

\begin{document}
\begin{tikzpicture}[node distance = 3cm]

    \tikzset{ctype/.style={circle,thick,minimum
      width=18mm},
      htype/.style={},
      p6type/.style={thick,minimum width=20mm},
      p4type/.style={thick,minimum width=20mm},
      arrow/.style={-,very thick}}
\begin{pgfonlayer}{background}
      \node at (0,0) [circle,draw=black!70,fill=gray!20,minimum size = 8.5cm] {};
    \end{pgfonlayer}\begin{pgfonlayer}{foreground}

      \node at (0,0)  [ctype,draw=black!70,fill=brown!20] 
      (c0)   {\textcolor{black}{\Huge{$\text{\Letter}$}}};

    \end{pgfonlayer}
\begin{pgfonlayer}{foreground}

\node [p6type,regular polygon, regular polygon sides=3,draw=black!70,fill=blue!50, above left of = c0]  (p0)   {};

\node [p6type,regular polygon, regular polygon sides=3,draw=black!70,fill=blue!50, above of = c0]  (p1)   {};

\node [p6type,regular polygon, regular polygon sides=3,draw=black!70,fill=blue!50, above right of = c0]  (p2)   {};

\node [p6type,regular polygon, regular polygon sides=3,draw=black!70,fill=blue!50, below right of = c0]  (p3)   {};

\node [p6type,regular polygon, regular polygon sides=3,draw=black!70,fill=blue!50, below of = c0]  (p4)   {};

\node [p6type,regular polygon, regular polygon sides=3,draw=black!70,fill=blue!50, below left of = c0]  (p5)   {};

\draw [arrow, dashed] (c0) to (p3) node[midway, above] {\Large{?}};

\draw [arrow, dashed] (c0) to (p4) node[midway, left] {\Large{?}};

\draw [arrow, dashed] (c0) to (p5) node[midway, above] {\Large{?}};

\end{pgfonlayer}

\draw [arrow] (c0) to (p0);

\draw [arrow] (c0) to (p1);

\draw [arrow] (c0) to (p2);

                %  \begin{scope}

                %       \clip (-4.1,-4.1) rectangle (4.1,0);

                %       \node at (0,0) [circle,draw=gray!80,fill=gray!80,minimum size = 8.1cm] {};

                % \end{scope}

        \begin{pgfonlayer}{background}
      \node at (0,0) [circle,draw=black!70,fill=gray!20,minimum size = 8.5cm] {};
    \end{pgfonlayer}

\end{tikzpicture}

\end{document}