### 2014-12-19 MT/exp_03_MT-IVT-Small/pics/Var03_bell_triangle/0-0.tex

\documentclass{article}

\usepackage[palatino]{mypackages}\usepackage{mycommands}

\usepackage[]{mathdesign}
%\usepackage{MnSymbol}
\usepackage{wasysym}
\usepackage{ascii}
\usepackage{marvosym}
\usepackage{pifont}
\usepackage{phaistos}
\usepackage{epsdice}
\usepackage{skak}

\usepackage{verbatim}
\usepackage[active,tightpage]{preview}
\PreviewEnvironment{tikzpicture}
\setlength\PreviewBorder{20pt}

\usetikzlibrary{shapes.geometric}
\pgfdeclarelayer{foreground}
\pgfsetlayers{background,main,foreground}

\begin{document}
\begin{tikzpicture}[node distance = 3cm]

    \tikzset{ctype/.style={circle,thick,minimum
      width=18mm},
      htype/.style={},
      p6type/.style={thick,minimum width=14mm},
      p4type/.style={thick,minimum width=14mm},
      arrow/.style={-,very thick}}
\begin{pgfonlayer}{background}
      \node at (0,0) [circle,draw=black!70,fill=gray!20,minimum size = 8.5cm] {};
    \end{pgfonlayer}\begin{pgfonlayer}{foreground}

      \node at (0,0)  [ctype,draw=black!70,fill=brown!20] 
      (c0)   {\textcolor{black}{\Huge{$\text{\bell}$}}};

    \end{pgfonlayer}
\begin{pgfonlayer}{foreground}

\node [p6type,regular polygon, regular polygon sides=3,draw=black!70,fill=green!50, above left of = c0]  (p0)   {};

\node [p6type,regular polygon, regular polygon sides=3,draw=black!70,fill=green!50, above of = c0]  (p1)   {};

\node [p6type,regular polygon, regular polygon sides=3,draw=black!70,fill=green!50, above right of = c0]  (p2)   {};

\node [p6type,regular polygon, regular polygon sides=3,draw=black!70,fill=green!50, below right of = c0]  (p3)   {};

\node [p6type,regular polygon, regular polygon sides=3,draw=black!70,fill=green!50, below of = c0]  (p4)   {};

\node [p6type,regular polygon, regular polygon sides=3,draw=black!70,fill=green!50, below left of = c0]  (p5)   {};

\end{pgfonlayer}

                 \begin{scope}

                      \node at (0,0) [circle,draw=gray!80,fill=gray!80,minimum size = 8.1cm] {};

                \end{scope}

        \begin{pgfonlayer}{background}
      \node at (0,0) [circle,draw=black!70,fill=gray!20,minimum size =
      8.5cm] {};
      \node at (0,0) [rectangle,minimum size = 11cm] {};
    \end{pgfonlayer}\end{tikzpicture}

\end{document}