### 2015-02-03 text/paper/revision_01/paper.tex
K2 1st revision initial export

\documentclass[fleqn,reqno,10pt]{article}

\usepackage[natbib=true,
            style=authoryear-comp,
            backend=bibtex,
            sortcites=false,
            maxnames=2,
            doi=false,
            url=false]{biblatex}
\bibliography{MyRefGlobal}
% \bibliography{paper}

\usepackage[final,            % override "draft" which means "no nothing"
            colorlinks,       % rather than outlining them in boxes
            linkcolor=gray,   % override truly awful color choices
            citecolor=gray,   %   (ditto)
            urlcolor=gray,    %   (ditto)
            plainpages=false, % to overcome complaints with multiple
            pdfpagelabels,    % multiple page 1-s due to preface
            hypertexnames=false % solves warning, but interferes with
                                % index and \autoref apparently
            ]{hyperref}

\usepackage{amsmath}            % Formeln
\usepackage{amsfonts}           % Fonts for Formulas
\usepackage{amssymb}
\usepackage[final]{graphicx}
\usepackage{booktabs}
\usepackage{enumerate}
\usepackage{lipsum}
\usepackage{txfonts} % for strict implication symbols
\usepackage{soul}
\usepackage{relsize} % provides command \relsize{+/-x} for relative
                     % font size changes
\usepackage[german,english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc} 
\usepackage{subfig}
\usepackage{xypic}
\usepackage{url}
\usepackage{tikz}
\usetikzlibrary{arrows,shapes,automata,backgrounds,petri,fit,decorations.pathmorphing}
\usepackage{refcount}
\usepackage{xspace} % for \xspace in definition of acronyms etc.
\usepackage{attrib} % for right-aligned references at the end of
                    % quotes; part of Frankenstein bundle, but 
\renewcommand\PreTrib {}   % overwrite these commands, from attrib.sty
\renewcommand\PostTrib {}  % to suppress additional brackets around
                           % attributions
\usepackage{eurosym}
\usepackage{pgfplots}
\usepackage{todonotes}

\usepackage{gb4e_micha}

\usepackage{mycommands}

% \usepackage{geometry}
% \linespread{1.2}

%%%% Additional Macros

\newcommand{\lit}{\acro{lit}}
\newcommand{\glb}{\acro{glb}}
\newcommand{\loc}{\acro{loc}}

\newcommand{\as}{\acro{as}}
\renewcommand{\es}{\acro{es}}
\renewcommand{\AE}{\as}
\newcommand{\GE}{\es}
\newcommand{\ea}{\acro{ea}}

\newcommand{\lc}{\acro{lc}}
\newcommand{\ec}{\acro{ec}}
\newcommand{\LC}{\lc}
\newcommand{\EC}{\ec}

\newcommand{\exh}{\ensuremath{\mathrm{Exh}}}
\newcommand{\alt}{\ensuremath{\mathrm{Alt}}}

\newcommand{\ivt}{\acro{ivt}}

\renewcommand{\mymark}[1]{\textbf{#1}}

\pgfplotscreateplotcyclelist{mylist}{% 
    {solid,black},
    {dotted,black}, 
    {dashed,black}, 
    {loosely dotted,gray}, 
    {solid,black},
    {dotted,black},
    {dashed,black},
    {loosely dotted}}

\usetikzlibrary{calc}
\usepackage{nicefrac}

%%%% Document

\title{Embedded Scalars, Preferred Readings \& Prosody: {A}n
  Experimental Revisit} 
\author{% Fabian
  % Schlotterbeck, Michael Franke and Petra Augurzky
} 
\date{}

\begin{document}
\maketitle

\begin{abstract}
  The scalar item \emph{some} is widely assumed to receive a pragmatic
  meaning enrichment to \emph{some but not all} if it occurs in matrix
  position. The question under which circumstances this enrichment can
  occur in certain embedded positions plays an important role in
  deciding how to delineate semantics and pragmatics. % Some claim
  % that it requires marked intonation for the relevant pragmatic
  % enrichment to occur in embedded position; others allow for it more
  % freely.
  We present new experimental data that bear on this theoretical
  issue. In distinction to previous experimental approaches, we
  presented sentence material auditorily in order to explicitly
  control prosodic markedness of the scalar item. Moreover, our
  experiment sheds light on the relative preferences or salience of
  candidate readings. The presented data turn out to be problematic
  for both a traditional Gricean view of scalar implicature
  calculation, as well as the more recent grammatical approach.
\end{abstract}

\section{Introduction}
\label{sec:introduction}

The existential quantifier \emph{some} is usually assumed to receive a
semantic interpretation similar to logical $\exists$, so that
(\ref{bsp:Plain-SI-Target}) is literally true even when Hans solved
all of the problems. But use of \emph{some} is also usually considered
to invite comparison with (at least) the semantically stronger
universal quantifier \emph{all}
\citep[c.f.][]{Horn1972:On-the-Semantic,Gazdar1979:Pragmatics:-Imp,AtlasLevinson1981}. This
scalar comparison can lead to an upper-bound meaning enrichment, e.g.,
when an utterance of (\ref{bsp:Plain-SI-Target}) is taken to invite
the inference in (\ref{bsp:Plain-SI-Implicature}).

\begin{exe}
  \ex \label{bsp:Plain-SI}
    \begin{xlist}
      \ex \label{bsp:Plain-SI-Target} Hans solved some of the
        problems.
      \ex \label{bsp:Plain-SI-Implicature} $\implicates$ Hans solved
        some but not all of the problems.
      \ex \label{bsp:Plain-SI-Alternative} Hans solved all of the problems.
    \end{xlist}
\end{exe}

\noindent The classical explanation of this inference, following the
pioneering work of \citet{Grice1975:Logic-and-Conve}, is that
(\ref{bsp:Plain-SI-Implicature}) is a pragmatic inference, a so-called
\emph{quantity implicature}, derived by an abductive inference to the
best explanation of why informed, knowledgable and cooperative
speakers would utter (\ref{bsp:Plain-SI-Target}) when they could also
utter the semantically stronger and relevant
(\ref{bsp:Plain-SI-Alternative}) \citep[see][for a recent
overview]{Geurts2010:Quantity-Implic}.

% If this upper-bound inference would occur often and
% systematically enough, then it may well be that also embedded
% occurrences of \emph{some} get enriched, in some fashion or other, to
% contribute the enriched meaning \emph{some but not all} also under the
% scope of other logical operators.
% Indeed, even
% \citeauthor{Grice1975:Logic-and-Conve} envisaged this possibility when
% he wrote: ``It may not be impossible for what starts life, so to
% speak, as a conversational implicature to become conventionalized''
% \citep[p.58]{Grice1975:Logic-and-Conve}.

Whereas the existence of such implicatures is rather uncontroversial
for plain occurrences of \emph{some}, it is still an issue of debate
whether comparable enrichments are also found in so-called
``embedded'' cases, where the scalar item \emph{some} occurs in the
scope of other logical operators. This paper deals with two such
cases.  In \as-sentences (mnemonic for \textit{all \dots some \dots})
as in (\ref{bsp:AE}) the scalar item \emph{some} is embedded the under
universal quantifier \emph{all}. In \es-sentences (mnemonic for
\textit{exactly one \dots some \dots}) like (\ref{bsp:GE}) \emph{some}
takes scope under the non-monotonic quantifier \emph{exactly one}.
Current theories, have proposed at least three candidate readings
for \as- and \es-sentences: (i) a \emph{literal reading} like in (\ref{bsp:AE-Literal}) and
(\ref{bsp:GE-Literal}) where \emph{some} has only its literal meaning;
(ii) a \emph{global reading} like in (\ref{bsp:AE-Global}) and
(\ref{bsp:GE-Global}) where the meaning of (\ref{bsp:AE}) and
(\ref{bsp:GE}) is enriched with the negation of the corresponding
sentences (\ref{bsp:AE-Alternative}) and (\ref{bsp:GE-Alternative})
respectively; and also (iii) a \emph{local reading} like in
(\ref{bsp:AE-Local}) and (\ref{bsp:GE-Local}) where \emph{some} is
interpreted as \emph{some but not all} in the scope of the embedding
quantifier.

\begin{exe}
  \ex \label{bsp:AE} \mymark{All} of the students read {\mymark{some}} of the
  papers. \hfill{(\as)}

  \begin{xlist}
  \ex \label{bsp:AE-Literal} \mymark{All} of the students read
    {\mymark{some and maybe all}} of the papers. \hfill (\as-\lit)
  \ex \label{bsp:AE-Global}
    \mymark{All} of the students read \mymark{some and maybe all} 
    and  \hfill (\as-\glb)\\
    it's not the case that \mymark{all} of the students read \mymark{all} of the papers.
  \ex \label{bsp:AE-Local}
    \mymark{All} of the students read {\mymark{some  but not all}} of the
    papers. \hfill (\as-\loc)
  \end{xlist}
\end{exe}

\begin{exe}
\ex  \label{bsp:GE} \mymark{Exactly one} of the students read {\mymark{some}} of the
  papers. \hfill{(\es)}

  \begin{xlist}
  \ex \label{bsp:GE-Literal} \mymark{Exactly one} of the students read
    {\mymark{some and maybe all}} of the papers. \hfill (\es-\lit)
  \ex \label{bsp:GE-Global}
    \mymark{Exactly one} of the students read \mymark{some and maybe all} 
    and  \hfill (\es-\glb)\\
    it's not the case that \mymark{exactly one} of the students read \mymark{all} of the papers.
  \ex \label{bsp:GE-Local}
    \mymark{Exactly one} of the students read {\mymark{some  but not all}} of the
    papers. \hfill (\es-\loc)
  \end{xlist}
\end{exe}

\begin{exe}
\ex \label{bsp:AE-Alternative} \mymark{All} of the students read
  {\mymark{all}} of the papers.

\ex \label{bsp:GE-Alternative} \mymark{Exactly one} of the students
  read {\mymark{all}} of the papers.
\end{exe}

Given this plurality of hypothesized readings, the question arises:
how salient are these theoretically conceivable readings
compared to each other (if they are detectable by empirical means at
all)?  Addressing this question empirically is relevant because it
lies at the heart of a current debate about the nature of quantity
inferences and the demarkation between grammar (i.e. syntax/semantics)
and domain-general reasoning procedures (i.e. pragmatics). The controversy
is usually seen as one between two main competing theoretical positions
\citep[see][for a detailed comparison of the two]{ChemlaSingh2014:Remarks-on-the-}.
Simplifying things for now, these two types of theories make clearly different
predictions about how salient the local readings, in particular, should be.
The first position which we refer to as \mymark{traditionalism} (Chemla \& Singh’s ``Gricean approach'') holds that local readings should not be salient at all.
The second one which we call \mymark{grammaticalism}
(Chemla \& Singh’s ``grammatical approach'') holds that local readings
are unproblematically available and even at least as salient
as the literal ones (see Section \ref{sec:theories-predictions} for a more
thorough exposition of the theories and their predictions).

That matters actually are not as simple as just described is reflected
by lively discussions about which conclusions are permitted by previous
experimental research \citep[e.g.][]{GeurtsPouscoulous2009:Embedded-Implic,CliftonDube2010:Embedded-Implic,ChemlaSpector2010:Experimental-Ev,BenzGotzner2014:Embedded-implic}. The results of these studies
were discussed controversially because of several specific reasons. Firstly,
and most importantly, previous studies presented target sentences visually.
However, traditionalists often acknowledge the availability of local readings for
prosodically marked utterances, albeit not as the result of
a standard scalar implicature but of a different pragmatic process triggered by
prosodic markedness \citep[e.g.][]{Horn2006:The-Border-Wars,Geurts2009:Scalar-Implicat,ChemlaSpector2010:Experimental-Ev,Geurts2010:Quantity-Implic,Tielvan-Tiel2012:Embedded-Scalar,GeurtsTielvan-Tiel2013:Embedded-Scalar}. Under this view, evidence for the availability of local readings \citep[see][]{CliftonDube2010:Embedded-Implic,ChemlaSpector2010:Experimental-Ev} may be
an effect of “silent prosody”  \citep[see e.g.][]{Bader98,Fodor98}. Secondly, as \citet{Tielvan-Tiel2012:Embedded-Scalar} and \citet{GeurtsTielvan-Tiel2013:Embedded-Scalar} argue,
the alleged evidence for the availability of local readings of \as-sentences, might well be
confounded with “typicality effects”  (see Section~\ref{sec:previous-studies} for discussion).
Finally, previous studies have only accumulated limited evidence pertaining to the relative
salience of candidate readings of sentences like (\ref{bsp:AE}) and (\ref{bsp:GE}) (see
Section~\ref{sec:previous-studies} for discussion).

More generally, two prerequisites have to be met in order to make use of
experimental data to decide between theoretical positions regarding the
computation of quantity inferences. Firstly, both traditionalism and grammaticalism
have to be amended with additional assumptions to derive testable
predictions about the relative salience of candidate readings. For example,
hypothezised effects of prosodic marking have to be taken into account before
we can test traditionalism empirically using sentences  (\ref{bsp:AE}) and (\ref{bsp:GE}).
We will address this extant position as the \mymark{prosodic markedness hypothesis}
and argue that, if prosody is hypothesized to play a central role, it is vital
to present target sentences auditorily or to control for effects of  “silent prosody”
that subjects may apply when reading a sentence. Secondly, the relative
salience of candidate readings has to be operationalized properly, i.e. we
have to devise methods to measure relative salience. For example, typicality
effects and logical dependencies between readings have to be controlled for
when evaluating speaker judgments in terms of the salience of candidate readings.
Both of these prerequisites are related to formulating what \citet{ChemlaSingh2014:Remarks-on-the-}
call ``linking functions'' which connect the core theories to the dependent
variables of an experiment. Without proper linking functions it is hard
(or impossible) to test traditionalism or grammaticalism empirically.

This paper presents results form an experimental study designed to deal with
the mentioned problems. The focus is on testing certain formulations of traditionalism
and grammaticalsim that are explicit enough to derive testable predictions about the
relative salience of candidate readings. For practical reasons, we choose what we consider
the most prominent of such formulations. In order to address the prosodic markedness
hypothesis, we presented sentence materials auditorily and explicitly manipulated contrastive
stress on embedded scalar items. In order operationalize the relative salience of candidate
readings, we employed an \mymark{incremental verification task} in which parts of a to-be-evaluated
picture were hidden and only revealed piece-wise at the participants’ request until they felt able
to give a binary truth-value judgement \citep[see][]{Conroy2008}. This way, we catered for
potential pictorial confounds, such as from typicality. To analyze the obtained data in terms
of relative salience of candidate readings, we furthermore employed a statistical model of
how the salience of readings guided participants judgments in the incremental verification task. We
validated this method by including control conditions testing ambiguous sentences with a
well-known preference on available readings. We argue that this method provides
categorical behavioral data that is both clearly indicative of the reading preferences
among participants and independent of any potential pictorial confounds, such as
the mentioned typicality effects.

The data obtained by this method is \emph{prima facie} problematic for both
traditionalism and grammaticalism (or at least what we consider the most
prominent versions thereof). However, we want to stress from the outset
that, while the experimental results allow us to derive restrictions on which
linking functions are conceivable for traditionalism and grammaticalism, conclusions
about the general validity of the two core theories will be limited
\citep[cf.][]{ChemlaSingh2014:Remarks-on-the-} because these hardly make falsifiable
predictions without the necessary linking functions. Contrary to the predictions
of common versions of traditionalism, we observed many answers indicative of
local readings of \es-sentences, irrespective of prosodic markedness. But also
prominent versions of grammaticalism are incongruent with the observed data.
Backed up by the \mymark{strongest meaning hypothesis}
\citep{DalrympleKanazawa1998:Reciprocal-Expr} as a
disambiguation criterion \citep[as suggested by][]{ChierchiaFox2008:The-Grammatical},
grammaticalism predicts local readings to be the preferred ones, but
we found no evidence for that. More strongly even, the preference order among
putative readings suggested by our data cannot even in principle be
explained by comparisons of logical strength among putative
readings. In sum, our data shed doubt on both traditionalism and grammaticalism
in their currently most common forms. We conclude the paper by some thoughts on
how either position could be amended in the light of our data.

The remainder of the paper is structured as
follows. Section~\ref{sec:get-know-your} elaborates on the three kinds
of relevant readings for our target
sentences. Section~\ref{sec:theories-predictions} works out the
predictions of traditionalism and grammaticalism (plus prominent linking functions) 
about availability and relative salience of readings for these
sentences. Section~\ref{sec:previous-studies} reviews related
experimental studies to the extent necessary to motivate our own
approach. Section~\ref{sec:exp} eventually contributes our
experimental design, analyses and results.

\section{Get to know your readings}
\label{sec:get-know-your}

Three readings are \emph{prima facie} conceivable for the \as- and
\es-sentences in (\ref{bsp:AE}) and (\ref{bsp:GE}). These are
logically dependent in intricate ways.

\paragraph{\as-sentences.}

An \as-sentence like (\ref{bsp:AE}), repeated below, has a literal
reading (\lit) as in (\ref{bsp:AE-Literal}), a global reading (\glb)
as paraphrased in (\ref{bsp:AE-Global}) and a local reading (\loc) as
in (\ref{bsp:AE-Local}).

\begin{exer}{bsp:AE}

  \ex \mymark{All} of the students read {\mymark{some}} of the
  papers.

  \begin{xlist}
  \ex \mymark{All} of the students read
    {\mymark{some and maybe all}} of the papers. \hfill (\as-\lit)
  \ex
    \mymark{All} of the students read \mymark{some and maybe all} 
    and  \hfill (\as-\glb)\\
    it's not the case that \mymark{all} of the students read \mymark{all} of the papers.
  \ex
    \mymark{All} of the students read {\mymark{some  but not all}} of the
    papers. \hfill (\as-\loc)
  \end{xlist}
\end{exer}

\noindent These readings stand in a strict entailment relation: the
local reading asymmetrically entails the global reading, which
asymmetrically entails the literal reading:
\begin{exe}
  \ex \label{bsp:Entailments-AS} \loc $\subset$ \glb $\subset$ \lit
\end{exe}
\glb entails \lit because, in general, global readings are defined as
the conjunction of the literal reading and the negated
(relevant/feasible) alternative(s) of the to-be-interpreted
utterance. This entailment is asymmetric, because the information that
not all of the students read all of the papers is not entailed by the
literal reading. To see that \loc $\subset$ \glb, notice that the case
where all of the students read some but not all of the papers is a
special case of the class of situations where all of the students read
some (and maybe all), while not all of the students read all of the
papers.

Given these entailment relations, there are four kinds of situations,
the names for which we borrow from
\citet{ChemlaSpector2010:Experimental-Ev}, that we can distinguish
based on different truth values for our candidate readings. These are
given in Table~\ref{fig:situations-AE}.
%
\begin{table}
  \centering
  \subfloat[][\as-sentences]{
  \label{fig:situations-AE}
    \begin{tabular}{lccc}
    \toprule
    situation    & \multicolumn{3}{c}{reading} 
  \\ 
  \cmidrule(r){2-4}
     & \lit & \glb & \loc \\ \midrule
    false   & 0 & 0 & 0 \\
    literal & 1 & 0 & 0 \\
    weak    & 1 & 1 & 0 \\
    strong  & 1 & 1 & 1 \\ \bottomrule
  \end{tabular}
} \qquad \qquad
  \subfloat[][\es-sentences]{
  \label{fig:situations-GS}
  \begin{tabular}{lccc}
    \toprule
    situation    & \multicolumn{3}{c}{reading} 
  \\ 
  \cmidrule(r){2-4}
    & \textsc{lit} & \textsc{glb} & \textsc{loc} \\ \midrule
    false   & 0 & 0 & 0 \\
    literal & 1 & 0 & 0 \\
    local   & 0 & 0 & 1 \\
    all     & 1 & 1 & 1 \\ \bottomrule
  \end{tabular}
}
\caption{Possible truth-value distributions for readings of \as- and \es-sentences}
\label{tab:situations}
\end{table}
%
Examples of these kinds of situations are shown in
Figure~\ref{fig:AS-distinguishing-pics}, where the dots on the right
of each diagram represent students, the dots on the left represent
papers and an arrow from a student to a paper indicates that the
student read the paper. % Notice that other arrangements of arrows might
% equally well serve as examples for the various situations.

\begin{figure}[t]
  \centering

\subfloat[fig:false][false]{
  \label{fig:false-AE}

\begin{tikzpicture}[node distance = 1cm]
        % Nodes

        \node (X-1) {$\Large{\bullet}$};

        \node (X-2) [below of = X-1] {$\Large{\bullet}$};

        \node (X-3) [below of = X-2] {$\Large{\bullet}$};

        \node (Y-1) [right of = X-1, node distance = 2cm]{$\Large{\bullet}$};

        \node (Y-2) [below of = Y-1] {$\Large{\bullet}$};

        \node (Y-3) [below of = Y-2] {$\Large{\bullet}$};

        \node (Y-4) [below of = Y-3] {$\Large{\bullet}$};

        % table added from here

        \node (X-4) [below of = X-3, node distance = 2cm] {};

        \node (table) [right of = X-4] {
          \begin{tabular}{ll}
            \lit & false \\
            \glb & false \\
            \loc & false 
          \end{tabular}};

        % Arrows

        \path [draw=blue,->] (X-1) -> (Y-1);

        \path [draw=blue,->] (X-1) -> (Y-2);

        \path [draw=red,->] (X-3) -> (Y-3);

        \path [draw=red,->] (X-3) -> (Y-4);

      \end{tikzpicture}

}
\subfloat[fig:literal][literal]{
  \label{fig:literal-AE}

  \begin{tikzpicture}[node distance = 1cm]
        % Nodes

        \node (X-1) {$\Large{\bullet}$};

        \node (X-2) [below of = X-1] {$\Large{\bullet}$};

        \node (X-3) [below of = X-2] {$\Large{\bullet}$};

        \node (Y-1) [right of = X-1, node distance = 2cm]{$\Large{\bullet}$};

        \node (Y-2) [below of = Y-1] {$\Large{\bullet}$};

        \node (Y-3) [below of = Y-2] {$\Large{\bullet}$};

        \node (Y-4) [below of = Y-3] {$\Large{\bullet}$};

        % table added from here

        \node (X-4) [below of = X-3, node distance = 2cm] {};

        \node (table) [right of = X-4] {
          \begin{tabular}{ll}
            \lit & true \\
            \glb & false \\
            \loc & false 
          \end{tabular}};

        % Arrows

        \path [draw=blue,->] (X-1) -> (Y-1);

        \path [draw=blue,->] (X-1) -> (Y-2);

        \path [draw=blue,->] (X-1) -> (Y-3);

        \path [draw=blue,->] (X-1) -> (Y-4);

        \path [draw=black,->] (X-2) -> (Y-1);

        \path [draw=black,->] (X-2) -> (Y-2);

        \path [draw=black,->] (X-2) -> (Y-3);

        \path [draw=black,->] (X-2) -> (Y-4);

        \path [draw=red,->] (X-3) -> (Y-1);

        \path [draw=red,->] (X-3) -> (Y-2);

        \path [draw=red,->] (X-3) -> (Y-3);

        \path [draw=red,->] (X-3) -> (Y-4);

      \end{tikzpicture}

}
\subfloat[fig:weak][weak]{
  \label{fig:weak}

\begin{tikzpicture}[node distance = 1cm]
        % Nodes

        \node (X-1) {$\Large{\bullet}$};

        \node (X-2) [below of = X-1] {$\Large{\bullet}$};

        \node (X-3) [below of = X-2] {$\Large{\bullet}$};

        \node (Y-1) [right of = X-1, node distance = 2cm]{$\Large{\bullet}$};

        \node (Y-2) [below of = Y-1] {$\Large{\bullet}$};

        \node (Y-3) [below of = Y-2] {$\Large{\bullet}$};

        \node (Y-4) [below of = Y-3] {$\Large{\bullet}$};

        % table added from here

        \node (X-4) [below of = X-3, node distance = 2cm] {};

        \node (table) [right of = X-4] {
          \begin{tabular}{ll}
            \lit & true \\
            \glb & true \\
            \loc & false 
          \end{tabular}};

        % Arrows

        \path [draw=blue,->] (X-1) -> (Y-1);

        \path [draw=blue,->] (X-1) -> (Y-2);

        \path [draw=black,->] (X-2) -> (Y-1);

        \path [draw=black,->] (X-2) -> (Y-2);

        \path [draw=black,->] (X-2) -> (Y-3);

        \path [draw=black,->] (X-2) -> (Y-4);

        \path [draw=red,->] (X-3) -> (Y-3);

        \path [draw=red,->] (X-3) -> (Y-4);

      \end{tikzpicture}

}
\subfloat[fig:strong][strong]{
  \label{fig:strong}

\begin{tikzpicture}[node distance = 1cm]
        % Nodes

        \node (X-1) {$\Large{\bullet}$};

        \node (X-2) [below of = X-1] {$\Large{\bullet}$};

        \node (X-3) [below of = X-2] {$\Large{\bullet}$};

        \node (Y-1) [right of = X-1, node distance = 2cm]{$\Large{\bullet}$};

        \node (Y-2) [below of = Y-1] {$\Large{\bullet}$};

        \node (Y-3) [below of = Y-2] {$\Large{\bullet}$};

        \node (Y-4) [below of = Y-3] {$\Large{\bullet}$};

        % table added from here

        \node (X-4) [below of = X-3, node distance = 2cm] {};

        \node (table) [right of = X-4] {
          \begin{tabular}{ll}
            \lit & true \\
            \glb & true \\
            \loc & true 
          \end{tabular}};

        % Arrows

        \path [draw=blue,->] (X-1) -> (Y-1);

        \path [draw=blue,->] (X-1) -> (Y-2);

        \path [draw=black,->] (X-2) -> (Y-2);

        \path [draw=black,->] (X-2) -> (Y-3);

        \path [draw=red,->] (X-3) -> (Y-3);

        \path [draw=red,->] (X-3) -> (Y-4);

      \end{tikzpicture}

}

\caption{Examples of distinguishing situations for \as-sentences}
  \label{fig:AS-distinguishing-pics}
\end{figure}

\paragraph{\es-sentences.}

The case for \es-sentences like (\ref{bsp:GE}) is similar but a
little more complicated because the embedding quantifier is
non-monotonic. Again, we consider a literal reading as in
(\ref{bsp:GE-Literal}), repeated below, a global reading as in
(\ref{bsp:GE-Global}) and a local reading as in (\ref{bsp:GE-Local}).

\begin{exer}{bsp:GE}
\ex \mymark{Exactly one} of the students read {\mymark{some}} of the
  papers.

  \begin{xlist}
  \ex \mymark{Exactly one} of the students read
    {\mymark{some and maybe all}} of the papers. \hfill (\es-\lit)
  \ex 
    \mymark{Exactly one} of the students read \mymark{some and maybe all} 
    and  \hfill (\es-\glb)\\
    it's not the case that \mymark{exactly one} of the students read \mymark{all} of the papers.
  \ex 
    \mymark{Exactly one} of the students read {\mymark{some  but not all}} of the
    papers. \hfill (\es-\loc)
  \end{xlist}
\end{exer}

\noindent Entailment relations in this case are non-linearly ordered:
\begin{exe}
  \ex \label{bsp:Entailments-GE}       \loc\ $\supset$ \glb\ $\color{mycol}{\subset}$ \lit 
    % \hspace{1cm} and \hspace{1cm}  \lit\ $\not \supset$ \loc
\end{exe}
By definition of global readings, $\glb$ entails $\lit$. This
entailment is asymmetric because the extra information that it is not
the case that exactly one student read all of the papers is not
entailed by the literal reading (\ref{bsp:GE-Literal}). However,
unlike for \as-sentences, \loc is not stronger than \glb, but
asymmetrically entailed by the latter. To see this, notice that the
global reading is equivalent to:

\begin{exe}
  \exp{bsp:GE-Global} \mymark{Exactly one} of the students read \mymark{some but not
    all} and \\
  \mymark{everybody else} read \mymark{none} of the papers.
\end{exe}

\noindent Finally, \loc and \lit are logically independent: all
combinations of truth-values for \loc and \lit are possible.

Given these entailment relations, there are again four different
situations corresponding to the four possible distributions of truth
values for candidate readings. These are given in
Table~\ref{fig:situations-GS} on page~\pageref{fig:situations-GS} and
named following \citet{ChemlaSpector2010:Experimental-Ev}. Examples
for each situation are given in
Figure~\ref{fig:ES-distinguishing-pics}.

\begin{figure}[t]
  \centering

\subfloat[fig:false][false]{
  \label{fig:false-GE}

  \begin{tikzpicture}[node distance = 1cm]
        % Nodes

        \node (X-1) {$\Large{\bullet}$};

        \node (X-2) [below of = X-1] {$\Large{\bullet}$};

        \node (X-3) [below of = X-2] {$\Large{\bullet}$};

        \node (Y-1) [right of = X-1, node distance = 2cm]{$\Large{\bullet}$};

        \node (Y-2) [below of = Y-1] {$\Large{\bullet}$};

        \node (Y-3) [below of = Y-2] {$\Large{\bullet}$};

        \node (Y-4) [below of = Y-3] {$\Large{\bullet}$};

        % table added from here

        \node (X-4) [below of = X-3, node distance = 2cm] {};

        \node (table) [right of = X-4] {
          \begin{tabular}{ll}
            \lit & false \\
            \glb & false \\
            \loc & false 
          \end{tabular}};

        % Arrows

        \path [draw=blue,->] (X-1) -> (Y-1);

        \path [draw=blue,->] (X-1) -> (Y-2);

        \path [draw=red,->] (X-3) -> (Y-3);

        \path [draw=red,->] (X-3) -> (Y-4);

      \end{tikzpicture}

}
\subfloat[fig:literal][literal]{
  \label{fig:literal-GE}

      \begin{tikzpicture}[node distance = 1cm]
        % Nodes

        \node (X-1) {$\Large{\bullet}$};

        \node (X-2) [below of = X-1] {$\Large{\bullet}$};

        \node (X-3) [below of = X-2] {$\Large{\bullet}$};

        \node (Y-1) [right of = X-1, node distance = 2cm]{$\Large{\bullet}$};

        \node (Y-2) [below of = Y-1] {$\Large{\bullet}$};

        \node (Y-3) [below of = Y-2] {$\Large{\bullet}$};

        \node (Y-4) [below of = Y-3] {$\Large{\bullet}$};

        % table added from here

        \node (X-4) [below of = X-3, node distance = 2cm] {};

        \node (table) [right of = X-4] {
          \begin{tabular}{ll}
            \lit & true \\
            \glb & false \\
            \loc & false 
          \end{tabular}};

        % Arrows

        \path [draw=black,->] (X-2) -> (Y-1);

        \path [draw=black,->] (X-2) -> (Y-2);

        \path [draw=black,->] (X-2) -> (Y-3);

        \path [draw=black,->] (X-2) -> (Y-4);

      \end{tikzpicture}

}
\subfloat[fig:local][local]{
  \label{fig:local}

            \begin{tikzpicture}[node distance = 1cm]
        % Nodes

        \node (X-1) {$\Large{\bullet}$};

        \node (X-2) [below of = X-1] {$\Large{\bullet}$};

        \node (X-3) [below of = X-2] {$\Large{\bullet}$};

        \node (Y-1) [right of = X-1, node distance = 2cm]{$\Large{\bullet}$};

        \node (Y-2) [below of = Y-1] {$\Large{\bullet}$};

        \node (Y-3) [below of = Y-2] {$\Large{\bullet}$};

        \node (Y-4) [below of = Y-3] {$\Large{\bullet}$};

        % table added from here

        \node (X-4) [below of = X-3, node distance = 2cm] {};

        \node (table) [right of = X-4] {
          \begin{tabular}{ll}
            \lit & false \\
            \glb & false \\
            \loc & true 
          \end{tabular}};

        % Arrows

        \path [draw=blue,->] (X-1) -> (Y-1);

        \path [draw=blue,->] (X-1) -> (Y-2);

        \path [draw=blue,->] (X-1) -> (Y-3);

        \path [draw=blue,->] (X-1) -> (Y-4);

        \path [draw=black,->] (X-2) -> (Y-2);

        \path [draw=black,->] (X-2) -> (Y-3);

        \path [draw=black,->] (X-2) -> (Y-4);

        \path [draw=red,->] (X-3) -> (Y-1);

        \path [draw=red,->] (X-3) -> (Y-2);

        \path [draw=red,->] (X-3) -> (Y-3);

        \path [draw=red,->] (X-3) -> (Y-4);

      \end{tikzpicture}

}
\subfloat[fig:all][all]{
  \label{fig:all}

      \begin{tikzpicture}[node distance = 1cm]
        % Nodes

        \node (X-1) {$\Large{\bullet}$};

        \node (X-2) [below of = X-1] {$\Large{\bullet}$};

        \node (X-3) [below of = X-2] {$\Large{\bullet}$};

        \node (Y-1) [right of = X-1, node distance = 2cm]{$\Large{\bullet}$};

        \node (Y-2) [below of = Y-1] {$\Large{\bullet}$};

        \node (Y-3) [below of = Y-2] {$\Large{\bullet}$};

        \node (Y-4) [below of = Y-3] {$\Large{\bullet}$};

        % table added from here

        \node (X-4) [below of = X-3, node distance = 2cm] {};

        \node (table) [right of = X-4] {
          \begin{tabular}{ll}
            \lit & true \\
            \glb & true \\
            \loc & true 
          \end{tabular}};

        % Arrows

        \path [draw=blue,->] (X-1) -> (Y-1);

        \path [draw=blue,->] (X-1) -> (Y-2);

      \end{tikzpicture}

}

  \caption{Examples of distinguishing situations for \es-sentences}
  \label{fig:ES-distinguishing-pics}

\end{figure}

\section{Theories and predictions}
\label{sec:theories-predictions}

There are two main theoretical positions which make different 
predictions about the readings of \as- and \es-sentences \citep[see][for
overview]{Horn2006:The-Border-Wars,Geurts2010:Quantity-Implic,Sauerland2012:The-Computation,ChemlaSingh2014:Remarks-on-the-}. We will refer to these as \mymark{traditionalism} and \mymark{grammaticalism}
and treat each one in turn.  As already mentioned in the introduction there is some
leeway in assessing the predictions for each approach depending on which of several
additional assumptions to adopt. Therefore, we will distinguish different varieties
of both of them. While these are certainly not the only possible varieties, they are,
in our perception, the most prominent ones. Focusing on exactly these thus has
practical relevance. Their predictions are summarized in Table~\ref{tab:predictions}
on page~\pageref{tab:predictions}. Below, in the discussion section
(section~\ref{sec:discussion-1}), further conceivable varieties of traditionalism
and grammaticalism will be discussed.

Although strictly speaking not a part of a Gricean theory of scalar
implicature calculation, traditionalists frequently concede that local
readings are in principle available, albeit not as a standard scalar
implicature, but as a different pragmatic process triggered by
prosodic markedness. We address this \emph{prosodic markedness
  hypothesis} separately in Section~\ref{sec:inton-mark-hypoth},
alongside a brief revision of the relevant experimental data.

\subsection{Traditionalism}
\label{sec:traditionalism}

Traditionalism gets its name from its conservative stance towards
Grice's original theory of conversational implicatures
\citep{Grice1975:Logic-and-Conve}. Many authors have defended
traditionalist positions in this sense. Of the more recent literature,
we would consider as traditionalist, among others, contributions by
\citet{Spector2006:Scalar-Implicat},
\citet{Sauerland2004:Scalar-Implicat},
\citet{Russell2006:Against-Grammat},
\citet{vanRooijSchulz:ExhaustiveInterpretation},
\citet{Geurts2010:Quantity-Implic} or
\citet{Franke2011:Quantity-Implic}.

According to Grice, conversational implicatures, of which quantity
implicatures are a special case, are to be thought of as
rationalizations of speaker behavior. Central in this reasoning is the
assumption that the speaker's behavior is efficient (if not optimal)
and goal-oriented. Usually, the assumed goal of conversation is the
cooperative exchange of relevant information from speaker to hearer.

Consequently, the \mymark{Gricean recipe}
\citep[see][]{Geurts2010:Quantity-Implic} for deriving a simple scalar
inference like that in (\ref{bsp:Plain-SI-Implicature}) from an
utterance of (\ref{bsp:Plain-SI-Target}) is as follows: if the issue
whether Hans solved only some or all of the problems is relevant, then
a cooperative and knowledgable speaker would utter
(\ref{bsp:Plain-SI-Alternative}) if in a position to do so; hence, one
of the most natural explanations of why such a speaker has not uttered
(\ref{bsp:Plain-SI-Alternative}), but only (\ref{bsp:Plain-SI-Target})
is that she is uncertain of whether (\ref{bsp:Plain-SI-Alternative})
is true; but on the assumption that she is knowledgeable (competent,
opinionated, informed \dots) it follows that
(\ref{bsp:Plain-SI-Implicature}) should in fact be true.\fn{We are
  glossing here somewhat swiftly over the more nuanced details of the
  derivation of implicatures targeting the speaker's epistemic state
  \citep[e.g.][]{Gazdar1979:Pragmatics:-Imp,Soames1982:How-Presupposit}. Section~\ref{sec:discussion-1}
  will come back to this in more detail.}

\begin{exer}{bsp:Plain-SI}
  \ex 
    \begin{xlist}
      \ex \label{bsp:Plain-SI-Target} Hans solved \mymark{some} of the problems.
      \ex \label{bsp:Plain-SI-Implicature} $\implicates$ Hans solved
        \mymark{some but not all} of the problems.
      \ex  \label{bsp:Plain-SI-Alternative}  Hans solved \mymark{all} of the problems.
    \end{xlist}
\end{exer}

As for \as- and \es-sentences, the Gricean recipe derives global
readings in a straightforward way. Consequently, traditionalism
predicts that both literal and global readings are available:  literal
readings, because these are the result of semantic interpretation
and form the starting point of pragmatic reasoning; global readings
because these may be arrived at by the Gricean recipe.

Traditionalism's predictions about relative salience of literal and
global readings depend on whether certain auxiliary assumptions
necessary to derive global readings by the Gricean recipe are
plausibly met. These auxiliary assumptions include relevance of the
extra information provided by the global reading, mutual awareness
that the stronger alternative has been a speaker option, the speaker's
competence about the issue, etc. Normally, traditionalist accounts
would assume that these extra assumptions are met. In that case,
traditionalism would predict that global readings should be preferred
over literal readings. On the other hand, if these extra assumptions
are implausible, traditionalism would predict that literal readings
are preferred over global ones \citep[see][for experimental evidence
indicating that manipulation of speaker knowledgeability indeed
affects the rate of quantity implicatures drawn by the listener
for unembedded occurrences of \emph{some}]{GoodmanStuhlmuller2013:Knowledge-and-I}.

Consequently, there are at least two varieties of traditionalism that,
depending on which extra assumptions we make, yield slightly
different predictions: the \mymark{the strong variety of traditionalism}
maintains that the extra assumptions of the Gricean recipe hold
(e.g., as a matter of strong default) and therefore predicts that global
readings are preferred over literal ones; \mymark{the weak variety of
traditionalism} holds that the auxiliary assumptions are more fragile
and context dependent and thus predicts that literal readings may
often be preferred over global ones. In Table~\ref{tab:predictions}
(p.~\pageref{tab:predictions}), this prediction of weak traditionalism
is captured by a preference of literal over global readings because it
may \emph{a priori} seem implausible from the perspective of weak
traditionalism that all the necessary assumptions to derive global
readings are met. \fn{We mention for clarity that although weak traditionalism does
  not predict global readings to be generally preferred, it might still predict
  an \emph{epistemically weak implicature}, similar to the global
  reading, that the speaker is uncertain whether the stronger
  alternative is true. (See also the discussion in
  Section~\ref{sec:discussion-1})} For weak traditionalism, it is of course an empirical
task to determine under which circumstances the auxiliary assumptions
are met.

Does traditionalism also predict local readings to be available? The
answer is slightly different for \as- and the \es-sentences. For
\es-sentences, traditionalism does not predict that a local reading is
available, at least not as a quantity implicature by conjoining the
literal meaning with a suitable set of negated alternatives
\citep[see][]{GeurtsPouscoulous2009:Embedded-Implic,ChemlaSpector2010:Experimental-Ev}. The
literal and the local reading of \es-sentences are logically
independent. But if $X$ and $Y$ are logically independent
propositions, then there is no proposition $Z$ such that $X$ would be
equivalent to $Y \wedge \neg Z$.

On the other hand, as for \as-sentences, there is a traditionalist way
of deriving the local reading, namely by assuming that not only
(\ref{bsp:AE-Alternative}) is an alternative to (\ref{bsp:AE}), but
also the sentence in (\ref{bsp:AE-Alternatives-Extended}) \citep[e.g.][]{Sauerland2004:Scalar-Implicat}.

\begin{exer}{bsp:AE}
  \ex \mymark{All} of the students solved \mymark{some} of the problems.
\end{exer}

\begin{exer}{bsp:AE-Alternative}
  \ex \label{bsp:AE-Alternatives-Standard}\mymark{All} of the students solved \mymark{all} of the problems.
\end{exer}

\begin{exe}
\ex \label{bsp:AE-Alternatives-Extended} \mymark{Some} of the students solved \mymark{all} of the problems.
\end{exe}

\noindent Clearly, if we conjoin a literal reading of (\ref{bsp:AE})
with the negation of (\ref{bsp:AE-Alternatives-Extended}) we obtain
exactly the local reading. (Notice that the negation of
(\ref{bsp:AE-Alternatives-Extended}) entails the negation of
(\ref{bsp:AE-Alternative}), so it makes no difference (not) to add it
in this case.) Consequently, traditionalism does predict that the
local reading of \as-sentences is available if
(\ref{bsp:AE-Alternatives-Extended}) is an available alternative. We
therefore introduce another distinction: \mymark{restricted
  traditionalism} assumes that (\ref{bsp:AE-Alternatives-Extended}) is
not available and therefore predicts that the local reading for
\as-sentences is not available; in contrast, \mymark{unrestricted
  traditionalism} assumes that
(\ref{bsp:AE-Alternatives-Extended}) is available and so is the
local reading.\fn{The term traditionalism may be misleading here
because traditionalists usually assume that scalar implicatures should
only be computed when a sentence is uttered in a context in which a
logically stronger alternative is relevant \cite{ChemlaSingh2014:Remarks-on-the-}. However, the alternative
in (\ref{bsp:AE-Alternatives-Extended}) is not stronger than
(\ref{bsp:AE}). For completeness sake, we still decided to subsume
“unrestricted traditionalism” under traditionalism because, once
alternatives like (\ref{bsp:AE-Alternatives-Extended}) are acknowledged,
the local reading of \as-sentences can be derived in a
straightforward traditionalist manner \citep[see also][]{ChemlaSpector2010:Experimental-Ev}
who call this type of theory ``unrestricted globalism'').}

As far as preferences over readings of \as-sentences are concerned,
even unrestricted traditionalism would not predict that the local
reading is preferred over the global reading. This is because the
derivation of the local reading via
(\ref{bsp:AE-Alternatives-Extended}) hinges on yet another auxiliary
assumption, namely the availability of the alternative in
(\ref{bsp:AE-Alternatives-Extended}), which seems, by comparing
numbers of replacements, much less obvious and therefore presumably is
less readily available than the alternative in
(\ref{bsp:AE-Alternative}). Consequently, we derive the following
predictions about preferences for unrestricted traditionalism: weak
unrestricted traditionalism prefers the literal over the global over
the local reading, while strong unrestricted traditionalism prefers
the global over the local over the literal reading (see also
Table~\ref{tab:predictions}).

\subsection{Grammaticalism}
\label{sec:grammaticalism}

Recently, a lot of evidence based on intuitive judgements and more
theoretical considerations has been discussed in favor of
\mymark{grammaticalism}
\citep[see][]{Chierchia2006:Broaden-Your-Vi,Fox2007:Free-Choice-and,Magri2011:Another-Argumen,Sauerland2012:The-Computation,ChierchiaFox2008:The-Grammatical}. % When
% it comes to the predictions of available readings of \as- and
% \es-sentences, grammaticalism appears like the conjunction of
% traditionalism and conventionalism. But conceptually speaking,
% grammaticalism goes down a road of its own.
Grammaticalism approaches quantity implicatures by postulating a
silent operator that can be variably applied during compositional
computation of a sentence's truth-value
\citep{Chierchia2006:Broaden-Your-Vi}, if necessary multiple times
\citep{Fox2007:Free-Choice-and}. This silent operator is referred to
as $\mathrm{O}(\cdot)$ or $\exh(\cdot)$, because it is assumed to be
similar in effect to the meaning of particle \emph{only} or of the
mechanism of \emph{exhaustive interpretation}
\citep{GroenendijkStokhofThesis1984,Stechowvon-StechowZimmermann1984:Term-Answers-an,Rooijvan-RooijSchulz2013:Exhaustive-Inte,vanRooijSchulz:ExhaustiveInterpretation,Fox2007:Free-Choice-and}. For
our purposes, it is enough to note that $\exh(\cdot)$ is a poly-typed
function that enriches an expression $X$, which need not be a full
proposition, based on a set $\alt(X)$ of (suitable, relevant)
alternatives to $X$ that yields an enriched meaning of the
form:\fn{More sophisticated formulations of exhaustive interpretation
  have been proposed
  \citep[e.g.][]{Schulz2005:A-Pragmatic-Sol,vanRooijSchulz:ExhaustiveInterpretation,Spector2006:Scalar-Implicat,Fox2007:Free-Choice-and}.}
\begin{exe}
  \ex \label{bsp:Exh-Def} $\exh(X,\alt(X)) = X \bigwedge_{A \in
      \alt(X)} \neg A$.
\end{exe}

As this operator can apply at various scope sites, the grammatical
approach predicts that all three readings of \as- and \es-sentences
are available: literal readings arise if no exhaustification operator
is applied; global readings arise if the exhaustification operator
takes sentence-wide scope as in (\ref{Grammar-Global}); and local
readings arise if the exhaustification operator takes scope under the
respective quantifiers as in (\ref{Grammar-Local}).

\begin{exe}
  \ex \label{Grammar-Global}
    \begin{xlist}
      \ex \label{Grammar-Global-AE} \mymark{$\exh$}(\mymark{All} of the students solved
        \mymark{some} of the problems).
      \ex \label{Grammar-Global-GE} \mymark{$\exh$}(\mymark{Exactly one} of the students solved
        \mymark{some} of the problems).
    \end{xlist}
\end{exe}

\begin{exe}
  \ex \label{Grammar-Local}
    \begin{xlist}
      \ex \label{Grammar-Local-AE} \mymark{All} of the students \mymark{$\exh$}(solved
        \mymark{some} of the problems).
      \ex \label{Grammar-Local-GE} \mymark{Exactly one} of the
        students \mymark{$\exh$}(solved
        \mymark{some} of the problems).
    \end{xlist}
\end{exe}

The grammatical approach, as described so far, does not
make testable predictions about the relative salience of our three
candidate readings. In order to derive predictions about reading preferences,
a selection principle is needed that determines which of the
many readings made available by the grammar actually obtain or are
preferred. Proponents of grammaticalism are aware of this and have
suggested some tentative disambiguation
strategies. \citet{Magri2011:Another-Argumen}, for example, opts for
obligatory presence of $\exh$ with variability in the set of relevant
alternatives.  In the approach of \citet{Magri2011:Another-Argumen},
alternatives may be pruned entirely in the extreme case. In that case, all occurrences
of $\exh$ become vacuous resulting in a literal reading. However,
\citet{Magri2011:Another-Argumen} does not formulate an explicit theory
about which alternatives enter into the computation of sentence meaning
under what circumstances. Therefore, it is not clear what his approach
predicts about the relative salience of the candidate readings.

Other proponents of grammaticalism assume that insertion of $\exh$ is optional.
These approaches treat the ambiguity between our candidate readings as a
structural one which resembles syntactic ambiguities. In order to derive
predictions about reading preferences a mechanism of ambiguity resolution
is needed in these approaches. One proposal is that readings are the
more preferred the better they answer the contextually given question
under discussion \citep[e.g.][]{Fox2007:Free-Choice-and,GualminiHulsey2008:The-Question-An}.
Another proposal is to select readings by comparison of logical strengths, as suggested by the \mymark{strongest meaning hypothesis}
of \citet{DalrympleKanazawa1998:Reciprocal-Expr}
\citep[see][]{FoxSpector:Economy-and-Emb,ChierchiaFox2008:The-Grammatical,Chierchia2012:FC-Nominals-and,Spector2014:Global-Positive}. For now, we focus
exclusively on this latter selection principle, because it seems to be
the most prominent in the literature and because it makes the clearest
predictions. In the discussion section (section~\ref{sec:discussion-1})
alternative selection principles will be reconsidered.

\citet{ChierchiaFox2008:The-Grammatical} discuss two concrete variants
of strength-based disambiguation: fix a sentence with propositional
content $S$ and candidate readings $C(S)$, where $C(S)$ contains $S$
and all readings derivable from inserting $\exh(\cdot)$ at suitable
scope sites;\fn{Strictly speaking, we should compare \emph{parses} of
  sentences with $\exh(\cdot)$ at various scope sites with respect to
  the readings that these parses give rise to
  \citep[see][]{ChierchiaFox2008:The-Grammatical}, but we may ignore
  this detail here for ease of exposition.} then define for $X,Y \in
C(S)$ either
\begin{exe}
\ex \label{bsp:SMH-Total} 
  $X$ is preferred to $Y$ iff $X \subset Y$ 
  \attrib{\citep[(104)]{ChierchiaFox2008:The-Grammatical}}
\end{exe}
or
\begin{exe}
\ex \label{bsp:SMH-Partial}  $X$ is preferred to $Y$ iff $X \subset Y$
  and $X$ and $Y$ are readings obtained from applications of
  $\exh(\cdot)$ at exactly the same scope sites, except for
  one. \attrib{\citep[(105)]{ChierchiaFox2008:The-Grammatical}} 
\end{exe}
Since the former is entailed by the latter, we will speak of
\mymark{weak grammaticalism} and \mymark{strong grammaticalism},
depending on whether (\ref{bsp:SMH-Total}) or (\ref{bsp:SMH-Partial})
is used.

These two varieties of grammaticalism give rise to different
predictions about preferred readings. The preference for readings that
weak grammaticalism predicts mirrors the entailment relations in
(\ref{bsp:Entailments-AS}) and (\ref{bsp:Entailments-GE}): for
\as-sentences, weak grammaticalism predicts that local readings are
preferred over global readings which in turn are preferred over
literal readings; for \es-sentences weak grammaticalism predicts that
global readings are most preferred, while literal and local readings
are not ranked with respect to preference. Strong grammaticalism, on
the other hand, predicts that \as-sentences preferably get either a
local or a global reading rather than a literal reading, but does not
rank these former two with respect to each other because they differ
in more than one application of the $\exh(\cdot)$ operator. For
\es-sentences, then, strong grammaticalism does not predict \emph{any}
preference relations between candidate readings at all.

\medskip

Taking stock, traditionalism and grammaticalism make different
predictions about the relative salience (and thus preference) of our candidate
readings. Most notably, traditionalism does not predict local
readings for \es-sentences to be available at all while grammaticalism does.
Grammaticalism predicts that all candidate readings are available and,
employing the strongest meaning hypothesis as selection principle, also
comes with a clear prediction about preference: literal meanings are dispreferred.
All predictions are summarized in Table~\ref{tab:predictions}.

\begin{table}[t]
  \centering
  \begin{tabular}{lcc}
    & \multicolumn{2}{c}{preference \& availability}
    \\ \cmidrule(r){2-3}
    theoretical position
    & \as
    & \es
    \\ \midrule
    traditionalism
    \\
    \ \ \ \ weak restricted 
    & \lit > \glb 
    & \lit > \glb
    \\
    \ \ \ \ weak unrestricted
    & \lit > \glb > \loc 
    & \lit > \glb
    \\
    \ \ \ \ strong restricted
    & \glb > \lit 
    & \glb > \lit
    \\
    \ \ \ \ strong unrestricted
    & \glb > \loc > \lit 
    & \glb > \lit
    \\
    % conventionalism
    % \\
    % \ \ defaultism
    % & \loc >  \lit 
    % & \loc >  \lit
    % \\
    % \ \ literalism
    % & \lit >  \loc 
    % & \lit >  \loc
    % \\
    grammaticalism
    \\
    \ \ \ \ weak
    & \loc > \glb > \lit 
    & \glb > \lit, \loc
    \\
    \ \ \ \ strong
    & \glb, \loc > \lit 
    & \glb, \loc , \lit
    \\
  \end{tabular}
  \caption{Predictions of the relevant theoretical positions. Listed
    are the predicted preference relations. If a reading is not
    listed, it is predicted to be unavailable.}
  \label{tab:predictions}
\end{table}

% \begin{table}[t]
%   \centering
%   \begin{tabular}{lcccccc}
%     & \multicolumn{4}{c}{availability} 
%     & \multicolumn{2}{c}{preference}
%     \\ \cmidrule(r){2-5}  \cmidrule(r){6-7}
%     theoretical position
%     & \as-\glb
%     & \as-\loc
%     & \es-\glb
%     & \es-\loc
%     & \as
%     & \es
%     \\ \midrule
%     traditionalism
%     \\
%     \ \ weak restricted 
%     & $\checkmark$
%     & $-$
%     & $\checkmark$
%     & $-$
%     & \lit > \glb 
%     & \lit > \glb
%     \\
%     \ \ weak unrestricted
%     & $\checkmark$
%     & $\checkmark$
%     & $\checkmark$
%     & $-$
%     & \lit > \glb > \loc 
%     & \lit > \glb
%     \\
%     \ \ strong restricted
%     & $\checkmark$
%     & $-$
%     & $\checkmark$
%     & $-$
%     & \glb > \lit 
%     & \glb > \lit
%     \\
%     \ \ strong unrestricted
%     & $\checkmark$
%     & $\checkmark$
%     & $\checkmark$
%     & $-$
%     & \glb > \loc > \lit 
%     & \glb > \lit
%     \\
%     conventionalism
%     \\
%     \ \ defaultism
%     & $-$
%     & $\checkmark$
%     & $-$
%     & $\checkmark$
%     & \loc >  \lit 
%     & \loc >  \lit
%     \\
%     \ \ literalism
%     & $-$
%     & $\checkmark$
%     & $-$
%     & $\checkmark$
%     & \lit >  \loc 
%     & \lit >  \loc
%     \\
%     grammaticalism
%     \\
%     \ \ weak
%     & $\checkmark$
%     & $\checkmark$
%     & $\checkmark$
%     & $\checkmark$
%     & \loc > \glb > \lit 
%     & \glb > \lit, \loc
%     \\
%     \ \ strong
%     & $\checkmark$
%     & $\checkmark$
%     & $\checkmark$
%     & $\checkmark$
%     & \glb, \loc > \lit 
%     & \glb, \loc >  \lit
%     \\
%   \end{tabular}
%   \caption{Predictions of the relevant theoretical positions}
%   \label{tab:predictions}
% \end{table}

\subsection{The Prosodic Markedness Hypothesis}
\label{sec:inton-mark-hypoth}

Although restricted traditionalism does not of itself predict local readings to be
available at all and even unrestricted traditionalism does not predict local
readings for \es-sentences,  many authors who adhere to a traditionalist position
propose that local readings are available if the embedded scalar item
is prosodically marked by contrastive stress \citep[e.g.][]{Horn2006:The-Border-Wars,Geurts2009:Scalar-Implicat,Geurts2010:Quantity-Implic,GeurtsTielvan-Tiel2013:Embedded-Scalar}. This
is taken to be continuous with other cases where pragmatic enrichments can
take scope under logical operators if the relevant focal accent is present (\ref{bsp:Stress}).

\begin{exe}
\ex \label{bsp:Stress} This chili is not \textsc{spicy}, it's \textsc{SPICY}.
\end{exe}

\noindent The \emph{prosodic markedness hypothesis}, as we will call
it here, is then an addition to the traditionalist's explanatory
repertoire. Under this extra assumption, local readings are a marked
phenomenon, and not the product of the same process as run-of-the-mill
scalar implicatures. If supplemented with the prosodic markedness
hypothesis, traditionalists would therefore predict that local
readings are available for \as- and \es-sentences, but only if
prosodicly marked by contrastive stress.

To date, empirical investigations of this assumption are still scarce.  One
notable exception is a pilot study reported in
\citet{Frazier08}, in which participants were instructed to read
embedded and non-embedded versions of English \as-sentences. In the
embedded sentences (e.g. {\it All of the students wrote some of the
  official memos}), {\it some} was either capitalized or
non-capitalized. Prior to the experiment, participants were informed
that capitalizing corresponded to contrastive stress. The task
consisted in a forced-choice questionnaire with two given alternative
paraphrases  (such as {\it All of the students wrote some but not all of the
  official memos} vs. {\it All of the students wrote at least some of
  the official memos}). Interestingly, strengthened interpretations
occurred rather often across all sentences in this task (59 \%), but
were not affected by capitalizing.\fn{Note that, unfortunately, the
  paper does not provide percentages for the embedded conditions
  alone, but only the means across sentence types.}
\citeauthor{Frazier08} concludes that accentuation did not affect the
silent reading data, but raises the possibility that prosodic effects
may have been present at least in a subset of the items (see
footnote~1 in \citet{Frazier08}, p.~330).

Another study relevant for the present considerations comes from
\citet{SchwarzClifton2008:Strengthening-o}, examining the effects of
contrastive accentuation of \emph{or} in non-embedded sentences like
{\it Mary will invite Fred or/OR Sam to barbecue}. After listening to
these sentences, two alternative paraphrases were presented to
the participants (e.g. {\it She will invite Fred or Sam or possibly both} vs. {\it
  She will invite Fred or Sam but not both}).  The authors found a
significant impact of overtly realized contrastive accents on
implicature inferences: Besides a general advantage for
non-strengthened readings, focal accents significantly increased the
strengthened interpretation (from 16.4 \% to 28.6 \%). These results
are compatible with the "focus strengthening hypothesis" proposed by
the authors.

To sum up so far, though an effect of contrastive accentuation has
been claimed to be essential for triggering local implicatures,
experimental evidence is currently mixed, suggesting potential effects
of presentation mode (visual vs.~auditory), item-specific
characteristics, as well as the specific construction under
investigation (e.g., embedded cases vs.~non-embedded cases such as in
\citet{SchwarzClifton2008:Strengthening-o}). It is for these reasons
that an experimental investigation into the availability of readings
of \as- and \es-sentences should present sentential material
auditorily and explicitly address the prosodic markedness
hypothesis.

\section{Previous studies}
\label{sec:previous-studies}

In order to test the divergent predictions of the relevant theoretical
positions, a number of empirical studies have been carried out. We
will focus on the works of
\citet{GeurtsPouscoulous2009:Embedded-Implic},
\citet{CliftonDube2010:Embedded-Implic} and
\citet{ChemlaSpector2010:Experimental-Ev} \citep[c.f.][for related
discussion]{BenzGotzner2014:Embedded-implic}. Unfortunately, these
studies do not provide conclusive evidence about the availability and
preference of relevant readings and do not speak to prosodic
markedness effects.

\subsection{\citet{GeurtsPouscoulous2009:Embedded-Implic}}
\label{sec:Geurts-and-Pouscoulous}

\citet{GeurtsPouscoulous2009:Embedded-Implic} conducted a
picture-verification task (their Experiment~3) to find out whether
local readings of \as- and \es-sentences are available.\fn{We will
  focus here on a subset of the conditions tested by
  \citet{GeurtsPouscoulous2009:Embedded-Implic}. Actually, these
  authors did not use \es-sentences, but sentences where scalar
  \emph{some} was embedded under non-monotonic quantifier
  \emph{exactly two} (with appropriate pictures, of course). This case
  is a little more complex, but we will gloss over this here, treating
  their data, as if it had been obtained for \es-sentences.} Subjects
were presented with pictures similar to in Figure~\ref{fig:weak} and
\ref{fig:local} (pages \pageref{fig:weak} and \pageref{fig:local})
where the local reading gets a different truth-value than the literal
and the global reading. For \as-sentences, the local reading is false
for the critical picture in Figure~\ref{fig:weak} whereas the literal
and global readings are true; for \es-sentences the local reading is
true for the critical picture in Figure~\ref{fig:local}, whereas the
literal and global readings are false.

The results of \citeauthor{GeurtsPouscoulous2009:Embedded-Implic} were
strikingly unambiguous: there were \emph{no} responses indicative of a
local reading; \emph{all} of the subjects judged \as-sentences true in
a situation like in Figure~\ref{fig:weak} and \emph{all} of the
subjects judged \es-sentences false in situations like
\ref{fig:local}.

These results were criticized on theoretical grounds (e.g.,
\citet{Sauerland2010:Embedded-Implic}, but see
\citet{Ippolito2010:Embedded-Implic} for support), as well as based on
empirical observations
\citep{CliftonDube2010:Embedded-Implic,ChemlaSpector2010:Experimental-Ev}. In
the following, we will review these empirical studies briefly, as they
provide the background for our own experiment. % We will first consider
% a comment by \citet{CliftonDube2010:Embedded-Implic}, which explicitly
% distinguishes between the availability of a reading and its
% preference.

\subsection{\citet{CliftonDube2010:Embedded-Implic}}
\label{sec:clifton-dube}

In a reply to \citeauthor{GeurtsPouscoulous2009:Embedded-Implic}'s
study. \citet{CliftonDube2010:Embedded-Implic} raised the question
whether the use of a picture-verification paradigm might have been
infelicitous for testing the availability of strong readings, at least
in the case of \as-sentences. Asking whether a sentence fits a picture
might have created a bias for accepting sentences also on a weaker and
probably dispreferred reading. Clifton and Dube's study was therefore
aimed at finding out about a potential preference relation between
local and literal readings. To this end, Clifton and Dube developed a
picture-choice task where subjects were presented with an \as-sentence
and a pair of pictures, hence introducing the option to choose between
different alternatives. Subjects were asked to ``indicate which shape
is best described by the sentence'' and could choose either picture,
or options `both' and `neither.' There were two versions of this
experiment, differing in which kind of picture pairs were presented on
critical trials. In version 1, the picture pair consisted of the weak
and strong situations in Figures~\ref{fig:weak} and
\ref{fig:strong}. The response percentages observed by
\citeauthor{CliftonDube2010:Embedded-Implic} were:

\begin{center}
  \begin{tabular}{cccc}
    weak & strong & both & neither
    \\ \midrule 
 3 & 39 & 57 & 1 
  \end{tabular}
\end{center}

\noindent That the majority answer is ``both'' could be taken as
evidence that the literal reading is the preferred one. But the almost
40\% of choices for the strong situation, so
\citeauthor{CliftonDube2010:Embedded-Implic} argue, is indicative of
the availability of the local reading. In version 2 of the experiment,
the picture pair consisted of the literal and weak situations in
Figures~\ref{fig:literal-AE} and \ref{fig:weak}. In this case,
response percentages were:

\begin{center}
  \begin{tabular}{cccc}
    weak & literal & both & neither
    \\ \midrule 
    28 & 6 & 50 & 17 
  \end{tabular}
\end{center}

\noindent Again, the majority response ``both'' might speak for a
preference for the literal reading, but, as
\citeauthor{CliftonDube2010:Embedded-Implic} propose, the 17\% of
``neither'' answers in this case again suggest that the local reading
is available. In sum,
\citeauthor{CliftonDube2010:Embedded-Implic} take these results to
contradict \citeauthor{GeurtsPouscoulous2009:Embedded-Implic}'s
findings. Local readings are, after all, attested if subjects are
given a choice as to which situation they consider most fitting for an
\as-sentence.

The diverging results of
\citeauthor{GeurtsPouscoulous2009:Embedded-Implic} and
\citeauthor{CliftonDube2010:Embedded-Implic} seem to indicate that
participants' choices for readings of \as-sentences might have been
affected by the specific experimental paradigm used \citep[but
see][for critique of the latter
design]{GeurtsTielvan-Tiel2013:Embedded-Scalar}. Unfortunately,
\citet{CliftonDube2010:Embedded-Implic} did not test both universal
and non-monotonic quantifiers within the same
experiment. Additionally, it would be informative to also probe into
the availability of global readings. The study of
\citet{ChemlaSpector2010:Experimental-Ev} did both of that.

\subsection{\citet{ChemlaSpector2010:Experimental-Ev}}
\label{sec:Chemla-Spector}

\citet{ChemlaSpector2010:Experimental-Ev} also took issue with
\citeauthor{GeurtsPouscoulous2009:Embedded-Implic}'s design, arguing
that, firstly, the pictorial material used in
\citeauthor{GeurtsPouscoulous2009:Embedded-Implic}'s study was unduly
difficult; that, secondly, these pictures also may have failed to make
the local reading sufficiently relevant; and that, thirdly, the
restriction to a categorial choice (whether the sentence fits the
picture of not) may induce a bias against non-preferred readings in
cases where candidate readings stand in entailment relations
\citep[see][for this latter
criticism]{Sauerland2010:Embedded-Implic}. To meet these potential
problems, \citet{ChemlaSpector2010:Experimental-Ev} presented subjects
with pictorial material like that in Figure~\ref{fig:Chemla-Spector},
which was assumed to be easier to assess and better at highlighting
the relevance of the local readings. Albeit in a different format, the
pictures used by \citeauthor{ChemlaSpector2010:Experimental-Ev} were
instantiations of the situation types in
Figures~\ref{fig:AS-distinguishing-pics} and
\ref{fig:ES-distinguishing-pics}. Additionally, subjects were asked,
not for categorial judgements, but for \textit{graded judgements}:
subjects could freely click on a scale, as shown in
Figure~\ref{fig:Chemla-Spector}, to indicate how much they considered
a picture fitting for a given sentence \citep[see][for more on this
method]{Chemla2009:Presuppositions}.

\begin{figure}[t]
  \centering

  \subfloat[][Example of the \as-weak condition of
  \citet{ChemlaSpector2010:Experimental-Ev}]{
      \label{fig:Chemla-Spector}
  \begin{minipage}{0.45\linewidth}
    \begin{center}
      \includegraphics[scale=0.25]{pictures/paper/Chemla_Spector_2010_Critical_AE.png}\\
      \includegraphics[scale=0.25]{pictures/paper/Chemla_Spector_2010_RatingBar.png}
    \end{center}

  \end{minipage}
    }
   \begin{minipage}{6cm}
     \subfloat[][Results \as-condition]{
    \label{fig:CS-Results-as}
        \includegraphics[width=5.5cm]{pictures/paper/Chemla_Spector_2010_Results_AE.png}
  }

  \subfloat[test][Results \es-condition]{
    \label{fig:CS-Results-es}
    \includegraphics[width=5.5cm]{pictures/paper/Chemla_Spector_2010_Results_GE.png}
  }  
   \end{minipage}
        \caption{Example trial and results of
          \citeauthor{ChemlaSpector2010:Experimental-Ev}'s
          (\citeyear{ChemlaSpector2010:Experimental-Ev}) study.}
\end{figure}

% \begin{figure}[t]
%   \centering
%       \includegraphics[scale=0.25]{../pictures/paper/Chemla_Spector_2010_Critical_AE.png}\\
%       \includegraphics[scale=0.25]{../pictures/paper/Chemla_Spector_2010_RatingBar.png}
%         \caption{Example of critical trial for \as-condition by
%           \citeauthor{ChemlaSpector2010:Experimental-Ev}'s
%           (\citeyear{ChemlaSpector2010:Experimental-Ev})}
%   \label{fig:Chemla-Spector}
% \end{figure}

\citeauthor{ChemlaSpector2010:Experimental-Ev} hypothesized that the
degree to which a sentence is rated acceptable is proportional to the
number of available true readings. Observed averaged clicking
positions are shown in Figures~\ref{fig:CS-Results-as} and
\ref{fig:CS-Results-es}.
%
% \begin{figure}
%   \centering
%   \subfloat[][\as-condition]{
%     \label{fig:CS-Results-as}
%         \includegraphics[width=5.5cm]{../pictures/paper/Chemla_Spector_2010_Results_AE.png}
%   }
%   \subfloat[test][\es-condition]{
%     \label{fig:CS-Results-es}
%     \includegraphics[width=5.5cm]{../pictures/paper/Chemla_Spector_2010_Results_GE.png}
%   }
%   \caption{Results from
%     \citeauthor{ChemlaSpector2010:Experimental-Ev}'s
%     (\citeyear{ChemlaSpector2010:Experimental-Ev}) study}
%   \label{fig:Chemla-Spector-Results}
% \end{figure}
%
According to \citeauthor{ChemlaSpector2010:Experimental-Ev}, the
crucial piece of evidence for the availability of local readings for
\as-sentences is that these sentences yielded higher graded
acceptability scores for the strong situation than for the weak
situation (although these differ only with respect to the truth value
of the local reading). Evidence for the availability
of the local reading for \es-sentences comes from the difference
between the local and the literal situation. Strikingly, \es-sentences
received an average 73\% degree of acceptability in the local
situation although the literal and global readings are false in this
case.

\subsection{Reflection}
\label{sec:local-read-categ}

Summing up, three experimental studies have presented diverging pieces
of evidence. % \citet{GeurtsPouscoulous2009:Embedded-Implic} did not
% observe any truth value-judgments that would indicate local readings
% in a picture verification task and thus conclude that these readings
% are unavailable. \citet{CliftonDube2010:Embedded-Implic} and
% \citet{ChemlaSpector2010:Experimental-Ev} employed different designs
% intended to be sensitive enough to reveal local readings. Both studies
% found putative evidence in support of local readings.
Differences in results are likely due to differences in experimental
design, in particular due to (i) the type of elicited judgement and
(ii) the possibility of conflating pictorial effects.

\paragraph{Judgement Type.} We could draw a distinction between
categorical truth-value judgments and other potentially more sensitive
measures. The former are possibly not sensitive enough to reveal
dispreferred readings with small samples, but the latter may be. We
could then hypothesize along with
\citeauthor{CliftonDube2010:Embedded-Implic} and
\citeauthor{ChemlaSpector2010:Experimental-Ev} that local readings are
available, but so dispreferred that they do not affect categorical
truth-value judgments.

On the other hand, \citet{Crain1998} argue in favor of truth-value
judgments as a means of detecting dispreferred readings. Consequently,
traditionalists could reply that the putative evidence for local
readings in allegedly more sensitive tasks might reflect something
other than (strongly) truth-relevant speaker-intended meaning
enrichments. Hence no case can be made for lexicalism or
grammaticalism on the basis of these data \citep[see][for arguments
along these
lines]{GeurtsTielvan-Tiel2013:Embedded-Scalar,Tielvan-Tiel2014:Quantity-Matter,Tielvan-Tiel2012:Embedded-Scalar}.

To resolve this issue, we would ideally like to have a design that
elicits categorical truth-value judgements, but is still possibly
sensitive enough to detect dispreferred readings.

\paragraph{Pictorial Effects.} It is often possible to present
different numbers of links between icons in pictures like in
Figures~\ref{fig:AS-distinguishing-pics}
and~\ref{fig:ES-distinguishing-pics}, without changing the
truth-values of relevant readings. Indeed, the pictorial material used
in previous studies differed in this respect, and that might explain
some of the differences in results. If that is so, then we cannot
ascribe the effects reported by
\citet{CliftonDube2010:Embedded-Implic} and
\citet{ChemlaSpector2010:Experimental-Ev} to the availability of local
readings with certainty. \citet{ChemlaSpector2010:Experimental-Ev}
also acknowledge the possibility that the \emph{typicality} of a
picture with respect to some meaning may affect graded truth-value
judgments. They show that graded judgments of \as-sentences differ
significantly for different pictures that agree on the truth value of
relevant readings but differ with respect to the amount of connections
between icons. \citeauthor{ChemlaSpector2010:Experimental-Ev} suggest
that typicality of the pictorial material can account for these
differences, but submit that this does not explain away the high
acceptance of pictures like Figure \ref{fig:weak}.

The latter point is disputed by
\citet{Tielvan-Tiel2012:Embedded-Scalar}, who demonstrates that a huge
chunk of variance in the responses to \as-sentences can be explained
as typicality effects, dispensing the need to appeal to the
distribution of different readings. In support of this idea,
\citet{Tielvan-Tiel2012:Embedded-Scalar} elicited what he calls the
typicality structures associated with the quantifiers {\it all} and
{\it some} \citep[as done also
by][]{DegenTanenhaus2011:Making-Inferenc} and predicted the judgments
of \as-sentences obtained by
\citeauthor{ChemlaSpector2010:Experimental-Ev} based on these
data. Thereby, he obtained an excellent model fit.

We are generally sympathetic towards
\citeauthor{Tielvan-Tiel2012:Embedded-Scalar}'s innovative line of
reinterpretation of the data, but note, as
\citet{ChemlaSpector2010:Experimental-Ev} already observed, that his
typicality-based explanation does not extend to \es-sentences in an
obvious way. Moreover, typicality itself is not a satisfactory
primitive in a putative explanation of the use of quantifiers, but
rather something that needs explanation itself \citep[c.f.][for
critical reflection on typicality-based
explanations]{Cummins2014:Typicality-made}. Indeed, it is possible to
explain typicality judgements as the outcome of Gricean speaker
preferences for informative descriptions, just like those that
rationalize quantity implicatures \citep{Franke2014:Typical-use-of-}.
For these reasons, we believe that typicality and other pictorial
effects need to be taken seriously in the experimental design, but do
not necessarily give a satisfactory account of the observed variation
all by themselves.

\paragraph{Upshot.} Taken together, we would ideally like a design
that evokes categorical truth value judgments in a way that is as
independent as possible of potentially conflating properties of the
pictorial material and that is nonetheless sensitive enough to detect
local readings even if they are dispreferred. The sensitivity to
dispreferred readings should be controlled for. On top of this, it
would be desirable to obtain further information about the relative
preferences among all putative readings, so as to be able to decide
between theoretical positions, as outlined in
Section~\ref{sec:theories-predictions}. % Neither of the previous
% studies gives us that, either because they simply have not
% investigated all the relevant comparisons
% (\citet{GeurtsPouscoulous2009:Embedded-Implic},
% \citet{CliftonDube2010:Embedded-Implic}), or because they cannot
% safely derive the required information for danger of conflating
% preferences for readings with preferences for pictures
% (\citet{ChemlaSpector2010:Experimental-Ev}).

Finally, we need to control for ``silent prosody.'' Several studies
suggest that numerous factors might influence accent placement and
prosodic phrasing even while reading, amongst them default
accentuation, constituent length, rhythmic phenomena, and individual
variation
\citep[e.g.][]{Bader98,Fodor98,Steinhauer01,Fodor02,Augurzky08,BreenWatson2011:Intonational-ph,Kentner12}. If
we adopt the prosodic markedness hypothesis described in
Section~\ref{sec:inton-mark-hypoth}, as many traditionalists do, we
predict that the availability of a local reading hinges on the
realization of contrastive stress on the scalar item
\citep[e.g.][]{Horn2006:The-Border-Wars,Geurts2009:Scalar-Implicat,Geurts2010:Quantity-Implic,Tielvan-Tiel2012:Embedded-Scalar,GeurtsTielvan-Tiel2013:Embedded-Scalar}.
But if prosody can have this role, it is necessary to control for
silent accent placement. Ideally, therefore, we should present
sentences auditorily and systematically manipulate contrastive stress.

\section{An Incremental Verification Task}
\label{sec:exp}

\subsection{Design}
\label{sec:design}

\paragraph{General Idea.} In order to test the availability and
preferences of different readings of \as- and \es-sentences in German,
we were looking for a way to unambiguously map responses from a
picture verification task to specific readings. We were particularly
interested in the question whether local readings would be observed at
all, and whether their presence or absence would be modulated by a
contrastive accent on the scalar item. Moreover, we were interested in
obtaining information about the relative preferences of the different
readings. In addition, we wanted to minimize potential picture
typicality effects. To achieve this (see
Section~\ref{sec:discussion-1} for discussion), we used an
\mymark{incremental verification task} (\acro{ivt}), which is a
modified version of picture verification
\citep[see][]{Conroy2008}. The general idea is that subjects are
requested to judge sentence material with respect to pictures that do
not necessarily contain all the information relevant for judging a
certain reading true or false. In that case, participants can demand
that more information be revealed. Participants are instructed to make
a truth-value judgment as soon as they are able to. When they do, the
trial ends.

\paragraph{Target Conditions.} We present sequences of pictures
depicting a set of four identical central elements (e.g., letters),
which could be connected to surrounding elements (e.g., triangles), as
shown in Figures~\ref{fig:exseqAS}--\ref{fig:exec}. Initially,
any potential connections between central and surrounding elements
were covered by dark gray color (see
Figure~\ref{fig:exseqAS1}). Sentences to be judged were presented
auditorily, and participants were asked to uncover the picture until
they felt able to give a truth value judgment. Three options were
available for participants at each step: (i) judge the sentence as
true, (ii) judge the sentence as false, (iii) demand more information
(unavailable when the picture was fully uncovered). Trials ended when
a truth-value judgement was made. Importantly, each of the three
potentially available readings corresponded to a specific step in the
uncovering process where the truth value of that reading (and only of
that reading) could be assessed for the first time. We refer to this
step in the sequence as the {\it critical position} of a given
reading. The critical position and the corresponding truth-value
judgment differed between \as- and \es-sentences, as described
presently. (This is, partly, because of the logical dependencies
between readings, and, partly, also to rule out positional biases.)

\begin{figure}[]
	\centering
	\subfloat[][Step 1]{ 
		\fbox{\includegraphics[width=3.5cm]{pictures/ae_01_1.pdf}}
	    \label{fig:exseqAS1}
	}
	\subfloat[][Step 2]{
		\fbox{\includegraphics[width=3.5cm]{pictures/ae_01_2.pdf}}
	    \label{fig:exseqAS2}
	}
	\subfloat[][Step 3 (\lit true)]{ 
		\fbox{\includegraphics[width=3.5cm]{pictures/ae_01_3.pdf}}
	    \label{fig:exseqAS3}
	}
        \\
	\subfloat[][Step 4]{ 
		\fbox{\includegraphics[width=3.5cm]{pictures/ae_01_4.pdf}}
	    \label{fig:exseqAS4}
	}
	\subfloat[][Step 5 (\glb true)]{
		\fbox{\includegraphics[width=3.5cm]{pictures/ae_01_5.pdf}}
	    \label{fig:exseqAS5}
	}
	\subfloat[][Step 6]{ 
		\fbox{\includegraphics[width=3.5cm]{pictures/ae_01_6.pdf}}
	    \label{fig:exseqAS6}
	}
        \\
	\subfloat[][Step 7 (\loc false)]{ 
		\fbox{\includegraphics[width=3.5cm]{pictures/ae_01_7.pdf}}
	    \label{fig:exseqAS7}
	} 
                \hspace*{7.85cm} 
	\caption[]{Example picture sequence for \as-sentences.}
	\label{fig:exseqAS}
\end{figure}

Consider the German \as-sentence in (\ref{ex:as}), included in our study.
\begin{exe}
\ex \gll \mymark{Alle} diese Briefe sind mit \mymark{einigen} ihrer Dreiecke
  verbunden.\label{ex:as}\\
All these letters are with some their triangles connected.\\
\trans All of these letters are connected to some of their triangles.
\end{exe}
Figure~\ref{fig:exseqAS} shows the corresponding picture sequence. The
critical positions are on step 3, 5 and 7. Pictures that were
interspersed between these positions served as
spillover-pictures. These were used in order to counter potentially
delayed judgments and they additionally served as distractor
items. Prior to step 3 there is not enough information to judge any
candiate reading true or false. The situation at step 3 in
Figure~\ref{fig:exseqAS3} is true under a literal reading, while the
local and global readings cannot be evaluated yet. On step 5 in
Figure~\ref{fig:exseqAS5} the literal reading is still true, but now
also the global reading can be confirmed. Finally, decisions
concerning the local reading are possible as soon as all connections
have been uncovered at step 7 in Figure~\ref{fig:exseqAS7}. Note that
the situation at step 7 is \emph{in}compatible with a local
reading. This enables us to separate local readings, which would yield
\emph{false} judgements at this position, from literal and global
readings, which would yield \emph{true} judgements. In sum,
\emph{true} or \emph{false} answers on particular positions in the
incrementally revealed picture sequence can be mapped uniquely to
candidate readings. All other \emph{true} or \emph{false} answers were
counted as errors.

Due to the non-linear entailment order of readings described in
Section~\ref{sec:get-know-your}, \es-sentences require a slightly
different sequence of unfolding, which yields a different order in
which truth-value judgments can be made.
%
\begin{figure}[t]
	\centering
	\subfloat[][Step 1]{ 
		\fbox{\includegraphics[width=3.5cm]{pictures/ge_01_1.pdf}}
	    \label{fig:exseqES1}
	}
	\subfloat[][Step 2 (\glb false)]{
		\fbox{\includegraphics[width=3.5cm]{pictures/ge_01_2.pdf}}
	    \label{fig:exseqES2}
	}
	\subfloat[][Step 3 (\lit false)]{ 
		\fbox{\includegraphics[width=3.5cm]{pictures/ge_01_3.pdf}}
	    \label{fig:exseqES3}
	}
        \\
	\subfloat[][Step 4]{ 
		\fbox{\includegraphics[width=3.5cm]{pictures/ge_01_4.pdf}}
	    \label{fig:exseqES4}
	}
	\subfloat[][Step 5 (\loc true)]{
		\fbox{\includegraphics[width=3.5cm]{pictures/ge_01_5.pdf}}
	    \label{fig:exseqES5}
	}
                \hspace*{3.9cm} 
	\caption[]{Example sequence for \es-sentences.}
	\label{fig:exseqES}
\end{figure}
%
For a sentence like (\ref{ex:es}), we obtain an unambiguous mapping
between truth judgements and readings with the sequence in
\ref{fig:exseqES}.
\begin{exe}
\ex \label{ex:es} \gll \mymark{Genau} eine der Glocken ist mit
  \mymark{einigen} ihrer Halbkreise verbunden.\\ 
  Exactly one of-the bells is with some its semicircles connected.\\
  \trans Exactly one bell is connected to some of its semicircless.
\end{exe}
The critical positions are steps 2, 3 and 5. At step 2 in
Figure~\ref{fig:exseqES2}, a global reading would yield a \emph{false}
judgement. The literal reading can be evaluated at step 3 in
Figure~\ref{fig:exseqES3}, where it would trigger a \emph{false}
judgement. Finally, confirming \ref{fig:exseqES5} by a \emph{true}
response indicates a local reading. Note that global and literal
readings again differ from local readings with respect to their truth
values.

\paragraph{Positional controls.} To make sure that subjects understood
the task, i.e., gave truth-value judgements at the first possible
position in a sequence, we included a set of control conditions. These
also controlled for response biases. For each of the three readings in
the \as- and \es-conditions, we constructed an unambiguous control
sentence as in (\ref{bsp:controls-as}) and (\ref{bsp:controls-es}),
requiring the same judgement at the identical position in the same
sequence used for the respective targets. For instance, example
(\ref{bsp:controls-as-1}) requires a \emph{true}-response analogous to
the \as-sentence in (\ref{ex:as}) under its literal reading at step 3
in Figure~\ref{fig:exseqAS3}, (\ref{bsp:controls-as-2}) corresponds to
the \as-sentence under its global reading and
(\ref{bsp:controls-as-3}) corresponds to its local reading. With
regard to the \es-sentences, controls like (\ref{bsp:controls-es-1})
correspond to the global reading, (\ref{bsp:controls-es-2}) to the
literal and (\ref{bsp:controls-es-3}) to the local reading. If,
independently of sentence meaning, there was any bias to respond in a
certain way at any point in the sequences, such as to preferably
unravel the whole picture, this should affect control sentences to the
same degree as it affects target conditions.

\begin{exe}
  \ex \label{bsp:controls-as}
    \begin{xlist}
\ex \label{bsp:controls-as-1} \gll Alle Briefe sind mit mindestens drei ihrer Dreiecke verbunden.\\
  All letters are with at-least three their triangles connected.\\
  \trans All letters are connected to at least three of their triangles.
\ex \label{bsp:controls-as-2} \gll Mindestens ein Brief ist mit genau f\"unf seiner Dreiecke verbunden.\\
  At-least one letter is with exactly five his triangles connected.\\
  \trans At least one letter is connected with exactly five of its triangles.
\ex \label{bsp:controls-as-3} \gll Jeder Brief ist mit mindestens vier seiner Dreiecke verbunden.\\
  Every letter is with at-least four his triangles connected.\\
  \trans Every letter is connected to at least four of its triangles.
\end{xlist}
\end{exe}

\begin{exe}
\ex \label{bsp:controls-es}
  \begin{xlist}
\ex \label{bsp:controls-es-1} \gll Alle Glocken sind mit weniger als
  vier ihrer Halbkreise verbunden. \\ 
All bells are with fewer than four their semicircles connected.\\
\trans All bells are connected with fewer than four of their semicircles.  
\ex \label{bsp:controls-es-2} \gll Alle Glocken sind mit allen ihren Halbkreisen verbunden.\\
All bells are with all their semicircles connected.\\
\trans All bells are connected to all of their semicircles. 
\ex \label{bsp:controls-es-3} \gll Mindestens drei Glocken sind mit
  allen ihren Halbkreisen verbunden.\\ 
  At-least three bells are with all  their semicircles connected.\\
  \trans At least three bells are connectd with all of their semicircles.
\end{xlist}
\end{exe}

\paragraph{Preference-related controls.} In order to test whether the
order of critical positions for different readings within a sequence
had an influence on responses, we included a second type of control
conditions.  These controls served as a test whether the incremental
verification task could possibly detect at least ordinal preference
relations among multiple candidate readings. Note that for our target
sentences, the logical entailment relations between readings always
require local readings to be evaluated at the end of a sequence. It
could thus in principle be possible that subjects never reach this
point, because they give truth-value judgements earlier, thereby
ending the trial. In that case it would be unclear whether these
decisions had been affected by a general unavailability of local
readings or by the fact that one of the earlier presented readings is
the preferred one. By including globally ambiguous structures with a
known preference over available readings, we thus tested whether
participants in our task occasionally choose dispreferred readings,
even if these were available only at a critical position following the
preferred readings. Finally, these conditions also controlled whether 
prosodic information can, in principle, shift answer patterns in the 
present task.

We therefore included globally ambiguous \emph{attachment ambiguities} 
as in (\ref{bsp:target-related}), which have been shown to exhibit 
interpretive preferences and can be disambiguated by prosodic information.

\begin{exe}
\ex \gll Der Brief ist mit Kreisen und Vierecken mit Sonnen
  verbunden. \label{bsp:target-related}\\
The letter is with circles and squares with suns connected.\\
The letter is connected with circles and squares with suns.
\begin{xlist}
  \ex \label{bsp:target-related-LC} The letter is connected with squares containing suns, and it is
    also connected with circles. \hfill{(\lc)}
  \ex \label{bsp:target-related-EC} The letter is connected with circles and squares, both of which
    are containing suns.  \hfill{(\ec)} 
\end{xlist}

\end{exe}
In attachment ambiguities involving post-nominal modification, 
an adjunct like a relative clause or a prepositional phrase 
(\acro{pp}) can be attached to one of two preceding hosts. 
For instance, in (\ref{bsp:target-related}), the
\acro{pp} \emph{with suns} can be attached to the preceding noun
\emph{squares}, resulting in the so-called \emph{late-closure} (\lc)
reading ((\ref{bsp:target-related-LC}). Alternatively, the \acro{pp}
can be attached to the whole conjunctive \acro{np} \emph{circles and
  squares}, corresponding to the \emph{early-closure} (\ec) reading
(\ref{bsp:target-related-EC}).  \lc-readings have been generally 
observed to be preferred over \ec-readings \citep[e.g.][]{Fodor98}
for an overview of different languages; but see \citet{CliftonCarlsonFrazier2002}
for counterevidence). Crucially, an \lc-preference for \acro{pp}-attachment 
structures has also been attested for German \citep[][]{KoniecznyHemforth2000}.

For each sentence, two sequences were designed: one in which the
preferred \lc-reading could be judged first (Figure~\ref{fig:exlc}),
%
\begin{figure}[]
	\centering
	\subfloat[][Step 2]{ 
		\fbox{\includegraphics[width=5cm]{pictures/lc_01_2.pdf}}
	    \label{fig:exlc2}
	}
	\subfloat[][Step 3 (\lc true)]{
		\fbox{\includegraphics[width=5cm]{pictures/lc_01_3.pdf}}
	    \label{fig:exlc3}
	}
        \\
	\subfloat[][Step 4]{
		\fbox{\includegraphics[width=5cm]{pictures/lc_01_4.pdf}}
	    \label{fig:exlc4}
	}
	\subfloat[][Step 5 (\lc true)]{
		\fbox{\includegraphics[width=5cm]{pictures/lc_01_5.pdf}}
	    \label{fig:exlc5}
	}
        \\
	\subfloat[][Step 6 (\ec false)]{
		\fbox{\includegraphics[width=5cm]{pictures/lc_01_6.pdf}}
	    \label{fig:exlc6}
	}
	\subfloat[][Step 7 (\lc true)]{
		\fbox{\includegraphics[width=5cm]{pictures/lc_01_7.pdf}}
	    \label{fig:exlc7}
	}
	\caption[]{Example of an \lc-sequence for sentence
          (\ref{bsp:target-related}) where the \lc-reading
          (\ref{bsp:target-related-LC}) can be judged first. The first
          step where all connections are covered is omitted.}
	\label{fig:exlc}
\end{figure}
%
and another in which the dispreferred \ec-reading could be judged
first (Figure~\ref{fig:exec}).
%
\begin{figure}[]
	\centering
	\subfloat[][Step 2]{ 
		\fbox{\includegraphics[width=5cm]{pictures/ec_01_2.pdf}}
	    \label{fig:exec1}
	}
	\subfloat[][Step 3 (\ec false)]{
		\fbox{\includegraphics[width=5cm]{pictures/ec_01_3.pdf}}
	    \label{fig:exec2}
	}
        \\
	\subfloat[][Step 4]{
		\fbox{\includegraphics[width=5cm]{pictures/ec_01_4.pdf}}
	    \label{fig:exec2}
	}
	\subfloat[][Step 5 (\lc true)]{
		\fbox{\includegraphics[width=5cm]{pictures/ec_01_5.pdf}}
	    \label{fig:exec2}
	}
	\caption[]{Example of an \ec-sequence for sentence
          (\ref{bsp:target-related}) where the \ec-reading
          (\ref{bsp:target-related-EC}) can be judged first. The first
          step where all connections are covered is omitted.}
	\label{fig:exec}
\end{figure}
%
We also tested structurally disambiguated versions like
(\ref{bsp:target-related2}), one for each sentence type, on each
sequence type for proper comparison.

\begin{exe}

\ex \gll Der Brief ist mit Kreisen, die Sonnen beinhalten, und Vierecken
  verbunden. \label{bsp:target-related2}\\
The letter is with circles, which suns contain and squares connected.\\
The letter is connected with circles containing suns, and with squares.

\end{exe}

\paragraph{Testing Effects of prosody.} Whereas contrastive accents
are usually assumed to be realized by a special contour in English
(i.e. L+H*, see \cite{Pierrehumbert90}), the exact phonological
classification of contrast in German is still under debate
\citep[see][]{Uhmann91,Fery93,Grabe98,Toepel06,Sudhoff10}.  Still,
all of these approaches have in common that from an acoustic
perceptive, prosodic prominence of contrastively accented constituents
is realized by

\begin{enumerate}[(i)]
\item a higher F0 maximum or a higher pitch range (difference 
between minimal and maximal F0 values), and
\item a longer duration of the accented element as opposed to its 
non-accented counterpart.
\end{enumerate}

\noindent Our sentences were read by a phonetically trained female
native speaker of German familiar with the concept of contrastive
focus. Both \as-sentences and \es-sentences were recorded in two
versions each. In the accented version, a contrastive stress was
placed on {\it einige}. The second version had neutral prosody.  If
accentuation is the driving force for local readings, we would expect
a higher proportion of local readings in the former than in the latter
version of the sentences.

Preference-related controls also differed prosodically. Modifier
attachment ambiguities have been shown to be sensitive to differences
in prosodic phrasing. Though there has been discussion as to whether
speakers reliably \textit{produce} disambiguating cues in such
constructions
\citep[e.g.][]{Allbritton96,Kraljik05,Snedeker03,Schafer00}, overt
prosodic boundaries are generally acknowledged to guide interpretive
processes of attachment ambiguities in \textit{comprehension}
\citep[e.g.][]{Augurzky06,Beach1991,CliftonCarlsonFrazier2002,KjelgaardSpeer1999,
  SchaferSpeerWarrenWhite2000,Steinhauer99}.  Specifically, it has
been shown that prosodically separating a modifier from the directly
preceding material supports an \ec-reading, whereas separating the two
potential attachment sites is supportive of an \lc-reading
\citep[e.g.][]{CliftonCarlsonFrazier2002,Fodor02,Jun2003}.  Thus, a
phrase boundary between {\it squares} and {\it with} in
(\ref{bsp:target-related}) supports an \ec-reading, whereas a prosodic
phrase boundary following {\it circles} supports an \lc-reading.

To test whether prosodic information is taken into account in our
task, we presented sentences in both of these prosodic variants. In
addition, in order to test prosody-independent preferences, we also
included a neutral version without any pronounced boundary. Further,
we also presented unambiguous sentences corresponding to the
\ec-reading. As in the case of the \as- and \es-sentences,
the unambiguous counterparts served as a baseline to control for
response biases that are independent of sentence meaning.

\subsection{Methods}
\label{sec:materials}

\subsubsection{Procedure}
\label{sec:procedure}

Each experimental trial proceeded as described above. The sentence was
presented using active PC loudspeakers and the picture materials were
presented on a computer screen. Participants responded using the
keyboard. By pressing one of four buttons they could (i) listen to the
sentence (up to three times), (ii) request additional pictorial
information, (iii) respond ``yes, fits'' or (iv) ``no, does not fit''
(German: ``passt'' or ``passt nicht''). At the beginning of an
experimental session participants received written instructions on
their task. Then they completed an interactive training consisting of
15 trials in which they were familiarized with the task.  After the
training the actual experiment started. Participants were tested
individually in a silent room. Each session was divided into two
blocks and participants were told that they could take breaks between
blocks. In total, an experimental session took about one hour. At the
end of the session participants received 8 euros compensation.

\subsubsection{Materials}

\paragraph{Target sentences and positional controls.}

We constructed a set of 15 items for German \as- and \es-sentences,
respectively, analogous to the examples in (\ref{ex:as}) and
(\ref{ex:es}). Sentences were introduced by a subject {\acro{dp}},
including the quantifiers {\it alle diese} ("all of these") or {\it
  genau eine(r) der} ("exactly one of these"), as well as the head
noun which denoted different icons, such as letters or bells. Subject \acros{DP}
were followed by an auxiliary, and a \acro{pp} containing {\it
  einige}, a possessive pronoun, and a noun denoting a geometrical
object, e.g., {\it einigen seiner Dreiecke/Quadrate} ("some of its
triangles/squares"). The possessive pronoun was used in order to fix
relative quantifier scope to surface scope, thus ensuring that {\it
  einige} has embedded status. The last word was the main verb {\it
  verbunden} ("connected").  Each sentence was recorded in a stressed
and an unstressed version of scalar {\it einige}. As described above,
three control sentences were created for each experimental item,
corresponding to the critical positions in the course of uncovering
the accompanying sequences.  The positional controls were recorded
with neutral prosody.  Items were evenly distributed across 5 lists
using a Latin square design.

\paragraph{Preference-related control sentences.}
For controlling preferences, 30 sentences that were ambiguous between
a \lc and an \ec-reading, as well as 30 of their disambiguated
counterparts were constructed as described above. In these sentences,
subject \acros{dp} were always denoting icons and were followed by an
auxiliary. In the ambiguous sentences, the auxiliary was followed by
two {\it with}-\acros{pp} (German: ``mit'').  The first of these
\acros{pp} consisted in a coordination of two nouns denoting
geometrical objects, whereas the nouns in the second \acro{pp} denoted
icons. In contrast to these sentences, the unambiguous sentences
contained only one \acro{pp} with a conjunction the first part of
which was modified by a relative clause.

\paragraph{Acoustic properties.}
% A phonetically trained female native speaker of German was instructed
% to realize two prosodic versions of each target sentence, and three
% versions of the ambiguous preference-related controls. In addition to
% these sentences, the unambiguous preference-related controls, the
% positional controls and a set of unrelated fillers were also
% recorded. For these constructions, the speaker was instructed to
% realize prosodic contours as neutral and natural as possible.

For all target sentences (\as and \es), the determiner \emph{einige} was
produced with a contrastive pitch accent as well as with a neutral
accent. % Contrastive accents in German are realized by an increase in F0 (range)
% as well as by a longer duration (see Section~\ref{sec:design} above). 
A set of 15 
experimental items was recorded for each condition (\as vs.~\es, \emph{accented} 
vs.~\emph{unaccented}), resulting in a total number of 60 target sentences.

In contrast to the accent manipulation, the ambiguous
preference-related controls differed with respect to prosodic
phrasing. Prosodic phrase boundaries in German are realized by a rise
in F0 as well as by a durational increase on the final part of the
constituent preceding the boundary (prefinal lengthening) plus an
optional pause \citep[e.g.][]{Vaissiere83,Fery93}.  Boundaries for
these control sentences were either realized at the position
separating the second \acro{pp} from the preceding material
(\emph{late boundary}, corresponding to an \ec-reading) or directly
preceding the second conjunct in the first \acro{pp} (\emph{early
  boundary}, corresponding to an \lc-reading). As the prosodic
realization of the targets involved the comparison between an accented
and a neutral variant, we also included a third version of the
ambiguous preference-related controls without any pronounced
boundaries. For each prosodic variant, a set of 30 items was read,
yielding 90 preference-related controls.

Altogether a total number of 300 sentences consisting of 60 target
sentences, 90 ambiguous preference-related controls, 30 unambiguous
preference-related controls, 90 positional control sentences and 30
unrelated fillers was recorded. The session was recorded in an
acoustically shielded booth (44.1\,kHz sampling rate, 16 bit amplitude
resolution).

Before entering the judgment task, experimental items and
preference-related controls were analyzed with respect to their acoustic
properties. As both accented elements as well as prosodic boundaries
were expected to differ with respect to their F0 and/or durational
properties, we calculated durational values as well as difference
values between minimal and maximal F0 for each word. Since targets
slightly differed with respect to the total number of words as well as
with respect to certain lexical properties (i.e., \emph{seinen}
vs.~\emph{ihren}, we considered the following analysis regions:

\begin{exe}
  \ex
    \begin{xlist}
      \ex $|_{\text{R}1}$~Alle  	\ $|_{\text{R}2}$~diese 
      \ $|_{\text{R}3}$~$\acro{np}_1$  
      \ $|_{\text{R}4}$~sind  \ $|_{\text{R}5}$~mit  \
      $|_{\text{R}6}$~einigen  \ $|_{\text{R}7}$~ihrer 
       \ $|_{\text{R}8}$~{$\acro{np}_2$}  \ $|_{\text{R}9}$~verbunden.
    \ex       $|_{\text{R}1}$~Genau einer  	\ $|_{\text{R}2}$~der 
      \ $|_{\text{R}3}$~$\acro{np}_1$  
      \ $|_{\text{R}4}$~ist  \ $|_{\text{R}5}$~mit  \
      $|_{\text{R}6}$~einigen  \ $|_{\text{R}7}$~seiner 
       \ $|_{\text{R}8}$~{$\acro{np}_2$}  \ $|_{\text{R}9}$~verbunden.
    \end{xlist}
    % \begin{tabbing}
    %   $|_{\text{R}1}$ Genau einer \= 	\ $|_{\text{R}2}$ diese \=
    %   \ $|_{\text{R}3}$ \acro{np}$_1$ \= 
    %   \ $|_{\text{R}4}$ sind \= \ $|_{\text{R}5}$ mit \= \
    %   $|_{\text{R}6}$ einigen \= \ $|_{\text{R}7}$ seiner 
    %   \= \ $|_{\text{R}8}$  \acro{np}$_2$ \= \ $|_{\text{R}9}$
    %   verbunden. \kill
    %   $|_{\text{R}1}$ Alle \> 	\ $|_{\text{R}2}$ dieser \>
    %   \ $|_{\text{R}3}$ \acro{np}$_1$ \> 
    %   \ $|_{\text{R}4}$ sind \> \ $|_{\text{R}5}$ mit \> \
    %   $|_{\text{R}6}$ einigen \> \ $|_{\text{R}7}$ ihrer 
    %   \> \ $|_{\text{R}8}$  \acro{np}$_1$ \> \ $|_{\text{R}9}$
    %   verbunden. \\
    %   $|_{\text{R}1}$ Genau einer \> 	\ $|_{\text{R}2}$ der \>
    %   \ $|_{\text{R}3}$ \acro{np}$_1$ \> 
    %   \ $|_{\text{R}4}$ ist \> \ $|_{\text{R}5}$ mit \> \
    %   $|_{\text{R}6}$ einigen \> \ $|_{\text{R}7}$ seiner 
    %   \> \ $|_{\text{R}8}$  \acro{np}$_1$ \> \ $|_{\text{R}9}$ verbunden.
    % \end{tabbing}
\end{exe}

\noindent Note that differences between Regions 1, 2, 4 and 7 can be expected
due to lexical differences between the \as and \es-conditions. As
preference-related controls did not differ with respect to their lexical
properties, we carried out word-by-word analyses for these
conditions. Durational values included the respective word plus any
following silent interval. We did not include the
disambiguated fillers in these analyses as they involved very
different sentence types (i.e., constructions involving prepositional
phrases vs.~relative clauses).

\begin{exe}
  \ex $|_{\text{R}1}$ \acro{D}  	\ $|_{\text{R}2}$ $\acro{np}_1$
      \ $|_{\text{R}3}$ ist 
      \ $|_{\text{R}4}$ mit  \ $|_{\text{R}5}$ $\acro{np}_2$   \
      $|_{\text{R}6}$ und  \ $|_{\text{R}7}$ $\acro{np}_3$ 
       \ $|_{\text{R}8}$  mit  \ $|_{\text{R}9}$ $\acro{np}_4$ \ $|_{\text{R}10}$~verbunden.
\end{exe}

For the durational analyses, constituents were automatically labeled
by the \emph{Aligner} tool \citep{Rapp98}, and the
obtained values were manually corrected afterwards. For the targets,
two-factorial \acros{anova} with the factors \acro{Quantifier} (\emph{all}
vs.~\emph{exactly one}) and \acro{Prosody} (accented vs.~unaccented) were
carried out. For the preference-related controls, we carried out
one-factorial \acros{anova} with the factor \acro{Prosody} (early
boundary, late boundary, neutral prosody). F0 values were extracted by
means of special Praat scripts
(\url{http://www.fon.hum.uva.nl/praat/}). For the present analyses,
differences between minimal and maximal F0 values for each region or
word were calculated. Again, two-factorial \acros{anova} were carried
out for statistical comparison of the target sentences, and one-factorial
\acros{anova} were carried out for preference-related controls.

\paragraph{Targets.} 
Differences between maximal and minimal F0 values for each of the
single words in the sentence are depicted in
Figure~\ref{fig:minmax_targets}. As expected, accented
determiners showed a larger F0 range compared to unaccented versions,
an effect which was confirmed by statistical analyses (effects on
the determiner: \acro{Prosody}: \emph{F}=94.8; \emph{p}<.001); 
Interaction of \acro{Quantifier} and \acro{Prosody}: \emph{F}=13.7; 
\emph{p}<.01). The observed interaction indicates that these differences
are even more pronounced in the \as-condition.

\begin{figure}[]
	\centering
	\subfloat[][\as-sentences]{
          \begin{tikzpicture}[scale=0.75]
  \begin{axis}[
    cycle list name = mylist,
    legend pos = north west,
    ymajorgrids,
    ymax = 250,
    y = 0.5,
    x tick label style = {rotate = 50, anchor = east},
    xtick = data,
    xticklabels = {alle,diese,N1,sind,mit,einigen,ihrer,N2,verbunden}
    ]
    \addplot+[line width = 1.3pt] coordinates {
      (0 ,17.6)
      (1 ,27.4)
      (2 ,82.6)
      (3 ,32.6)
      (4 ,62.4)
      (5 ,211.3333333)
      (6 ,57.6)
      (7 ,36.66666667)
      (8 ,47.13333333)};

  \addplot+[line width = 1.3pt] coordinates {
        (0 ,19.6        )
        (1 ,36.46666667 )
        (2 ,115.1333333 )
        (3 ,60.6        )
        (4 ,46          )
        (5 ,69.86666667 )
        (6 ,46.66666667)
        (7 ,71.4)
        (8 ,61.46666667)};

      \legend{accented, neutral}
  \end{axis}
\end{tikzpicture}
		% \fbox{\includegraphics[width=5.5cm]{pictures/paper/minmax_ae.png}}
	    \label{fig:minmax_ae}
	}
	\subfloat[][\es-sentences]{
\begin{tikzpicture}[scale=0.75]
  \begin{axis}[
    cycle list name = mylist,
    legend pos = north west,
    ymajorgrids,
    ymax = 250,
    y = 0.5,
    x tick label style = {rotate = 50, anchor = east},
    xtick = data,
    xticklabels = {genau,einer,dieser,N1,ist,mit,einigen,seiner,N2,verbunden}
    ]
    \addplot+[line width = 1.3pt] coordinates {
        (     0 , 59.06666667 )
        (     1 , 46.26666667 )
        (     2 , 19          )
        (     3 , 67.6        )
        (     4 , 17.6        )
        (     5 , 39.8        )
        (     6 , 174.9333333 )
        (     7 , 30.46666667 )
        (     8 , 32.46666667 )
        (     9 , 42.33333333 )};

  \addplot+[line width = 1.3pt] coordinates {
        (0 , 30.21428571 )
        (1 , 34.21428571 )
        (2 , 17.07142857 )
        (3 , 93.78571429 )
        (4 , 28.64285714 )
        (5 , 67.64285714 )
        (6, 75.5        )
        (7 , 25.71428571 )
        (8 , 54.85714286 )
        (9 , 55)};

      \legend{accented, neutral}

  \end{axis}
\end{tikzpicture}
		% \fbox{\includegraphics[width=5.5cm]{pictures/paper/minmax_ge.png}}
	    \label{fig:minmax_ge}
	}
	\caption[]{Differences between minimal and maximal F0 values (in Hz) for 
	the single regions for \as-sentences (a) and \es-sentences (b).}
	\label{fig:minmax_targets}
\end{figure}

\begin{figure}[]
	\centering
	\subfloat[][\as-sentences]{ 
\begin{tikzpicture}[scale=0.75]
  \begin{axis}[
    cycle list name = mylist,
    legend pos = north west,
    ymajorgrids,
    ymax = 800,
    ymin = 0,
    y = 0.125,
    x tick label style = {rotate = 50, anchor = east},
    xtick = data,
    xticklabels = {alle,diese,N1,sind,mit,einigen,ihrer,N2,verbunden}
    ]
    \addplot+[line width = 1.3pt] coordinates {
      (0 ,225.8333333)
      (1 , 245.3333333 )
      (2 , 397         )
      (3 , 172.3333333 )
      (4 , 230.1666667 )
      (5 , 445.8333333 )
      (6, 199.6666667)
      (7, 503)
      (8 ,722 )
    };

  \addplot+[line width = 1.3pt] coordinates {
        (0 , 216.9212    )
        (1 , 234.6666667 )
        (2 , 414.6666667 )
        (3 , 175.5       )
        (4 , 170.5       )
        (5 , 422         )
        (6 , 194.6666667 )
        (7 , 494.6666667 )
        (8 , 748)
};

      \legend{accented, neutral}

  \end{axis}
\end{tikzpicture}
		% \fbox{\includegraphics[width=5.5cm]{pictures/paper/duration_ae.png}}
	    \label{fig:duration_ae}
	}
	\subfloat[][\es-sentences]{
          \begin{tikzpicture}[scale=0.75]
  \begin{axis}[
    cycle list name = mylist,
    legend pos = north west,
    ymajorgrids,
    ymax = 800,
    ymin = 0,
    y = 0.125,
    x tick label style = {rotate = 50, anchor = east},
    xtick = data,
    xticklabels = {genau,einer,dieser,N1,ist,mit,einigen,seiner,N2,verbunden}
    ]
    \addplot+[line width = 1.3pt] coordinates {
(       0 , 307         )
(       1 , 256         )
(       2 , 128.6666667 )
(       3 , 398.6666667 )
(       4 , 164.6666667 )
(       5 , 226.6666667 )
(       6 , 420.6666667 )
(       7 , 235.3333333 )
(       8 , 494.6666667 )
(       9 , 704.6666667 )
    };

  \addplot+[line width = 1.3pt] coordinates {
(0 , 285.8333333 )
(1 , 244.6666667 )
(2 , 128         )
(3 , 400         )
(4 , 152.6666667 )
(5 , 170         )
(6 , 383.3333333 )
(7 , 249.3333333 )
(8 , 494.6666667 )
(9 , 764.6666667 )
};

      \legend{accented, neutral}

  \end{axis}
\end{tikzpicture}
		% \fbox{\includegraphics[width=5.5cm]{pictures/paper/duration_ge.png}}
	    \label{fig:duration_ge}
	}
	\caption[]{Durational values (in ms) for 
	the single regions for \as-sentences (a) and \es-sentences (b).}
	\label{fig:duration_targets}
\end{figure}

Figure~\ref{fig:duration_targets} shows the durational values for each
of the single regions in the sentence.  Though descriptively small,
the duration of accented determiners was significantly increased as
opposed to their non-accented counterparts (effects on the determiner:
\acro{Quantifier}: \emph{F}=31.4; \emph{p}<.001; \acro{Prosody}:
\emph{F}=23.8; \emph{p}<.001).  A list of the statistical results of
each sentential region is provided in
Appendix~\ref{sec:audit-sent-mater} (F0 values:
Table~\ref{tab:table-B}; durational values: Table~\ref{tab:table-A}).

\paragraph{Preference-related controls.} 
Differences between maximal and minimal F0 values for each of the
single words in the sentence as well as durational values are depicted in
Figure~\ref{fig:acoustics_fi}. As is
evident, the largest durational and F0 differences were
realized at the boundary regions (i.e. Region 5 and Region
7).

\begin{figure}[t]
	\centering
	\subfloat[][F0 values]{ 
\begin{tikzpicture}[scale=0.75]
  \begin{axis}[
    cycle list name = mylist,
    legend pos = north west,
    ymajorgrids,
    ymax = 250,
    ymin = 0,
    y = 0.55,
    x tick label style = {rotate = 50, anchor = east},
    xtick = data,
    xticklabels = {D1,N1,ist,mit,N2,und,N3,mit,N4,verbunden}
    ]
    \addplot+[line width = 1.3pt] coordinates {
(       0 ,27.65517241 )
(       1 ,91.20689655 )
(       2 ,32.51724138 )
(       3 ,27.20689655 )
(       4 ,53.72413793 )
(       5 ,19.5862069  )
(       6 ,210.8965517 )
(       7 ,15.48275862 )
(       8 ,66.06896552 )
(       9 ,68.72413793 )
    };

  \addplot+[line width = 1.3pt] coordinates {
(0 ,23.48148148) 
(1 ,63.07407407) 
(2 ,13.7037037 ) 
(3 ,22.7037037 ) 
(4 ,197.0740741) 
(5 ,34.03703704) 
(6 ,95.37037037) 
(7 ,15.22222222) 
(8 ,52.18518519) 
(9 ,123        ) };

  \addplot+[line width = 1.3pt] coordinates {
(0, 31.7        )
(1, 80.2        )
(2, 24.46666667 )
(3, 31.2        )
(4, 135.8333333 )
(5, 54.46666667 )
(6, 86.96666667 )
(7, 16.76666667 )
(8, 51.4        )
(9, 62.53333333 )};

      \legend{\textsc{ec},\textsc{lc},neutral}

  \end{axis}
\end{tikzpicture}
		% \fbox{\includegraphics[width=5.5cm]{pictures/paper/minmax_fi.png}}
	    \label{fig:acoustics_fi}
	}
	\subfloat[][Duration]{
\begin{tikzpicture}[scale=0.75]
  \begin{axis}[
    cycle list name = mylist,
    legend pos = north west,
    ymajorgrids,
    ymax = 1400,
    ymin = 0,
    y = 0.1,
    x tick label style = {rotate = 50, anchor = east},
    xtick = data,
    xticklabels = {D1,N1,ist,mit,N2,und,N3,mit,N4,verbunden}
    ]
    \addplot+[line width = 1.3pt] coordinates {
(0  , 105.75      )
(1  , 332.3333333 )
(2  , 168         )
(3  , 135.3333333 )
(4  , 561.3333333 )
(5  , 143.3333333 )
(6  , 1183.666667 )
(7  , 167         )
(8  , 384.6666667 )
(9  , 691.3333333 )
    };

  \addplot+[line width = 1.3pt] coordinates {
(0  , 100.75      )
(1  , 306.6666667 )
(2  , 167         )
(3  , 139.6666667 )
(4  , 1268.333333 )
(5  , 176.3333333 )
(6  , 535.3333333 )
(7  , 150.3333333 )
(8  , 381         )
(9  , 681.3333333 )
};

  \addplot+[line width = 1.3pt] coordinates {
(0  , 98.08333333 )
(1  , 313.6666667 )
(2  , 169.3333333 )
(3  , 141.3333333 )
(4  , 642.6666667 )
(5  , 150         )
(6  , 557.3333333 )
(7  , 151         )
(8  , 402.6666667 )
(9  , 730)
};

      \legend{\textsc{ec},\textsc{lc},neutral}

  \end{axis}
\end{tikzpicture}
		% \fbox{\includegraphics[width=5.5cm]{pictures/paper/duration_fi.png}}
	    \label{fig:duration_fi}
	}
	\caption[]{Differences between minimal and maximal F0 values (in Hz) (a) as well
	as durational values (in ms)(b) for 
	the single regions  .}
	\label{fig:acoustics_fi}
\end{figure}

Statistical analyses for F0 on Region 5 reveal that all prosodic
realizations (\emph{neutral} vs. (\emph{early boundary}
vs. (\emph{late boundary})) differ significantly from each other
(effect of \acro{Prosody}: \emph{F}=129.3; \emph{p}<.001; all single
comparisons \emph{p}<.001). On Region 7, F0 analyses again show a main
effect of \acro{Prosody}: \emph{F}=132.5; \emph{p}<.001. However, the
comparison between neutral prosody and a \emph{late boundary} did not
reach significance (\emph{p}>.2). Finally, durational analyses reveal
that all prosodic realizations (\emph{neutral} vs. (\emph{early
  boundary} vs. (\emph{late boundary})) differ significantly from each
other (Region 5: effect of \acro{Prosody}: \emph{F}=819.5;
\emph{p}<.001; all single comparisons \emph{p}<.001; Region 7: effect
of \acro{Prosody}: \emph{F}=1920.3; \emph{p}<.001; all single
comparisons \emph{p}<.001.) A list of the statistical results of each
sentential region is provided in Appendix~\ref{sec:audit-sent-mater}
(F0 values: Table~\ref{tab:table-D}; durational values:
Table~\ref{tab:table-C}).

In sum, our speaker reliably
produced (i) differences in accent realization for the target
sentences and (ii) the expected boundary realizations for the
preference-related controls. Whereas accented elements clearly
differed from their unaccented counterpart by showing an increase in
duration and F0 range, prosodic boundaries for our preference-related
controls were realized by pre-final lengthening (i.e., an increase in F0
and duration at the position preceding the boundary).

\paragraph{Pictures.}
The visual sequences accompanying each sentence consisted of pictures
like in Figures~\ref{fig:exseqAS}--\ref{fig:exlc}. In each picture,
four alternative sub-scenarios were presented, in which an icon was
surrounded by six geometrical shapes. Depending on the sentence
material, shapes and icons differed across sentence types and
conditions. For example, in our target conditions, 4 identical icons
were surrounded by geometrical shapes that were all of the same type,
whereas various icons and surrounding elements were used in each
sequence of the preference-related control structures (see
Section~\ref{sec:design} for more detailed exposition). Pictures for
unrelated fillers had critical positions at random uncovering stages,
including cases where no uncovering was needed for a truth-value
judgement.

\paragraph{Lists.}
Experimental items and preference-related controls were evenly
distributed across lists. For the \as- and \es-sentences together with
their respective controls, five lists were used employing a Latin
square design. Picture sentence-pairs for the preference-related
controls were spread across eight lists. Each target list was combined
with each list from the preference-related controls, thus yielding to
a total of forty lists. The thirty unrelated filler items were
included in each of these resulting lists.

\subsubsection{Participants.}

Forty native speakers of German took part in our study, none of which
had any prior exposure to logic or formal semantics. We excluded two
participants due to insufficient performance on controls ($\le 60\%$
correct answers).

\subsection{Results}
\label{sec:results}

In the following, we will report the results for targets and
preference-related controls separately. After providing the
descriptive results, we present statistical analyses using log-linear
models. We complement these with a Bayesian analysis that is sensitive
to the sequential nature of the incremental verification task and
probes specifically into the preference relations between candidate
readings.

\paragraph{Target conditions and positional controls.}
Performance on positional controls was overwhelmingly correct (90\%
correct responses on average with a variance of 0.012) indicating that
participants understood the task and were able to give correct
truth-value judgments at the adequate point in a sequence. In
general, they were not affected by inessential features of the picture
materials or general response biases.

% \begin{figure}[t] 
% \centering

% \includegraphics[width=0.9\textwidth]{positional_answers_critical.pdf}

% \caption{Answers for the \as- and \es-sequences. Bars show the total
%   number of \emph{yes} (up) and \emph{no} judgments (down) collected
%   at each step during a sequence. E.g., we recorded a total of 103
%   \emph{yes} judgments at the third position in the \as-sequence with
%   neutral intonation, which corresponds to 103 answers that were
%   classified as \emph{literal} answers. Notice that the \as-condition
%   had seven steps, the \es-condition only five.}
% \label{fig:JudgmentsK2}
% \end{figure}
%

\begin{table}
  \centering
  \begin{tabular}{lcccccccc@{\hskip 0.6cm}cccc}
    && \multicolumn{7}{c}{truth value at step}
    & \multicolumn{4}{c}{classification}
    \\ \cmidrule(l){3-9} \cmidrule(r){10-13}
    condition &
    & 1
    & 2
    & 3
    & 4
    & 5
    & 6
    & 7
    & \lit
    & \loc
    & \glb
    & err \\ \midrule
    \as-ntr 
    & T & 1 & 1 & \textbf{103} & 2 & \textbf{0} & 0 & 2 
    & 103 & 3 & 0 & 8 \\
    & F & 0 & 0 & 0 & 0 & 2 & 0 & \textbf{3} \\ 
    \addlinespace[0.25cm]
    \as-acc
    & T & 0 & 1 & \textbf{99} & 2 & \textbf{2} & 0 & 1 
    & 99 & 8 & 2 & 5 \\
    & F & 0 & 0 & 1 & 0 & 0 & 0 & \textbf{8} \\
    \addlinespace[0.25cm]
    \es-ntr
    & T & 0 & 11 & 3 & 1 & \textbf{14} & - & - 
    & 78 & 14 & 0 & 22 \\
    & F & 0 & \textbf{0} & \textbf{78} & 5 & 2 & - & - \\
    \addlinespace[0.25cm]
    \es-acc
    & T & 0 & 10 & 3 & 0 & \textbf{16} & - & - 
    & 78 & 16 & 0 & 20 \\
    & F & 0 & \textbf{0} & \textbf{78} & 5 & 2 & - & -
  \end{tabular}
  \caption{Observed truth-value judgments in critical conditions. The
    left-hand part of the table shows the number of \emph{true} and
    \emph{false} responses obtained at each step in the sequence. The
    right-hand part shows the count for our imposed classification
    scheme. The answer counts indicative of relevant readings are
    given in bold-face on the left-hand side.}
  \label{tab:count_data_CC}
\end{table}

%
% \begin{table}[h]
% \centering

% \begin{tabular}{lccccc}
%   condition & position & \multicolumn{4}{c}{response type} \\ \cmidrule(r){3-6}
%   && \emph{literal} & \emph{local} & \emph{global} & \emph{error} \\
%   \midrule \addlinespace[0.15cm]
%   \textbf{\as neutral} 
%   & Pos 1 & 35 &  0 & 0 & 3 \\
%   & Pos 2 & 35 &  1 & 0 & 2 \\
%   & Pos 3 & 33 &  2 & 0 & 3 \\ \addlinespace[0.15cm]
%   & \textbf{Total} & \textbf{103} &  \textbf{3} & \textbf{0} &
%   \textbf{8} \\
%   \addlinespace[0.25cm]
%   \textbf{\as accented} 
%   & Pos 1 & 33 &  2 & 1 & 2 \\
%   & Pos 2 & 33 &  3 & 1 & 1 \\
%   & Pos 3 & 33 &  3 & 0 & 2 \\ \addlinespace[0.15cm]
%   & \textbf{Total} & \textbf{99} &  \textbf{8} & \textbf{2} &
%   \textbf{5} \\
%   \addlinespace[0.25cm]
%   \textbf{\es neutral} 
%   & Pos 1 & 26 &  5 & 0 & 7 \\
%   & Pos 2 & 26 &  3 & 0 & 9 \\
%   & Pos 3 & 26 &  6 & 0 & 6 \\ \addlinespace[0.15cm]
%   & \textbf{Total} & \textbf{78} &  \textbf{14} & \textbf{0} &
%   \textbf{22} \\
%   \addlinespace[0.25cm]
%   \textbf{\es accented} 
%   & Pos 1 & 26 &  6 & 0 & 6 \\
%   & Pos 2 & 24 &  5 & 0 & 9 \\
%   & Pos 3 & 28 &  5 & 0 & 5 \\ \addlinespace[0.15cm]
%   & \textbf{Total} & \textbf{78} &  \textbf{16} & \textbf{0} &
%   \textbf{20}
% \end{tabular}

% \caption{Contingency table of classified responses for the critical
%   conditions. The factor position encodes whether participants saw the
%   condition in question for the first, second or third time during the
%   whole experiment.}
% \label{tab:CC-ContingencyTable}
% \end{table}
%

Judgements obtained for the four target conditions are listed in
Table~\ref{tab:count_data_CC}. We coded these answers as {\it
  literal}, {\it global} or {\it local} if they were as expected under
one of these readings and as {\it error} if not. The majority of
answers are indicative of a literal reading (\as-neutral: 90.4\%,
\as-accented: 86.8\%, \es-neutral: 68.4\%, \es-accented: 68.4\%). Only
two answers across all critical conditions fall into the category
indicative of global readings. Judgements indicative of local
readings, on the other hand, are more frequent (\as-neutral: 2.6\%,
\as-accented: 7.0\%, \es-neutral: 12.3\%, \es-accented: 14.0\%).

The number of responses that fell out of our classification scheme
differs substantially between \as- and \es-conditions. There are 7.0\%
/ 4.4\% unclassifiable judgements in the neutral / accented
\as-conditions, but 19.3\% / 17.6\% in the neutral / accented
\es-conditions respectively. A look at the left-hand side of
Table~\ref{tab:count_data_CC} reveals that there is some systematicity
in 'error' answers. Most answers that we classified as 'errors' have a
plausible explanation: random errors, a handful of mistakes resulting
from confusing buttons for \emph{true} and \emph{false} and a few
'spill-over errors' where participants gave responses one position too
late. There is one notable exception, though. A substantial number of
\emph{true} responses were given at the second position of
\es-sequences (9.7\% / 8.8\%) which is shown in
Figure~\ref{fig:exseqES2}. Although this position in the sequence is
the critical position to reveal a global reading, the latter would be
indicated by a \emph{false} response. It appears that subjects read
\es-sentences as if the upward-monotonic quantifier \emph{exactly one}
was a monotonic \emph{at least one}, without pragmatic enrichment of
\emph{some}. 16 out of the total 21 answers in this category come from
5 subjects, while the remaining 5 were single answers by 5 different
subjects.
% subjects 1, 9, 12, 21, 84
This suggests that at least for some subjects non-monotonic
quantifiers were problematic. (We will come back to this intriguing
observation later in Section~\ref{sec:discussion-1}, where we suggest
that a non-monotonic reading of \emph{exactly one} might also explain
the high number of \emph{true} answers in step 5 of the
\es-conditions.)

We fitted generalized linear models to the count data in
Table~\ref{tab:count_data_CC}, using factors \textsc{Condition} with
levels \as and \es, \textsc{Accent} with levels \emph{neutral} and
\emph{accented}, and \textsc{Trial} with levels 1 through 3 (coding
whether it was the first, second or third time that a participant saw
a critical condition during the experiment) to predict the dependent
factor \textsc{Reading} with levels \emph{literal}, \emph{local}, and
\emph{other}. The latter lumps \emph{global} and \emph{error}
responses together, because otherwise we would not have enough cell
counts to apply log-linear regression. We determined the ``best''
model of the data by a gradient search over hierarchical models in
terms of AICs, starting from the saturated model. The best model takes
main factors \textsc{Reading} and \textsc{Condition} into account, as
well as the interaction term \textsc{Reading:Condition} ($\chi^2 =
17.19$, $df=42$, $p = 0.99$, $AIC = 158.01$ (compared to $AIC=208.25$
of the saturated model)). Crucially, factors \textsc{Trial} and
\textsc{Accent} were dropped in the best model. This suggests that
response patterns only depend on the type of the sentence, but that
there were no learning or fatigation effects and that accentuation did
not influence answer patterns significantly (at this level of
abstraction).

\begin{table}
  \centering
  \begin{tabular}{lllll}
    coefficient & estimate & std.~error & $z$-value & $\Pr(>|z|)$ \\
    \midrule
    \textsc{Intercept} & 0.916  &  0.258 &  3.49  & < 0.001 \\
    \textsc{Reading}.\emph{literal} & 2.600 &   0.268 & 9.716 & < 0.001 \\
    \textsc{Reading}.\emph{local} & -0.310  &  0.397 & -0.781 &
    0.435 \\
    \textsc{Condition}.\es & 1.030 &  0.301 & 3.423 &  < 0.001 \\
    \textsc{Reading}.\emph{literal}:\textsc{Condition}.\es & -1.288 &
    0.319 & -4.036 &  < 0.001 \\
    \textsc{Reading}.\emph{local}:\textsc{Condition}.\es & -0.026 &
    0.463 &  -0.0579 &  0.955
  \end{tabular}
  \caption{Coefficients of the ``best'' log-linear model for the count
    data in Table~\ref{tab:count_data_CC}.}
  \label{tab:Coefficients-CC}
\end{table}

Inspection of the coefficients of the ``best'' model in
Table~\ref{tab:Coefficients-CC} suggests that the distinction between
levels \emph{local} and \emph{other} in factor \textsc{Reading} might
not be necessary: factor levels \emph{local} do not cause a
significant shift in counts, given reference level \emph{false}. In
order to test whether there is support for the postulation of local
readings, we therefore compared the previous ``best'' model to a
model that only differs from the former in that it subsumes the level
\emph{local} of factor \textsc{Reading} under level \emph{other} as
well \citep[e.g.][Chapter~15]{Crawley2007:The-R-Book}. Model
comparison reveals that there is no significant improvement in
explanatory power ($Pr(|\chi|) = 0.2689$) of the more complex model
that includes the level \emph{local} (residual df=30, residual
deviance = 10.497) over the simpler model without this level (residual
df=32, residual deviance = 13.124, AIC=157.268).

This latter analysis suggests that our data does not provide strong
evidence for the maintenance of belief in local readings in our
setting. The regression analysis tells us that the number of errors
and the number of local readings were similar in all conditions. But
the distribution of answers in Table~\ref{tab:count_data_CC} shows
that small error counts occur at many different steps, whereas we do
see a concentrated rise in answers at the critical position for local
readings. Also, so far we have not taken into account the sequential
structure of our task. This is where preference-related controls
are relevant.

\paragraph{Preference-related controls.} Positional answers for our
preference-related control conditions are shown in
Table~\ref{tab:count_data_TF}.\fn{We removed 19 data points for the
  \ec-conditions because of a coding mistake that presented subjects
  with incongruent sentence-picture pairs. Including these and
  classifying them as \emph{errors} does not change the qualitative
  results of the reported analyses.} Judgments were coded as \ec or as
\lc if they were as expected under one of these readings, and as
\emph{error} if they were not. As expected, most answers classify as
\lc-readings, fewer as \ec-readings. Moreover, prosody does seem to
have the expected effect as well: as compared to neutral phrasing,
\lc-cueing prosody increases the count for \lc-readings, while
\ec-cueing prosody increases the count for \ec-readings. If we compare
the number of \lc- or \ec-readings for each prosodic variant across
sequences, we see that there are always more answers indicative of the
relevant readings when that reading can be judged later in the
sequence. This suggests a general tendency to answer later in the
sequence, rather than at the earliest possible step.

\begin{table}
  \centering
  \begin{tabular}{lcccccccc@{\hskip 0.6cm}ccc}
    && \multicolumn{7}{c}{truth value at step}
    & \multicolumn{3}{c}{classification}
    \\ \cmidrule(l){3-9} \cmidrule(r){10-12}
    condition &
    & 1
    & 2
    & 3
    & 4
    & 5
    & 6
    & 7
    & \lc
    & \ec
    & err \\ \midrule
    \lc-ntr 
    & T & 0 & 0 & \textbf{87} & 1 & 0 & 9 & 0 
    & 87 & 32 & 13 \\
    & F & 0 & 0 & 3 & 0 & 0 & \textbf{32} & 0 \\
    \addlinespace[0.25cm]
    \lc-\lc-cue
    & T & 0 & 0 & \textbf{108} & 1 & 1 & 7 & 0 
    & 108 & 16 & 10 \\
    & F & 0 & 0 & 1 & 0 & 0 & \textbf{16} & 0 \\
    \addlinespace[0.25cm]
    \lc-\ec-cue
    & T & 0 & 0 & \textbf{51} & 4 & 0 & 7 & 0 
    & 51 & 59 & 25 \\
    & F & 2 & 1 & 7 & 0 & 0 & \textbf{59} & 1 \\
    \addlinespace[0.25cm]
    \ec-ntr
    & T & 0 & 0 & 0 & 0 & \textbf{92} & - & - 
    & 92 & 22 & 13 \\
    & F & 1 & 0 & \textbf{22} & 6 & 6 & - & - \\
    \addlinespace[0.25cm]
    \ec-\lc-cue
    & T & 0 & 0 & 0 & 0 & \textbf{112} & - & - 
    & 112 & 13 & 4 \\
    & F & 0 & 1 & \textbf{13} & 1 & 2 & - & - \\
    \addlinespace[0.25cm]
    \ec-\ec-cue
    & T & 0 & 0 & 0 & 0 & \textbf{65} & - & - 
    & 65 & 31 & 30 \\
    & F & 2 & 0 & \textbf{31} & 8 & 20 & - & - \\
    \addlinespace[0.25cm]

  \end{tabular}
  \caption{Observed truth-value judgments in preference-related controls.}
  \label{tab:count_data_TF}
\end{table}

% \begin{figure}[]
% \centering

%  \subfloat[][\LC-sequence]{

% \includegraphics[width=0.9\textwidth]{positional_answers_TF_LC.pdf}
% }

%  \subfloat[][\EC-sequence]{

% \includegraphics[width=0.9\textwidth]{positional_answers_TF_EC.pdf}
% }

% \caption{Answers for \lc-sequences and \ec-sequences. \emph{True}
%   judgements are plotted upwards for each step in the sequence,
%   \emph{false} judgements downwards.}
% \label{Fig:TF_judgments}
% \end{figure}

% \begin{table}[]
% \centering

% \begin{tabular}{lcccc}
%   condition & position & \multicolumn{3}{c}{response type} \\ \cmidrule(r){3-5}
%   && \emph{LC} & \emph{EC} & \emph{error}  \\
%   \midrule \addlinespace[0.15cm]
%   \textbf{\lc neutral} 
%   & Pos 1 & 27 &  7  & 4 \\
%   & Pos 2 & 23 &  11 & 4 \\
%   & Pos 3 & 26 &  9  & 3 \\ \addlinespace[0.15cm]
%   & \textbf{Total} & \textbf{76} &  \textbf{27} & \textbf{11}  \\
%   \addlinespace[0.25cm]
%   \textbf{\lc with \lc cue} 
%   & Pos 1 & 26 &  7 & 5  \\
%   & Pos 2 & 31 &  5 & 2  \\
%   & Pos 3 & 33 &  3 & 2  \\ \addlinespace[0.15cm]
%   & \textbf{Total} & \textbf{90} &  \textbf{15} & \textbf{9}  \\
%   \addlinespace[0.25cm]
%   \textbf{\lc with \ec cue} 
%   & Pos 1 & 14 &  17 & 7  \\
%   & Pos 2 & 15 &  16 & 7  \\
%   & Pos 3 & 15 &  17 & 6  \\ \addlinespace[0.15cm]
%   & \textbf{Total} & \textbf{44} &  \textbf{50} & \textbf{20}  \\
%   \addlinespace[0.25cm]
%   \textbf{\ec neutral} 
%   & Pos 1 & 22 &  6 & 5  \\
%   & Pos 2 & 28 &  6 & 4  \\
%   & Pos 3 & 29 &  7 & 2  \\ \addlinespace[0.15cm]
%   & \textbf{Total} & \textbf{79} &  \textbf{19} & \textbf{11}  \\
%   \addlinespace[0.25cm]
%   \textbf{\ec with \lc cue} 
%   & Pos 1 & 30 &  5 & 1  \\
%   & Pos 2 & 32 &  5 & 0  \\
%   & Pos 3 & 31 &  3 & 2  \\ \addlinespace[0.15cm]
%   & \textbf{Total} & \textbf{93} &  \textbf{13} & \textbf{3}  \\
%   \addlinespace[0.25cm]
%   \textbf{\ec with \ec cue} 
%   & Pos 1 & 19 &  10 & 7  \\
%   & Pos 2 & 19 &  7  & 9  \\
%   & Pos 3 & 19 &  7  & 11  \\ \addlinespace[0.15cm]
%   & \textbf{Total} & \textbf{57} &  \textbf{24} & \textbf{27}  \\
% \end{tabular}

% \caption{Contingency table of classified responses for the
%   preference-related controls. As before, the factor position encodes
%   whether participants saw the condition in question for the first,
%   second or third time during the whole experiment.}
% \label{tab:CC-ContingencyTable-TF}
% \end{table}

There are only very few answers that appear to be random
mistakes. Some answers in the `error' category are plausibly
`spill-over errors', where participants gave answers one step too
late. But on top of that, a substantial number of `error' answers
occurs on critical positions. For example, in the \lc-sequence there
are 23 \emph{true} answers at step 6 where \emph{false} answers would
be taken as indicative of \ec-readings. In the \ec-sequence there are
28 \emph{false} answers at step 5 where \emph{true} answers would be
taken as indicative of \lc-readings. Finally, prosodic cues for
dispreferred \ec-readings seem to have led to more mistakes than
other prosodic patterns.

As for the critical conditions, we computed logistic regression models
including factors \textsc{Reading} with levels \lc, \ec and
\emph{error}, \textsc{Accent} with levels \emph{neutral},
\emph{\lc-cue} and \emph{\ec-cue}, \textsc{Trial} with levels 1 to
3, and \textsc{Condition} with levels \lc and \ec for the picture
sequences. Starting search from the saturated model along a gradient
of decreasing AICs, we find that the ``best'' model relies on main
effects of \textsc{Reading}, \textsc{Accent} and \textsc{Condition}
and two two-way interactions \textsc{Reading:Condition} and
\textsc{Reading:Accent} ($\chi^2=18.34$, $df=42$, $p=0.99$,
$AIC=256.63$ (compared to $AIC=322.3$ of the saturated model)). Factor
\textsc{Trial} was dropped from the best model, suggesting that
answer patterns are stable over the duration of the experiment. Unlike
for the critical conditions, factor \textsc{Accent} is retained, and
it interacts with \textsc{Reading}, suggesting that prosodic cues
significantly influenced the dominant reading.

\begin{table}
  \centering
  \begin{tabular}{lllll}
    coefficient & estimate & std.~error & $z$-value & $\Pr(>|z|)$ \\
    \midrule
    \textsc{Intercept} & 2.119  &   0.162 & 13.079   & < 0.001 \\
    \textsc{Reading}.\emph{error} & -0.343 &  0.259 & -1.328 &  0.184
    \\
    \textsc{Reading}.\lc & 0.526 &  0.202 & 2.607 & < 0.01 \\
    \textsc{Condition}.\lc & 0.497 &  0.170  & 2.929 & < 0.01 \\
    \textsc{Accent}.\lc & -1.054 & 0.242 & -4.354 & < 0.001 \\
    \textsc{Accent}.\emph{neutral} & -0.112 & 0.180 &  -0.626 &  0.531
    \\
    \textsc{Reading}.\emph{error}:\textsc{Condition}.\lc & -0.521 &
    0.280 & -1.87 & 0.062 \\
    \textsc{Reading}.\lc:\textsc{Condition}.\lc & -0.583 & 0.195 &
    -2.997 &  < 0.01 \\
    \textsc{Reading}.\emph{error}:\textsc{Accent}.\lc & -0.103 &
    0.422 &  -0.245 & 0.807  \\
    \textsc{Reading}.\lc:\textsc{Accent}.\lc & 1.657 &  0.279 & 5.943
    & < 0.001 \\
    \textsc{Reading}.\emph{error}:\textsc{Accent}.\emph{neutral} & 0.112 &
    0.299  &  0.375 &  0.708 \\
    \textsc{Reading}.\lc:\textsc{Accent}.\emph{neutral} & 1.065  &
    0.222 &  4.800 & < 0.001 \\
  \end{tabular}
  \caption{Coefficients of the ``best'' log-linear model for the count
    data in Table~\ref{tab:count_data_TF}.}
  \label{tab:Coefficients-TF}
\end{table}

Closer examination of the coefficients of the fitted model, shown in
Table~\ref{tab:Coefficients-TF}, reveals that there were significantly
more \ec-readings in the \lc-condition than in the \ec-condition, and
more \lc-readings in the \ec-condition than in the \lc-condition. This
is evident from the significant \emph{negative} deviation from the
baseline of \ec-readings in the interaction coefficient for
\textsc{Reading}.\lc:\textsc{Condition}.\lc. The log-linear model
therefore indicates that there is a bias for late responses in a
sequence. Given this, the number of local responses in critical
conditions might be increased by a sequential bias to answer late. The
following Bayesian analysis further probes into this matter in order
to establish how strongly the sequence-effect might have emphasized
local readings and, from there, whether we should uphold belief in
local readings.

\subsubsection{Probabilistic salience competition.}

The previous regression-based analyses did not take the sequential
nature of the task into account. The following paragraphs therefore
spell out a generative Bayesian model which computes the
likelihood of answer patterns for different latent relative saliencies
of candidate readings and biases to answer early or late in a
sequence. We use the data to estimate the posterior likelihood of
these latent parameters to gain insight into the likely ordering
relations over the strength of latent salience of candidate readings.

\paragraph{The model.} We present a model that aims to capture the
competition between candidate readings, all of which have a given
latent level of relative salience, in the incremental verification
task, where some readings can be judged before others. Take the
AS-conditions with its three candidate readings. The relative strength
between these is given by a probability vector $\myvec{p}^\text{as-n}
=
\tuple{p^\text{as-n}_{\text{lit}},p^\text{as-n}_{\text{loc}},p^\text{as-n}_{\text{glb}}}$. We
think of $p^\text{as-n}_{\text{lit}}$, for instance, as the salience
of the literal reading for \as-sentences with neutral prosody,
relative to the global and local reading. If there were no potential
effects of sequentiality and no errors (or other readings that we
classify as errors), the vector $\myvec{p}^\text{as-n}$ would be our
prediction of observed frequencies of answer patterns. We look at four
such three-placed vectors for the critical conditions, one for each
target sentence-prosody pair. For the preference-related control
conditions, we look at three vectors $p^k =
\tuple{p^{\text{k}}_{\text{\lc}}, p^{\text{k}}_{\text{\ec}}}$, where
$k \in \set{\text{ntr}, \text{\lc-cue}, \text{\ec-cue}}$. These
vectors give the relative salience of \lc- and \ec-readings for
different types of prosody. The salience of readings is independent
of the pictorial sequence. If there is a tendency to judge sentences
earlier in the sequence, rather than later, that has to be attributed
to the sequential bias.

To add potential effects of sequentiality, consider bias factor $q \in
[0;1]$. This bias is a global parameter held constant over all
conditions, because we think of it as a general tendency to answer
early or late in the incremental verification task. The bias factor
captures whether there is a tendency to choose readings that appear
earlier ($q > .5$) or later ($q <.5$) in the sequence. In the
\as-sequences, the literal reading can be judged first, then the local
and then the global reading. So we will assume that $p_{\text{lit}}$
is multiplied by $q$, that $p_{\text{loc}}$ is multiplied by $(1-q)q$
and that $p_{\text{glb}}$ is multiplied by $(1-q)^2$, when determining
the eventual weights of readings.

Finally, we also allow for mistakes. To keep things simple (no
differential consideration for spill-overs, button mix-ups etc.), we
assume that there is a fixed error rate $e^k \in [0;1]$ for each
relevant condition $k$. We allow for different error rates for
different conditions, because we want to allow for the possibility
that some sentence types (e.g., non-monotonic quantifiers) or some
kind of prosody (e.g., \ec-cues) might be harder to process. A
decision can be wrong but incidentally coincide with an answer that is
indicative of a relevant reading. Error rates multiply in proportion
to the number of steps in a sequence: more choice, more chance to be
wrong.

Taking this together we calculate, for each pair $k$ of target
sentence type and prosody type, a probability vector $\myvec{t}^k =
\tuple{t^k_{\text{lit}},t^k_{\text{loc}},t^k_{\text{glb}},t^k_{\text{err}}}$
with which we expect an answer to be classified as literal, local,
global or as error. For example, for \as-sentences with neutral
prosody, where literal can be judged before global before local, we
get:
\begin{align*}
  t^{\text{as-n}}_{\text{lit}} & \propto q \
  p^\text{as-n}_{\text{lit}} + e^\text{as-n} &   t^{\text{as-n}}_{\text{loc}} & \propto (1-q)^2 \
  p^\text{as-n}_{\text{loc}} + e^\text{as-n} \\
  t^{\text{as-n}}_{\text{glb}} & \propto (1-q)q \
  p^\text{as-n}_{\text{glb}} + e^\text{as-n} &   t^{\text{as-n}}_{\text{err}} &
  \propto 12 e^\text{as-n}
\end{align*}
Two remarks. The vector $\myvec{t}^k$ should be normalized eventually,
which is why `$\propto$' is used above instead of `$=$.' Also, the
probability of error $t^{\text{as-n}}_{\text{err}}$ comes from the
observation that there are 15 positions in total in an \as-sequence at
which subjects can make an erroneous choice, but we have accounted for
three of these already as adding to the count of answers indicative of
relevant readings.

Similarly, we compute a target probability $\myvec{t}^k =
\tuple{t^k_{\text{late}},t^k_{\text{early}},t^k_{\text{err}}}$ for
each preference-related control sentence under prosody $k$, taking
into account which reading can be judged first and the length of the
respective sequences.

The full probabilistic model is visualized in
Figure~\ref{fig:model_graph}. As indicated there, we assume largely
uninformative priors. Any positional bias $q \in [0;1]$ is assumed to
be equally likely. The error rates for each condition should be small,
so we sample uniformly from interval $[0,.2]$. The biases
$\myvec{p}_k$ are also determined uniformly at random, by sampling from a
Dirichlet distribution with equal weights on all dimensions.

\begin{figure}[]
  \centering
  \begin{tikzpicture}[node distance = 2cm, 
    double distance = 2pt,
    minimum size=1.25cm,
    thick]

    \node[rectangle, draw=black, fill=black!30] (obs)
    {obs$_k$};

    \node[rectangle, draw=black, fill=black!30,
    right of = obs] (n) {$n_k$};

    \node[below of = n, node distance= 1.5cm]
    (anchor) {};

    \node[left of = anchor, node distance= 0.95cm] (k) {$k \in
      \set{\text{\as-ntr, \as-acc, \dots}}$};

    \node[circle, double, draw=black,
    above of = obs] (t) {$\myvec{t}_k$};

    \node[circle, draw=black, 
    above of = t] (p) {$\myvec{p}_k$};

    \node[circle, draw=black, 
    right of = p] (e) {$e_k$};

    \node[circle, draw=black, 
    left of = p] (q) {$q$};

    \draw[->] (t)--(obs);

    \draw[->] (n)--(obs);

    \draw[->,shorten >=0.05cm] (q)--(t);

    \draw[->,shorten >=0.05cm] (p)--(t);

    \draw[->,shorten >=0.05cm] (e)--(t);

    \begin{pgfonlayer}{background}
       \node [rounded corners,
       very thick,draw=black!40,fit={($(obs.south)+(0,-30pt)$) 
                                 ($(n.east)+(+5pt,0)$) 
                                 ($(e.north)+(0,+5pt)$) 
                                 ($(p.west)+(-5pt,0)$)}] {};
    \end{pgfonlayer}

    \begin{scope}[xshift = 6.5cm, node distance = 1cm]

      \node[] (obsd) {$\text{obs}_k \sim \text{Multinomial}(\myvec{t}_k,n_k)$};

      \node[above of = obsd] (td) {$\myvec{t}_k = \text{as described
          in text}$};

      \node[above of = td] (pd) {$\myvec{p}_k \sim
        \text{unbiased Dirichlet}$};

      \node[above of = pd] (ed) {$e_k \sim \mathcal{U}(0,.2)$};

      \node[above of = ed] (qd) {$q \sim \mathcal{U}(0,1)$};

    \end{scope}

  \end{tikzpicture}
  \caption{Probabilistic graphical model. All nodes represent
    variables, arrows represent functional relations between them.
    Using conventions of \citet{LeeWagenmakers2013:Bayesian-Cognit},
    the higher a node, the more latent it is. Square nodes represent
    discrete-valued variables, circular nodes represent
    continuous-valued variables. Gray nodes show the observable
    variables. Values at double-lined nodes are derived
    deterministically. The variables in the box are repeated for each
    of the $k$ relevant cases.}
  \label{fig:model_graph}
\end{figure}

\paragraph{Model fitting.} We used JAGS
\citep{Plummer2003:JAGS:-A-Program} to estimate the posteriors over
paramater values given our data. Results reported here are based on
two chains of 10.000 MCMC samples from the joint posterior
distribution, obtained after an initial burn-in of 10.000 steps. The
latter guaranteed convergence.

To check whether the model yields sensible results we first look at
the preference-related control conditions.
%
\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{pics/post_salience_TF.pdf}
  \caption{Estimated posteriors for the salience of \lc- and
    \ec-readings for different prosodic cues. Lines under curves
    indicate 95\% HDIs (see Footnote~\ref{fn:HDI}).}
  \label{fig:Posterior_TF}
\end{figure}
%
Estimates for the (marginalized) posteriors over relative salience
levels for target-related controls are plotted in
Figure~\ref{fig:Posterior_TF}. The results are exactly as we would
expect them to be. Given our data, we should believe that the
\lc-reading is most prominent. The level of prominence varies with
accentuation. Under prosody that we hypothesized would favor the
\lc-reading, the contrast is most pronounced. Under accentuation that
we hypothesized favors the \ec-reading we are no longer justified in
believing with certainty that the \lc-reading is preferred (by
comparison of 95\% HDIs, see below).\footnote{\label{fn:HDI} A 95\%
  Highest Density Interval (HDI) is a convex region of values over
  which 95\% of the distribution's probability density is distributed,
  such that no value outside of that region has higher probability
  density than any point within
  \citep{Kruschke2011:Doing-Bayesian-}. Intuitively speaking, the 95\%
  HDI is the set of values which we may believe to be true with some
  confidence; what falls outside this region is what can be excluded
  safely.}

\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{pics/post_salience.pdf}
  \caption{Posteriors over relative salience of target readings in
    critical conditions.}
  \label{fig:Posterior_T}
\end{figure}

Estimates of posteriors over salience of readings in the critical
conditions are given in Figure~\ref{fig:Posterior_T}. In the
\es-conditions, literal readings are by far the most salient, followed
by local, followed by global readings. There does not appear to be a
significant effect of prosody (by visual comparison of 95\%
HDIs). For \as-sentences, the same preference order holds in tendency,
but we cannot assert with full confidence that local readings are
attested or that they are preferred over global readings (see below).

Posteriors over error rates are conceptually less interesting, and we
will skip them here. The posterior over the sequential bias parameter
is shown in Figure~\ref{fig:post_q}. Since the 95\% is clearly
entirely below .5, we have reason for believing, given the model and
our data, in a bias for late responses.

\begin{figure}
  \centering
  \includegraphics[width=0.5\textwidth]{pics/post_q.pdf}
  \caption{Estimated Posteriors over global sequential bias parameter
    $q$.}
  \label{fig:post_q}
\end{figure}

\paragraph{Model validation.} As a crude sanity check that our model
is able to predict the observed data, we gathered 10.000 random values
from the posterior predictive distribution and found a highly
significant correlation between these and the observed data ($R^2 =
0.987$, $p < 0.001$).

\paragraph{Hypothesis testing.} We are chiefly interested in two
questions: (i) which readings are available? and (ii) what is the
preference order over readings?

The first question can be answered by looking at the HDIs of the
posteriors of the candidate readings. We would like to check the
hypothesis that a given reading is unavailable, i.e., its activation
value in the probabilistic model is equal to 0, or close to zero for
practical purposes (Region of Practical Equivalence, ROPE). We reject
the ``null-hypothesis'' that the true value is 0 if its ROPE lies
entirely outside the 95\% HDI; we accept the ``null-hypothesis'' if
the ROPE lies entirely within the 95\% HDI
\citep[][Ch.~12]{Kruschke2011:Doing-Bayesian-}. The following table
lists approximations of the 95\% HDIs for each condition and reading:

\begin{center}
  \begin{tabular}{lllllll}
    & lit & loc & glb & \\ \midrule
    \as-ntr & $(0.924 , 1)$ & $(0 , 0.068)$
    & $(0 , 0.022)$ \\
    \as-acc 
    & $(0.816 , 0.964)$
    & $(0.028 , 0.152)$
    & $(0 , 0.066)$ \\
    \es-ntr 
    & $(0.665 , 0.873)$
    & $(0.117 , 0.323)$
    & $(0 , 0.024)$ \\
    \es-acc 
    & $(0.634 , 0.851)$
    & $(0.139 , 0.356)$
    & $(0 , 0.022)$ 
  \end{tabular}
\end{center}

\noindent We should thus accept the hypothesis that there are no
global readings at all, and no local readings for \as-sentence under
neutral prosody. There is support for a belief in local readings in
the \as-acc condition up to a ROPE of $[0;0.028)$.

The preference relations between readings can be checked in a similar
way. We look at the posterior beliefs about the \emph{differences} of
activation strengths and ask whether these posterior beliefs allow us
to safely conclude that differences are bigger than zero or not. It is
obvious enough that literal readings are always the most preferred,
and that local readings are preferred over global readings for
\es-sentences. Figure~\ref{fig:PostDiffTAS} shows the posteriors over
differences for the only non-trivial cases, namely those between local
and global readings for \as-sentences. Since the 95\% HDIs include 0,
the data provides no ground for confidence that, given the model,
local readings of \as-sentences are preferred over global ones,
despite the obvious tendency. In sum, with the exception of remaining
uncertainty about local readings of \as-sentences, the model and the
data suggest that global readings are absent entirely, while literal
readings are preferred over local readings.

\begin{figure}
  \centering
  \subfloat[][neutral]{
      \includegraphics[width=0.45\textwidth]{pics/PostDiffT_ASntr.pdf}
  }
  \hfill
  \subfloat[][accented]{
      \includegraphics[width=0.45\textwidth]{pics/PostDiffT_ASacc.pdf}
  }
  \caption{Posteriors over differences between salience levels for
    local and global readings of \as-sentences. As zero is contained
    in the 95\% HDI for both neutral and accented variants, there is
    no justification for assuming a difference in salience.}
  \label{fig:PostDiffTAS}
\end{figure}

\subsection{Discussion}
\label{sec:discussion-1}

\paragraph{Summary \& Interpretation.} The present study investigated
prosodic effects on quantifier interpretation by employing a new
experimental paradigm, the incremental verification
task. Based on a current debate in the literature, we were
specifically interested in whether ``local readings'' of sentences
where \emph{some} is embedded under another quantifier can be found,
and how salient these readings are under varying prosody of auditorily
presented sentence material, in particular the presence or absence of
contrastive stress on embedded \emph{some}.

High scores of success in control conditions show that participants
understood the incremental verification task and were able to give
judgements at the right point in a picture sequence. Results for the
preference-related controls show that our design is able to detect
multiple readings and preferences among these. Crucially, the amount
of response types for \lc- and \ec-readings reflected the known
preference for the former, with a mild bias for readings that can be
judged later in the sequence. Also, the incremental verification task
was generally sensitive to prosody. However, significant prosodic
effects were restricted to preference-related controls. Contrastively
accenting scalar \emph{some} generally led to more local readings, but
that effect was not significant. Local readings occurred in
\es-sentences, independently of prosody, but required prosodic marking
in \as-sentences. Our analyses support the impression that inspection
of the raw data gives, namely that the general qualitative pattern of
reading preferences is the same for \as- and \es-conditions: literal
readings are preferred over local readings which are preferred over
global readings. More concretely, our analyses suggest the following
preference patterns (bracketed readings are not supported by our data
and the previous analyses):

\begin{exe}
  \ex \label{bsp:observed-preferences}
    \begin{xlist}
      \ex \as-ntr: \ \ \ \lit \ (>  \loc \ \  $\sim$ \glb)
      \ex \as-acc: \ \ \ \lit \ > \ \loc \ \ ( $\sim$ \glb)
      \ex \es-ntr: \ \ \ \lit \ > \ \loc \ \ ( > \glb)
      \ex \es-acc: \ \ \ \lit \ > \ \loc \ \ ( > \glb)
    \end{xlist}

\end{exe}

\noindent If we compare (\ref{bsp:observed-preferences}) with
Table~\ref{tab:predictions}, we quickly see that these findings are
not predicted by either traditionalism or grammaticalism. We address
each position separately in the following.

Traditionalism easily explains the absence of global readings. Weak
traditionalism, introduced in Section~\ref{sec:traditionalism}, for
instance, would hold that the extra assumptions needed to derive
global readings (e.g., a speaker competence assumption) are not
available or only allow for epistemically weak implicatures that would
show as responses indicative of literal readings in our task. In
support of this view, traditionalists could argue that the task's
rather high processing demands lead to a suppression of global
readings, in line with the findings of
\citet{NeysDe-NeysSchaeken2007:When-People-Are} that working memory
load negatively affects the number of scalar implicature responses in
sentence verification tasks. The small number of local readings in
\as-sentences with contrastive stress is easily explained by adopting
the prosodic markedness hypothesis. What traditionalism does not
predict is the high number of answers indicative of local readings in
the \es-conditions. Resorting to the prosodic markedness hypothesis
does not help, because local readings were observed for accented as
well as for unaccented \es-sentences. It would also remain unclear why
prosodic cues should have such differential effects on \as- and
\es-sentences.

Grammaticalism, on the other hand, fails to predict the pattern in
(\ref{bsp:observed-preferences}) entirely if it equates preference for
readings with logical strength. Grammaticalism predicts that literal
readings of \as-sentences are the least preferred, contrary to
(\ref{bsp:observed-preferences}). Again, we need to bear the caveat of
Section~\ref{sec:grammaticalism} in mind that we are only evaluating
grammaticalism with respect to a particular disambiguation criterion
here. The verdict, however, seems to be clear. If
(\ref{bsp:observed-preferences}) is the actual preference order over
readings, then meaning disambiguation in terms of logical strength is
generally on the wrong track, because no account that merely looks at
the logical strengths of readings can plausibly predict the preference
pattern in (\ref{bsp:observed-preferences}) for both \as- and
\es-sentences at the same time.

In sum, the present findings cannot be explained in terms of the
austere and stylized versions of traditionalism and grammaticalism
that we have set up so far. We should then ask what plausible
additional assumptions would enable either framework to accommodate
the data? Again, let us look at each position in turn.

Consider grammaticalism first. Our data contradicts the notion that
preferences follow logical strength. But grammaticalism could be
supplemented by a conceptually different selection
criterion. Selecting readings in terms of how well they answer the
contextually given question under discussion, as suggested by
\citet{Fox2007:Free-Choice-and,GualminiHulsey2008:The-Question-An}, is
an option. But it is quite unclear what the question under discussion
should be that guided judgements in our task. As long as a theory of
accommodating implicit questions under discussion is missing, this
selection criterion does not make any clear, testable
predictions. However, looking at the qualitative pattern in
(\ref{bsp:observed-preferences}), a \emph{prima facie} plausible
disambiguation criterion is ready at hand: if we assume that each
insertion of the $\exh(\cdot)$ operator is \emph{costly}
anti-proportional to the depth of its embedding under other logical
operators, then we derive a preference of literal readings over local
ones over global ones. Whether such an \emph{economy principle} is
sound and in line with further empirical data would then be an
interesting matter of further investigation. What this economy
principle does not explain yet is the quantitative differences in the
preference orders between \as- and \es-sentences. Still, if combined
with further information about the global properties of the sentences
in question, this might be an interesting alternative disambiguation
criterion for an empirically testable grammaticalism.

Traditionalism, on the other hand, has to explain the possibility of
local readings of prosodically unmarked \es-sentences, and, perhaps,
also the large number of local readings in accented \es-sentences
compared to the much lower number in accented \as-sentences. To do so
within a strict traditionalist framework seems difficult. A first
option might be to account for alleged local readings as the result of
some other phenomenon, such as typicality or contrast
\citep{Tielvan-Tiel2012:Embedded-Scalar,GeurtsTielvan-Tiel2013:Embedded-Scalar,Tielvan-Tiel2014:Quantity-Matter}.
Since this ties into a critical reflection on our design, we will
enlarge on the possibility of typicality- and contrast-based effects
in our setting. Eventually, we argue that contrast-effects alone are
unlikely to explain the observed high number of local answers in
\es-sentences, but that a ``modern traditionalist'' explanation of the
answer pattern might be available if we allow for the possibility that
\emph{exactly one} gets an unexpected ``referential interpretation.''

As discussed in Section~\ref{sec:previous-studies},
\citet{Tielvan-Tiel2012:Embedded-Scalar} and
\citet{GeurtsTielvan-Tiel2013:Embedded-Scalar} argue that the
distribution of responses to \as-sentences found by
\citet{CliftonDube2010:Embedded-Implic}, as well as
\citet{ChemlaSpector2010:Experimental-Ev} could be explained by
differences in how \emph{typical} the presented pictures were for an
\as-sentence. Notice that this does not apply in the present situation
where the task is to explain away the local readings of \es-sentences,
for which no alternative explanation in terms of typicality has been
offered. Moreover, to apply typicality-based explanations to the
incremental verification task, one would have to reason about the
typicality of parts of pictures or to figure in subjects' expectations
of pictorial typicality given their uncertainty about parts of the
picture that they have not yet seen. On intuitive grounds, we take
either extension to be implausible. But even if this line of
explanation can be plausibly extended to incrementally revealed
picture verification, it would still be unclear why typicality-effects
should show so strongly in categorical truth-value judgements, and why
there is such a stark contrast between \as- and \es-sentences.

A more promising alternative explanation for data indicative of local
readings of \es-sentences is pursued by
\citet{GeurtsTielvan-Tiel2013:Embedded-Scalar} in response to the data
reported by \citet{ChemlaSpector2010:Experimental-Ev}. The main idea
is that visual contrast between an item that is connected to all, and
one that is connected to only some but not all of the relevant
elements may trigger exceptional local enrichment of \emph{some}. In
support of this idea,
\citeauthor{GeurtsTielvan-Tiel2013:Embedded-Scalar} show that the
strength of agreement to an \es-sentence on a 7-point Likert-scale
depends on how strong the relevant visual contrast is. When an item
with only some connections is presented alongside two universally
connected elements, the mean rate of agreement with a suitable
\es-sentence is significantly higher than when only one universally
connected element is present. So, maybe our \es-sequences similarly
provoked local readings because of an overemphasized visual contrast
between \emph{some but not all} and \emph{all}.

The third step of our \es-sequences in Figure~\ref{fig:exseqES} is a
relevant case of direct contrast between a \emph{some-but-not-all-}
and an \emph{all}-situation, but it is a weak contrast, in the sense
of \citet{GeurtsTielvan-Tiel2013:Embedded-Scalar}, because there is
only one universally connected element. Moreover, our study elicited
categorical truth-value judgements. Unlike in
\citeauthor{ChemlaSpector2010:Experimental-Ev}'s
(\citeyear{ChemlaSpector2010:Experimental-Ev}) and
\citeauthor{GeurtsTielvan-Tiel2013:Embedded-Scalar}'s
(\citeyear{GeurtsTielvan-Tiel2013:Embedded-Scalar}) experiments, we
did not record degrees of agreement with a statement. For a purely
contrast-based explanation to work, it would have to be made plausible
why the (weak) visual contrast in a picture like in
Figure~\ref{fig:exseqES3} alone is enough to overrule a truth-value
judgment so as to contradict the semantic meaning (even when there is
no prosodic markedness to support such a reinterpretation).

On the other hand, there is a conceivable alternative explanation of
the puzzling data for \es-sentences that would also account for the
large and systematic error responses. Recall from
Section~\ref{sec:results} and Table~\ref{tab:count_data_CC} on page
\pageref{tab:count_data_CC} that a surprisingly high number of
\emph{true} answers was given at the second step of the
\es-sequence. This answer type does not correspond to any of the three
candidate readings that the previous literature has focused on. As
suggested earlier, this answer type can be explained under the
assumption that \emph{exactly one} gets a reading similar to
\emph{there is (at least) one}, while the scalar item \emph{some}
would just receive its semantic interpretation. Such a reading of
\emph{exactly one}, although surprising, could actually be supported
by some theories of numerals and modifiers
\citep[e.g.][]{Geurts2006:Take-five,MartyChemla2014:Phantom-reading},
as the outcome of an ``existential closure'' type-shifting rule in the
sense of \citet{Partee1987:Noun-phrase-int} to the effect that
\emph{exactly one} is mapped onto an existential reading of the form
``there is a group with cardinality exactly one.''

What happens if \emph{some} is additionally strengthened in some way
or other under such a reading of \emph{exactly one}? If \emph{some} is
pragmatically strengthened under the scope of \emph{at least one} in a
pure traditionalist manner by adding the negation of a sentence type
\emph{at least one \dots all \dots}, we would expect \emph{true}
judgements at position 3 of the \es-sequence. Although exactly 3
\emph{true} answers occurred there for each neutral and accented
\es-conditions respectively, the main point to notice is that this
construal does not explain \emph{true} judgements at position 5, which
we classified as local responses. These would therefore still seem
indicative of local readings. However, there is a slight variant of
the hypothesized existentially closed reading of \emph{exactly one},
that would explain \emph{true} judgments at position 5. This is a
reading of \emph{exactly one} as ``there is a \emph{unique} group with
cardinality exactly one.''  Although this is admittedly only a vague
sketch of a possible line of explanation, it is not entirely
implausible that such a uniqueness requirement has participants, in a
first step of evaluation, look for a distinguished referent, which
they can find no sooner than on position 5 when the whole picture is
unravelled. The search for uniqueness would result in a witness for
the existential quantifier that appears to take a pragmatically
strengthened reading of \emph{some} into account, but that is not
necessarily a ``local reading'' in the sense of
grammaticalism. Moreover, although not strictly required to explain
the behavioral data, we can furthermore explain how the predicate
\emph{is connected to some if its \dots} can be pragmatically enriched
in the desired fashion, along the lines of a ``modern traditionalist''
account that makes use of interpretive mechanisms originally developed
to account for certain discourse phenomena
\citep[][Chapter~8.4]{Geurts2010:Quantity-Implic}. The listener would
then conclude that the referent in question was not connected to all
of its surrounding elements, because otherwise the speaker would have
attributed that to the now fixed referent in question. This is
parallel, so Geurts suggests, to the reasoning that an utterance of
``Smith met a woman'' implicates that \emph{the woman referred to} was
not Smith's mother and not that Smith didn't meet his mother at all.

To wrap up, we suggested amendements to grammaticalism and
traditionalism that could help either position explain the data we
observed. Whether these suggestions can be vindicated by further
empirical research is a matter that we must leave open. It remains,
however, that our data provides interesting challenges to both
traditionalism and grammaticalism.

% \paragraph{Critique.} There are a number of conceivable objections to
% our design and our interpretation of the presented results. We will
% consider the most pressing ones in the following.

% To begin with, it could be questioned whether our design reliably
% indicates the preference relations in
% (\ref{bsp:observed-preferences}). One could object that it is not
% feasible to compare results from the preference-related controls to
% results from our target conditions, because the source of different
% readings is fundamentally different. In the former we have a syntactic
% ambiguity, in the latter a set of putative pragmatic enrichments. The
% former is a case of perceived ambiguity (some of our subjects reported
% this), the latter arguably is not
% \citep[see][]{GeurtsPouscoulous2009:Embedded-Implic}. So perhaps even
% if the distribution of responses is indicative of the known preference
% patterns in preference-related controls, this would not necessarily
% mean that the distribution of responses in target conditions is
% indicative of preference relations. Although we cannot rule out that
% the incremental verification task detects preferences among multiple
% readings only for some sources of variability in readings and not
% others, it is by no means clear why this should be so. The burden of
% argument is on the sceptic. But even if we assume that the incremental
% verification task is incapable of reliably indicating preferences
% among candidate pragmatic enrichments of the relevant kind, then it
% should at least be able to indicate availability of readings. So if we
% deny the observed response patterns the power to indicate preference,
% what should the information about availability of readings be that we
% may safely infer from it? It should be that local readings are
% available in \es-sentences and that global readings are unattested in
% \as- and \es-sentences. But that alone is not predicted by either
% traditionalism or grammaticalism. The former cannot explain the
% substantial number of local responses in \es-sentences, the latter
% does not explain the absence of total global responses.

% Another line of criticism that challenges the validity of
% (\ref{bsp:observed-preferences}) targets the Bayesian
% salience-competition model that we used to derive it. We readily admit
% that our model is highly simplistic. Several aspects are especially
% noteworthy here. Firstly, we did not distinguish different types of
% errors, and did not take into account that when errors occurred
% earlier in the sequence, later decision points could no longer be
% reached. This way the model may actually have unduly de-emphasized the
% salience of readings that can be judged later in a sequence. Another,
% more complicated model of the likelihood of answer patterns, might
% therefore have given support to different conclusions. Unfortunately,
% it is not clear how exactly such a more complex generating model
% should be set-up. Eventually, a full processing model, designed for
% data from incremental verification tasks, would be needed. We cannot
% offer such a model here, but can only the offer a first step in that
% direction. More importantly, though, the only part of
% (\ref{bsp:observed-preferences}) that might plausibly change for a
% more encompassing model along the lines sketched above is that the 3
% \emph{false} answers in the \as-conditions with neutral intonation
% might support the conclusion that the local reading is attested. But
% this hypothetical effect will certainly be no bigger than marginal,
% and so no genuinely different conclusions of theoretical relevance
% should be expected.

\section{Conclusions}
\label{sec:conclusions}

We tested experimentally the predictions of traditionalism and
grammaticalism with respect to the availability of and preferences
over three types of conceivable pragmatic enrichments of two types of
sentences, in which scalar \emph{some} occurred in the scope of either
a monotonic quantifier \emph{all} or the non-monotonic quantifier
\emph{exactly one}. To avoid potential confounds of the pictorial
material, such as through typicality or contrast
\citep{Tielvan-Tiel2012:Embedded-Scalar,GeurtsTielvan-Tiel2013:Embedded-Scalar},
we employed an incremental verification task \citep{Conroy2008}, in
which subjects were presented with partially covered pictures and
asked to uncover the picture sequentially until they felt able to give
a categorical truth-value judgements. In order to control for
potential confounds of ``silent prosody,'' sentence material was
presented auditorily and contrastive stress on embedded \emph{some}
was manipulated. This way we were able to also explicitly test the
prosodic markedness hypothesis, which is often used to supplement
traditionalism. Additionally, we included preference-related control
conditions in order to be able to deduce latent preferences for
candidate readings from the answer patterns we gathered. To this end,
we introduced a generative Bayesian model that gives probabilistic
predictions about observable answer patterns in our task, given
concrete instances of relative saliences of readings and other
parameters. Using the observed data, in particular in connection with
the preference-related control conditions, we inferred a posteriori
likely values of the latent model parameters.

Backed up by our analyses, we concluded that both traditionalism and
grammaticalism do not predict our data well. Traditionalism fails to
account for high numbers of responses indicating local readings even
for prosodically unmarked \es-sentences. Grammaticalism, insofar as
it builds on disambiguation by logical strength, fails to predict the
absence of answers indicative of global readings and the relative
abundance of answers indicative of literal readings.

The clearest upshot of theoretical importance of this study is that
strength-based meaning selection does not seem to be a good
disambiguation criterion for grammaticalism. We tentatively suggested
an alternative selection principle for grammaticalism in terms of
``parsing economy.'' Another important conclusion from our data is
that \es-sentences, using non-monotonic quantifier \emph{exactly one},
appeared to be prone to receive a ``shallow interpretation.'' This may
be used by traditionalism to explain answers indicative of local
readings for \es-sentences. But the point is relevant in general. The
possibility of ``shallow interpretation'' of non-monotonic quantifiers
should be considered also when interpreting the results of other
experimental designs \citep[e.g.][]{CliftonDube2010:Embedded-Implic,
  ChemlaSpector2010:Experimental-Ev}.

We have argued that auditory presentation of sentence material should
be controlled for if prosodic information is hypothesized to influence
availability and preferences of readings. We also argued that
theoretical positions must be formulated in such a way that they make
testable predictions about reading preferences where different
potential readings are allowed (contra a grammaticalism without
commitment to any testable disambiguation principle). We should then
try to obtain information about actual preference patterns by explicit
comparison with structures whose preference structures are
well-studied, ideally with some model of the data generating process
for optimal data analysis.

\newpage

\appendix

\section{The auditory sentence material}
\label{sec:audit-sent-mater}

\begin{table}[!hp]
  \centering

  \includegraphics[width=\textwidth]{pictures/Acoustics/Table-B.png}

  \caption{Difference between minimal and maximal F0 values in Hz for
    each of the single words in the target sentences. Region 6
    corresponds to the determiner \emph{einigen}.} 
  \label{tab:table-B}
\end{table}

\begin{table}[!hp]
  \centering

  \includegraphics[width=\textwidth]{pictures/Acoustics/Table-A.png}

  \caption{Durational values in ms for each of the single regions in the target sentences. 
  Region 6 corresponds to the determiner \emph{einigen}.}
  \label{tab:table-A}
\end{table}

\begin{table}[hp]
  \centering

  \includegraphics[width=\textwidth]{pictures/Acoustics/Table-D.png}

  \caption{Difference between minimal and maximal F0 values in Hz  
  for each of the single regions in
    the preference-related controls. Regions 5 and 7 correspond to the
    nouns preceding the boundaries.}  
  \label{tab:table-D}
\end{table}

\begin{table}[hp]
  \centering

  \includegraphics[width=\textwidth]{pictures/Acoustics/Table-C.png}

  \caption{Durational values in ms for each of the single regions in
    the preference-related controls. Regions 5 and 7 correspond to the
    nouns preceding the boundaries.}  
  \label{tab:table-C}
\end{table}

%\paragraph{Targets.} Table~\ref{tab:table-A} shows the durational values for each of
%the single regions in the sentence. Crucially, durational values of
%accented determiners were significantly increased as opposed to their
%non-accented counterparts. Overall, \acro{Quantifier} effects were
%observed, which might be attributed to the above-mentioned lexical
%differences, as well as to the fact that \es-structures generally
%contained fewer words, thus leading to a tendency of a durational
%decrease for each of the single words.
%
%Differences between maximal and minimal F0 values for each of the
%single words in the sentence are depicted in
%Table~\ref{tab:table-B}. As is descriptively evident, accented
%determiners showed a larger F0 range compared to unaccented versions,
%an effect which is confirmed by statistical analyses. An additional
%\acro{Quantifier}*\acro{Prosody} interaction indicates that these
%differences are even more pronounced in the \as-condition. Finally,
%comparable to the durational analyses, \acro{Quantifier} effects were
%observed when conditions exhibited lexical differences.
%
%Finally, differences between maximal and minimal F0 values for each of
%the single words in the sentence are given in
%Table~\ref{tab:table-D}. Again, the most reliable differences occur at
%the boundary regions.

%\paragraph{Preference-related controls.} Table~\ref{tab:table-C} shows the
%durational values of each of the single words in the sentence. As is
%descriptively evident, the largest durational differences were
%realized at the boundary regions (i.e. Region 5 and Region
%7). Interestingly, small differences between conditions also yielded
%significance at other positions, suggesting that our speaker produced
%the different conditions very consistently (i.e., with little
%variance).

\newpage

\printbibliography[heading=bibintoc]

\end{document}