### 2015-11-04 text/paper/revision_02/title_page/titlePage.tex
this version second revision sent

\documentclass[fleqn,reqno,12pt]{article}

%========================================
% Packages
%========================================

\usepackage[]{mypackages}
\bibliography{MyRefGlobal}
\usepackage{mycommands}

\usepackage{array}
% \usepackage{authblk}

\newcommand{\ia}{\acro{ia}}

\title{Embedded Scalars, Preferred Readings \& Intonation: {A}n
  Experimental Revisit} 
\author{Michael Franke, Fabian Schlotterbeck and Petra Augurzky
}

\date{}

\begin{document}
\maketitle

\thispagestyle{empty}

\vspace*{-1cm}
\begin{center}
  Universtit\"at T\"ubingen \\
  Seminar f\"ur Sprachwissenschaft \\
  Wilhelmstra\ss e 19\\
  72072 T\"ubingen\\
  Germany  \\
  \bigskip \href{mailto:mchfranke@gmail.com}{corresponding email: \tt
    mchfranke@gmail.com}
\end{center}

 \vspace*{1cm}

\begin{abstract}
  The scalar item \emph{some} is widely assumed to receive a meaning enrichment to \emph{some
    but not all} if it occurs in matrix position. The question under which circumstances this
  enrichment can occur in certain embedded positions plays an important role in deciding how to
  delineate semantics and pragmatics.  We present new experimental data that bear on this
  theoretical issue. In distinction to previous experimental approaches, we presented sentence
  material auditorily in order to explicitly control prosodic markedness of the scalar
  item. Moreover, our experiment sheds light on the relative preferences or salience of
  candidate readings. The presented data turn out to be unexpected under a traditional Gricean
  view, but also challenge the idea of disambiguation by logical strength in grammaticalist
  approaches.
\end{abstract}

\end{document}